---
title: "medfate - Mediterranean forest functioning"
output: github_document
---

```{r setup, echo = FALSE}

knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "README-")
```

## Introduction

Package **medfate** is designed to assist forest scientists to simulate forest functioning and dynamics, using cohort-based description of forest stands. The package provides functions to simulate the following processes:

+ Soil water balance (De Cáceres et al. 2015)
+ Plant hydraulics, transpiration and photosynthesis
+ Plant growth

The models are parameterized for species of the Mediterranean region (particularly for Catalonia, NE Spain), but forests with different composition could be modelled with different parameter sets.

## Package installation

Package **medfate** can be found at [CRAN](https://CRAN.R-project.org/package=medfate), where it is updated every few months. Users can also download and install the latest stable versions GitHub as follows (required package `devtools` should be installed/updated first):

```{r installation, eval = FALSE}
devtools::install_github("vegmod/medfate")
```

Additionally, users can have help to run package functions directly as package vignettes, by forcing their inclusion in installation: 
```{r installation vignettes, eval = FALSE}
devtools::install_github("vegmod/medfate", 
                         build_opts = c("--no-resave-data", "--no-manual"))
```

## Documentation

+ The package installation includes a number of *vignettes* that illustrate how to run simulation models in **medfate**. 

+ Additional documentation of the package, including a *reference book*, can be found at https://vegmod.github.io/rpackage/medfate. 

+ An interactive example application showing the behaviour of the water balance model can be found at http://vegmod.ctfc.cat/medfateweb.

## References

+ De Cáceres M, Martínez-Vilalta J, Coll L, Llorens P, Casals P, Poyatos R, Pausas JG, Brotons L. (2015) Coupling a water balance model with forest inventory data to predict drought stress: the role of forest structural changes vs. climate changes. Agricultural and Forest Meteorology 213: 77-90 (doi:10.1016/j.agrformet.2015.06.012).

