---
title: "Forest growth"
author: "Miquel De Caceres"
date: "`r Sys.Date()`"
description: >
   Learn how to conduct simulations including carbon balance, plant growth and mortality
output: 
  rmarkdown::html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(medfate)
```

## About this vignette

This document describes how to run the growth model of `medfate`, described in De Cáceres et al. (2023) and implemented in function `growth()`. All the details of the model design and formulation can be found at the corresponding chapters of the [medfate book](https://emf-creaf.github.io/medfatebook/index.html). 

Because the forest growth model builds on the water balance model, the reader is assumed here to be familiarized with `spwb()`. If not, we recommend reading vignette [*Basic water balance*](https://emf-creaf.github.io/medfate/articles/runmodels/BasicWaterBalance.html) before this one. 

## Preparing model inputs

Model inputs are explained in greater detail in vignettes [*Understanding model inputs*](https://emf-creaf.github.io/medfate/articles/intro/UnderstandingInputs.html) and [*Preparing model inputs*](https://emf-creaf.github.io/medfate/articles/intro/PreparingInputs.html). Here we briefly review the different steps required to run function `growth()`.

### Soil, vegetation, meteorology and species data

Soil physical characteristics needs to be specified using a `data frame` with soil layers in rows and physical attributes in columns. Soil physical attributes can be initialized to default values, for a given number of layers, using function `defaultSoilParams()`:

```{r}
examplesoil <- defaultSoilParams(4)
examplesoil
```

As explained in the package overview, models included in `medfate` were primarily designed to be ran on **forest inventory plots**. Here we use the example forest object provided with the package:
```{r}
data(exampleforest)
exampleforest
```

Importantly, a data frame with daily weather for the period to be simulated is required. Here we use the default data frame included with the package: 

```{r}
data(examplemeteo)
head(examplemeteo)
```

The weather variables required by the `growth()` function depend on the complexity of the water balance simulations underlying growth (i.e. on the control parameter `transpirationMode`, see below). 
 
Finally, all simulations in `medfate` require a data frame with species parameter values, for which we load using defaults for Catalonia (NE Spain):

```{r}
data("SpParamsMED")
```

### Simulation control

Apart from data inputs, the behaviour of simulation models can be controlled using a set of global parameters. The default parameterization is obtained using function `defaultControl()`:

```{r}
control = defaultControl("Granier")
```
Here we will run growth simulations using the basic water balance model (i.e. `transpirationMode = "Granier"`). The complexity of the soil water balance calculations can be changed by using `"Sperry"` as input to `defaultControl()`. 


### Growth input object

A last object, called `growthInput`, needs to be created before calling the simulation function. This is analogous to `spwbInput` and consists in the compilation of soil and cohort-level parameters needed for simulations. The object can be obtained by using function `growthInput()`:

```{r}
x <- growthInput(exampleforest, examplesoil, SpParamsMED, control)
```

All the input information for forest data and species parameter values can be inspected by printing different elements of the input object, whose names are: 
```{r}
names(x)
```
As with `spwbInput` objects, information about the cohort species is found in element `cohorts`  (i.e. code, species and name):
```{r}
x$cohorts
```

Element `above` contains the above-ground structure data that we already know, but with an additional columns that describes the estimated initial amount of *sapwood area*:
```{r}
x$above
```

Elements starting with `params*` contain cohort-specific model parameters. Some of them were already presented in previous vignettes ([*Basic water balance*](https://emf-creaf.github.io/medfate/articles/runmodels/BasicWaterBalance.html) and [*Advanced water/energy balance*](https://emf-creaf.github.io/medfate/articles/runmodels/AdvancedWaterEnergyBalance.html)). An important set of new cohort-specific parameters for the forest growth model are `paramsGrowth`:
```{r}
x$paramsGrowth
```
which includes maximum growth rates, senescence rates and maintenance respiration rages. Another important set of parameters is given in `paramsAllometries`:
```{r}
x$paramsAllometries
```
Note that in the previous models, allometries were already used to estimate above-ground structural parameters, but these were static during simulations. 

Elements starting with `internal*` contain state variables required to keep track of plant status. For example, the metabolic and storage carbon levels can be seen in `internalCarbon`:
```{r}
x$internalCarbon
```
and `internalAllocation` stores the carbon allocation targets:

```{r}
x$internalAllocation
```
Additional `internal*` elements are `internalMortality`, used to keep track of dead individuals; and `internalRings`, which stores state variables used to model sink limitations on wood formation.

## Executing the growth model

Having all the input information we are ready to call function `growth()`, which has the same parameter names as `spwb()`:
```{r}
G1<-growth(x, examplemeteo, latitude = 41.82592, elevation = 100)
```

At the end of daily simulations, the `growth()` function displays information regarding the carbon and water balance, which is mostly useful to check that balances are closed. 


Function `growth()` returns an object of class with the same name, actually a list:
```{r}
class(G1)
```
If we inspect its elements, we realize that some of them are the same as returned by `spwb()`:
```{r}
names(G1)
```

Some elements are common with the output of `spwb()`. In particular, `growthInput` contains a copy of the input object, whereas `growthOutput` contains the same object, but with values of state variables at the end of simulation. The new list elements, with respect to the output of function `spwb()`, are `LabileCarbonBalance` (components of the labile carbon balance), `PlantBiomassBalance` (plant- and cohort-level biomass balance), `PlantStructure` (daily series of structural variables) and `GrowthMortality` (daily growth and mortality rates).


## Inspecting model outputs

Users can extract, summarize or inspect the output of `growth()` simulations as done for simulations with `spwb()`. 

### Extracting output

Function `extract()` allow extracting model outputs in form of data frame:
```{r}
extract(G1, "forest", addunits = TRUE) |>
  tibble::as_tibble()
```
These data frames are easy handle in R or can be written into text files for post-processing with other programs.

### Plots

Several plots are available, in addition to all the plots that were available to display the results of `spwb()` simulations. Some of them are illustrated in the following subsections:

To inspect components of the plant carbon balance we can first display daily gross photosynthesis expressed as the carbon fixation relative to dry biomass:
```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=4, fig.alt = "Gross photosynthesis"}
plot(G1, "GrossPhotosynthesis", bySpecies = T)
```
Then we can draw the maintenance respiration costs (which include the sum of leaf, sapwood and fine root respiration) in the same units:
```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=4}
plot(G1, "MaintenanceRespiration", bySpecies = T)
```
Finally we can display the daily negative or positive balance of the plant storage, which determines changes in plant carbon pools:
```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=4}
plot(G1, "LabileCarbonBalance", bySpecies = T)
```

Carbon assimilation and respiration rates define the dynamics of stored carbon. The most important storage compartment is sapwood starch, whose dynamics can be shown using:
```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=4}
plot(G1, "StarchSapwood", bySpecies = T)
```

Leaf and sapwood area dynamics arising from the interplay between growth and senescence of tissues can be inspected using:
```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=4}
plot(G1, "LeafArea", bySpecies = T)
plot(G1, "SapwoodArea", bySpecies = T)
```

We can also inspect carbon balance at the stand level, here displayed at monthly resolution:
```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=4}
plot(G1, "CarbonBalance", summary.freq = "months")
```

### Interactive plots
Finally, recall that one can interactively create plots using function `shinyplot`, e.g.:
```{r, eval = FALSE}
shinyplot(G1)
```

## Growth evaluation

Evaluation of growth simulations will normally imply the comparison of predicted vs observed basal area increment (BAI) or diameter increment at a given temporal resolution. 

Here, we illustrate the evaluation functions included in the package using a fake data set, consisting on the predicted values and some added error.
```{r}
data(exampleobs)
```

Normally growth evaluations will be at annual scale, but here we only have one year of simulated growth. Assuming we want to evaluate the predictive capacity of the model in terms of monthly basal area increment for the pine cohort, we can plot the relationship between observed and predicted values using function `evaluation_plot()`:
```{r, fig=TRUE, echo=TRUE, fig.width=4, fig.height=4}
evaluation_plot(G1, exampleobs, "BAI", cohort = "T1_148", 
                temporalResolution = "month", plotType = "scatter")
```

And the following would help us quantifying the strength of the relationship:

```{r}
evaluation_stats(G1, exampleobs, "BAI", cohort = "T1_148", 
                 temporalResolution = "month")
```

The observed data set is fake and the evaluation is unrealistically good. For illustrative purposes, we also compare diameter increment values, here drawing the observed and predicted time series together: 

```{r, echo=TRUE, fig.width=7, fig.height=4}
evaluation_plot(G1, exampleobs, "DI", cohort = "T1_148", 
                temporalResolution = "day")
```

Again, actual comparisons will be done at coarser temporal resolution. For convenience, function `shinyplot()` also accepts an observed data frame as second argument, which allows performing model evaluation interactively:

```{r, eval = FALSE}
shinyplot(G1, exampleobs)
```

## References

+ De Cáceres M, Molowny-Horas R, Cabon A, Martínez-Vilalta J, Mencuccini M, García-Valdés R, Nadal-Sala D, Sabaté S, Martin-StPaul N, Morin X, D'Adamo F, Batllori E, Améztegui A (2023) MEDFATE 2.9.3: A trait-enabled model to simulate Mediterranean forest function and dynamics at regional scales. Geoscientific Model Development 16: 3165-3201 (https://doi.org/10.5194/gmd-16-3165-2023).
