---
title: "Forest dynamics"
author: "Miquel De Caceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(medfate)
```

## About this vignette

This document describes how to run the forest dynamics model of `medfate`, described in De Cáceres et al. (2023) and implemented in function `fordyn()`. This document is meant to teach users to run the simulation model with function `fordyn()`. Details of the model design and formulation can be found at the corresponding chapters of the [medfate book](https://emf-creaf.github.io/medfatebook/index.html). 

Because the model builds on the growth and water balance models, the reader is assumed here to be familiarized with `spwb()` and `growth()` (otherwise read vignettes [*Basic water balance*](https://emf-creaf.github.io/medfate/articles/runmodels/BasicWaterBalance.html) and [*Forest growth*](https://emf-creaf.github.io/medfate/articles/runmodels/ForestGrowth.html)). 



## Preparing model inputs
Any forest dynamics model needs information on climate, vegetation and soils of the forest stand to be simulated. Moreover, since models in `medfate` differentiate between species, information on species-specific model parameters is also needed. In this subsection we explain the different steps to prepare the data needed to run function `fordyn()`.

Model inputs are explained in greater detail in vignettes [*Understanding model inputs*](https://emf-creaf.github.io/medfate/articles/intro/UnderstandingInputs.html) and [*Preparing model inputs*](https://emf-creaf.github.io/medfate/articles/intro/PreparingInputs.html). Here we only review the different steps required to run function `fordyn()`.

### Soil, vegetation, meteorology and species data

Soil information needs to be entered as a `data frame` with soil layers in rows and physical attributes in columns. Soil physical attributes can be initialized to default values, for a given number of layers, using function `defaultSoilParams()`:

```{r}
spar = defaultSoilParams(4)
```

The soil input for water balance simulation is actually a list of class `soil` that is created using a function with the same name:
```{r}
examplesoil = soil(spar)
```

As explained in the package overview, models included in `medfate` were primarily designed to be ran on **forest inventory plots**. Here we use the example object provided with the package:
```{r}
data(exampleforestMED)
exampleforestMED
```

Importantly, a data frame with daily weather for the period to be simulated is required. Here we use the default data frame included with the package: 
```{r}
data(examplemeteo)
head(examplemeteo)
```

Finally, simulations in `medfate` require a data frame with species parameter values, which we load using defaults for Catalonia (NE Spain):

```{r}
data("SpParamsMED")
```

### Simulation control

Apart from data inputs, the behaviour of simulation models can be controlled using a set of global parameters. The default parameterization is obtained using function `defaultControl()`:

```{r}
control = defaultControl("Granier")
```
Here we will run simulations of forest dynamics using the basic water balance model (i.e. `transpirationMode = "Granier"`). The complexity of the soil water balance calculations can be changed by using `"Sperry"` as input to `defaultControl()`. However, when running `fordyn()` sub-daily output will never be stored (i.e. setting `subdailyResults = TRUE` is useless).

## Executing the forest dynamics model

In this vignette we will fake a ten-year weather input by repeating the example weather data frame ten times.
```{r}
meteo = rbind(examplemeteo, examplemeteo, examplemeteo, examplemeteo,
                    examplemeteo, examplemeteo, examplemeteo, examplemeteo,
                    examplemeteo, examplemeteo)
row.names(meteo) = seq(as.Date("2001-01-01"), 
                             as.Date("2010-12-29"), by="day")
```


Now we run the forest dynamics model using all inputs (note that no intermediate input object is needed, as in `spwb()` or  `growth()`):
```{r}
fd<-fordyn(exampleforestMED, examplesoil, SpParamsMED, meteo, control, 
           latitude = 41.82592, elevation = 100)
```
It is worth noting that, while `fordyn()` calls function `growth()` internally for each simulated year, the `verbose` option of the control parameters only affects function `fordyn()` (i.e. all console output from `growth()` is hidden). Recruitment and summaries are done only once a year at the level of function `fordyn()`.

## Inspecting model outputs

### Stand, species and cohort summaries and plots

Among other outputs, function `fordyn()` calculates standard summary statistics that describe the structural and compositional state of the forest at each time step. For example, we can access stand-level statistics using:

```{r}
fd$StandSummary
```
Species-level analogous statistics are shown using:
```{r}
fd$SpeciesSummary
```

Package `medfate` provides a simple `plot` function for objects of class `fordyn`. For example, we can show the interannual variation in stand-level basal area using:

```{r, fig=TRUE, fig.align="center", fig.width=7, fig.height = 3.5}
plot(fd, type = "StandBasalArea")
```

### Tree/shrub tables

Another useful output of `fordyn()` are tables in long format with cohort structural information (i.e. DBH, height, density, etc) for each time step:

```{r}
fd$TreeTable
```

The same can be shown for dead trees:
```{r}
fd$DeadTreeTable
```

### Accessing the output from function growth()

Since function `fordyn()` makes internal calls to function `growth()`, it stores the result in a vector called `GrowthResults`, which we can use to inspect intra-annual patterns of desired variables. For example, the following shows the leaf area for individuals of the three cohorts during the second year:

```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=3}
plot(fd$GrowthResults[[2]], "LeafArea", bySpecies = T)
```
Instead of examining year by year, it is possible to plot the whole series of results by passing a `fordyn` object to the `plot()` function:

```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=3}
plot(fd, "LeafArea")
```

Finally, we can create interactive plots for particular steps using function `shinyplot()`, e.g.:
```{r, eval = FALSE}
shinyplot(fd$GrowthResults[[1]])
```

## Forest dynamics including management
The package allows including forest management in simulations of forest dynamics. This is done in a very flexible manner, in the sense that `fordyn()` allows the user to supply an arbitrary function implementing a desired management strategy for the stand whose dynamics are to be simulated. The package includes, however, an in-built default function called `defaultManagementFunction()` along with a flexible parameterization, a list with defaults provided by function `defaultManagementArguments()`. 

Here we provide an example of simulations including forest management:

```{r}
# Default arguments
args <- defaultManagementArguments()
# Here one can modify defaults before calling fordyn()
#
# Simulation
fd<-fordyn(exampleforestMED, examplesoil, SpParamsMED, meteo, control, 
           latitude = 41.82592, elevation = 100,
           management_function = defaultManagementFunction,
           management_args = args)
```
When management is included in simulations, two additional tables are produced, corresponding to the trees and shrubs that were cut, e.g.:
```{r}
fd$CutTreeTable
```

## References

+ De Cáceres M, Molowny-Horas R, Cabon A, Martínez-Vilalta J, Mencuccini M, García-Valdés R, Nadal-Sala D, Sabaté S, Martin-StPaul N, Morin X, D'Adamo F, Batllori E, Améztegui A (2023) MEDFATE 2.9.3: A trait-enabled model to simulate Mediterranean forest function and dynamics at regional scales. Geoscientific Model Development (https://doi.org/10.5194/gmd-2022-243).