---
title: "Example_SurEau"
author: "Miquel De Caceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Preparation

Initialisation for `spwb()` function: 

```{r }
library(medfate)

# Soil
examplesoil <- defaultSoilParams(n = 4)

# Load example plot plant data
data(exampleforest)

# Prepare input for 'Sureau' transpiration mode
control <- defaultControl("Sureau")

# Create input for spwb
x <- forest2spwbInput(exampleforest,examplesoil,SpParamsMED, control)

```

### Sureau network input

Create Sureau network (similar) for each of the three plant cohort in `x`. This returns a list with three elements:
```{r}
cn <- initSureauNetworks(x)
length(cn)
```

You can inspect the structure of the network for the first object using:

```{r}
network <- cn[[1]]
str(network)
```

### Call to `semi_implicit_integration()`

If we modify the values of `Elim`, `Emin_L`  or `Emin_S`, then a call to `semi_implicit_integration()` should update the water potentials according to these flows. The water potentials prior to the call are:
```{r}
network$Psi_LApo
network$Psi_LSym
network$Psi_SApo
network$Psi_SSym
```

We now set the options and time step size:
```{r}
opt <- c("Lsym" = 1.0, "Ssym" = 1.0, 
         "Eord" = 1.0, "Lcav" = 1.0, "Scav" = 1.0,
         "CLapo" = 1.0, "CTapo" = 1.0)

# Number of small steps per hour sub-steps
nsmalltimesteps <- c(2, 4, 8, 16)

# Seconds of the small time step (four sub-steps per hour)
dt <- 3600.0 / nsmalltimesteps[2] 
```

We also specify the flows and call `semi_implicit_integration()`:
```{r}
network$Elim <- 0.1
network$Emin_L <- 0.01
network$Emin_S <- 0.01

semi_implicit_integration(network, dt, opt)
```

The function modifies the `network` object, whose water potentials are now:

```{r}
network$Psi_LApo
network$Psi_LSym
network$Psi_SApo
network$Psi_SSym
```


