---
title: "Forest dynamics"
author: "Miquel De Caceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{Forest dynamics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignettePackage{medfate}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(medfate)
```

## About this vignette

This document describes how to run the forest dynamics model of `medfate`. This document is meant to teach users to run the simulation model within R. Details of the model design and formulation can be found at https://emf-creaf.github.io/medfatebook/index.html. 

## Preparing model inputs
Any forest dynamics model needs information on climate, vegetation and soils of the forest stand to be simulated. Moreover, since models in `medfate` differentiate between species, information on species-specific model parameters is also needed. In this subsection we explain the different steps to prepare the data needed to run function `fordyn()`.

Model inputs are explained in greater detail in vignette '*Simulation inputs*'. Here we only review the different steps required to run function `fordyn()`.

### Soil, vegetation, meteorology and species data

Soil information needs to be entered as a `data frame` with soil layers in rows and physical attributes in columns. Soil physical attributes can be initialized to default values, for a given number of layers, using function `defaultSoilParams()`:

```{r}
spar = defaultSoilParams(2)
```

The soil input for water balance simulation is actually a list of class `soil` that is created using a function with the same name:
```{r}
examplesoil = soil(spar)
```

As explained in the package overview, models included in `medfate` were primarily designed to be ran on **forest inventory plots**. Here we use the example object provided with the package:
```{r}
data(exampleforestMED)
exampleforestMED
```

In the basic water balance, only mean temperature, precipitation and potential evapotranspiration is required, but radiation may also be necessary to simulate snow melt. 
```{r}
data(examplemeteo)
head(examplemeteo)
```

Finally, simulations in `medfate` require a data frame with species parameter values, which we load using defaults for Catalonia (NE Spain):

```{r}
data("SpParamsMED")
```

### Simulation control

Apart from data inputs, the behaviour of simulation models can be controlled using a set of global parameters. The default parameterization is obtained using function `defaultControl()`:

```{r}
control = defaultControl("Granier")
```
Here we will run simulations of forest dynamics using the basic water balance model (i.e. `transpirationMode = "Granier"`). The complexity of the soil water balance calculations can be changed by using `"Sperry"` as input to `defaultControl()`. However, when running `fordyn()` sub-daily output will never be stored (i.e. setting `subdailyResults = TRUE` is useless).

## Executing the forest dynamics model

In this vignette we will fake a eight-year weather input by repeating the example weather data frame eight times, while reducing precipitation of every two years by 60\%.
```{r}
meteowet = examplemeteo
meteodry = examplemeteo
meteodry$Precipitation = meteowet$Precipitation*0.6
meteo_01_08 = rbind(meteowet, meteodry, meteowet, meteodry,
                    meteowet, meteodry, meteowet, meteodry)
row.names(meteo_01_08) = seq(as.Date("2001-01-01"), 
                             as.Date("2008-12-29"), by="day")
```


Now we run the forest dynamics model using all inputs (note that no intermediate input object is needed, as in `spwb()` or  `growth()`):
```{r}
fd<-fordyn(exampleforestMED, examplesoil, SpParamsMED, meteo_01_08, control, 
           latitude = 41.82592, elevation = 100)
```
It is worth noting that, while `fordyn()` calls function `growth()` internally for each simulated year, the `verbose` option of the control parameters only affects function `fordyn()` (i.e. all console output from `growth()` is hidden). Recruitment and summaries are done only once a year at the level of function `fordyn()`.

## Inspecting model outputs

### Stand, species and cohort summaries and plots

Among other outputs, function `fordyn()` calculates standard summary statistics that describe the structural and compositional state of the forest at each time step. For example, we can access stand-level statistics using:

```{r}
fd$StandSummary
```

where we can observe an increase in stand basal area and leaf area index during years 1 and 3, but a reduction during the second (drier) year. Species-level analogous statistics are shown using:
```{r}
fd$SpeciesSummary
```

Package `medfate` provides a simple `plot` function for objects of class `fordyn`. For example, we can show the interannual variation in stand-level basal area using:

```{r, fig=TRUE, fig.align="center", fig.width=7, fig.height = 3.5}
plot(fd, type = "StandBasalArea")
```

### Tree/shrub tables

Another useful output of `fordyn()` are tables in long format with cohort structural information (i.e. DBH, height, density, etc) for each time step:

```{r}
fd$TreeTable
```

The same can be shown for dead trees, where we see larger mortality rates during the second (drier) year:
```{r}
fd$DeadTreeTable
```

### Accessing the output from function growth()

Since function `fordyn()` makes internal calls to function `growth()`, it stores the result in a vector called `GrowthResults`, which we can use to inspect intra-annual patterns of desired variables. For example, the following shows the leaf area for individuals of the three cohorts during the three consecutive years:

```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=3}
plot(fd$GrowthResults[[1]], "LeafArea", bySpecies = T)
plot(fd$GrowthResults[[2]], "LeafArea", bySpecies = T)
plot(fd$GrowthResults[[3]], "LeafArea", bySpecies = T)
```
where we see that the second (drier) year, resulted in a decrease in leaf area per individual, due to the sink limitation to leaf area growth caused by drought. Instead of examining year by year, it is possible to plot the whole series of results by passing a `fordyn` object to the `plot()` function:

```{r, fig=TRUE, echo=TRUE, fig.width=7, fig.height=3}
plot(fd, "LeafArea")
```
where we can see the leaf area drop of the second year along with the progressive recovery of the following years. 

Finally, we can create interactive plots for particular steps using function `shinyplot()`, e.g.:
```{r, eval = FALSE}
shinyplot(fd$GrowthResults[[1]])
```

## Forest dynamics including management
The package allows including forest management in simulations of forest dynamics. This is done in a very flexible manner, in the sense that `fordyn()` allows the user to supply an arbitrary function implementing a desired management strategy for the stand whose dynamics are to be simulated. The package includes, however, an in-built default function called `defaultManagementFunction()` along with a flexible parameterization, a list with defaults provided by function `defaultManagementArguments()`. 

Here we provide an example of simulations including forest management:

```{r}
# Default arguments
args <- defaultManagementArguments()
# Here one can modify defaults before calling fordyn()
#
# Simulation
fd<-fordyn(exampleforestMED, examplesoil, SpParamsMED, meteo_01_08, control, 
           latitude = 41.82592, elevation = 100,
           management_function = defaultManagementFunction,
           management_args = args)
```
When management is included in simulations, two additional tables are produced, corresponding to the trees and shrubs that were cut, e.g.:
```{r}
fd$CutTreeTable
```
