---
title: "Understanding model inputs"
author: "Miquel De Caceres"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, include=FALSE}
library(medfate)
```

## About this article

Any process-based model of forest functioning and dynamics needs information on **climate**, **vegetation** and **soils** of the forest stand to be simulated. Moreover, since `medfate` allows simulating cohorts belonging to different species,  **species-specific parameters** are also needed. Finally, **simulation control parameters** may need to be changed, depending on the goals of the simulation. This article explains data structures required as input to run simulations using the package so that the user can understand them. A companion article [*Preparing model inputs*](https://emf-creaf.github.io/medfate/articles/intro/PreparingInputs.html) provides a practical example to illustrate how to create model inputs and some common problems encountered.


## Species parameter tables

Simulation models in `medfate` require a data frame with species (taxon) parameter values. The package includes a default data sets to be readily used. The values of the parameter table were obtained from global trait data bases, bibliographic searches, fit to empirical data or expert-based guesses:

```{r}
data("SpParamsMED") # For the Spanish forest inventory (including taxon groups)
```

 A large number of parameters (columns) can be found in species parameter tables. Not all parameters are needed for all models. You can find parameter definitions in table `SpParamsDefinition`, which we reproduce below:

```{r, echo = FALSE}
data("SpParamsDefinition")
knitr::kable(SpParamsDefinition[,-2])
```

In order to understand the role of parameters in the model, you should read the details of model design and formulation included in the [**medfatebook**](https://emf-creaf.github.io/medfatebook/index.html). Details regarding how the species parameter tables are build can be found in article [*Species parameterization*](https://emf-creaf.github.io/medfate/articles/parametrization/SpeciesParameterization.html). 

## Vegetation

### Forest objects

Models included in `medfate` were primarily designed to be ran on **forest inventory plots**. In this kind of data, the vegetation of a sampled area is often described by several records of woody plants (trees and shrubs) along with their size and species identity. Forest plots in `medfate` are assumed to be in a data structure that follows closely the Spanish national forest inventory, but is simple enough to so that other forest sampling schemes can be mapped onto it. 

Each forest plot is represented in an object of class `forest`, a list that contains several elements. Among them, the most important items are two data frames, `treeData` (for trees) and `shrubData` (for shrubs):

```{r}
data(exampleforest)
exampleforest
```

Trees are expected to be primarily described in terms of species, diameter (DBH; cm) and height (cm), whereas shrubs are described in terms of species, percent cover (%) and mean height (cm). Root distribution has to be specified for both growth forms, in terms of the depths (mm) corresponding to 50% and 95% of cumulative fine root distribution. Functions are provided in package [medfateutils](https://emf-creaf.github.io/medfateutils/) to map variables in user data frames into tables `treeData` and `shrubData`. Information about the herb layer may be either absent or included in an aggregated way (i.e. without distinguishing cohorts).

While the former example illustrates the standard structure of a `forest` object, users may use an alternative description, based on leaf area index and crown ratio of woody cohorts and the herb layer:

```{r}
data(exampleforest2)
exampleforest2
```

This alternative `forest` form is suitable for water balance simulations, but does not allow simulating forest dynamics.

### Single-cohort forests


Although `medfate` has been designed to perform simulations on multi-cohort forests, it can also handle simulations where vegetation is described using a single cohort. Functions `tree2forest()` and `shrub2forest()` allow defining single-cohort forests from attributes. For example a holm oak (*Quercus ilex*) forest of 4-m height and having a leaf area index of $2\, m^2\cdot m^{-2}$ can be defined using:

```{r}
oak_forest <-tree2forest("Quercus ilex", Height= 400, LAI = 2)
```

The function will return a `forest` object where most attributes are empty:
```{r}
oak_forest
```

Since density and diameter have not been provided, simulations in this case will be restricted to water balance. Moreover, note that when defining single-cohort forests all possible interactions with functionally distinct plants are neglected.


### Aboveground and belowground data

We can use some functions to inspect how above-ground and below-ground information is represented in `medfate`. 

For example, we can use function `forest2aboveground()` on the object `exampleforest` to show how medfate completes above-ground information:
```{r}
above = forest2aboveground(exampleforest, SpParamsMED)
above
```
Note that the call to `forest2aboveground()` included the species parameter table, because species-specific allometric coefficients are needed to calculate leaf area from tree size or shrub percent cover and height. Moreover, note that the plant cohorts were given unique codes that tell us whether they correspond to trees ('T') or shrubs ('S'). 

Columns `N`, `DBH` and `Cover` describe forest structure and are required for simulating growth, but not for soil water balance, which only requires columns `SP`, `H` (in cm), `CR` (i.e. the crown ratio), `LAI_live`, `LAI_expanded` and `LAI_dead`. Therefore, one could use alternative forest description as starting point, i.e.:
```{r}
above2 = forest2aboveground(exampleforest2, SpParamsMED)
above2
```

Of course, the resulting data frame has missing values, whereas the other values are directly copied from `forest`.

Aboveground leaf area distribution (with or without distinguishing among cohorts) can be examined by calling function `vprofile_leafAreaDensity()`:

```{r, fig = TRUE, fig.width= 4, fig.height=3, fig.align= 'center', echo=TRUE}
vprofile_leafAreaDensity(exampleforest, SpParamsMED, byCohorts = F)
```
```{r, fig = TRUE, fig.width= 5, fig.height=3, fig.align= 'center', echo=TRUE}
vprofile_leafAreaDensity(exampleforest, SpParamsMED, byCohorts = T)
```

### Belowground data

Regarding **belowground** information, we need vectors with depths corresponding to 50% and 95% of fine roots, which we simply concatenate from our forest data:
```{r}
Z50 = c(exampleforest$treeData$Z50, exampleforest$shrubData$Z50)
Z95 = c(exampleforest$treeData$Z95, exampleforest$shrubData$Z95)
```
These parameters specify a continuous distribution of fine roots. Users can visually inspect the distribution of fine roots of `forest` objects by calling function `vprofile_rootDistribution()`:

```{r, fig = TRUE, fig.width= 5, fig.height=3, fig.align= 'center', echo=TRUE}
vprofile_rootDistribution(exampleforest, SpParamsMED)
```


## Soils

### Soil physical description

Simulation models in `medfate` require information on the physical attributes of soil, namely soil depth, texture, bulk density and rock fragment content. Soil physical attributes can be initialized to default values, for a given number of layers, using function `defaultSoilParams()`:

```{r}
spar <- defaultSoilParams(4)
print(spar)
```

where `widths` are soil layer widths in mm; `clay` and `sand` are the percentage of clay and sand, in percent of dry weight, `om` stands for organic matter, `bd` is bulk density (in $g \cdot cm^{-3}$) and `rfc` the percentage of rock fragments. Because soil properties vary strongly at fine spatial scales, ideally soil physical attributes should be measured on samples taken at the forest stand to be simulated. For those users lacking such data, soil properties are available via SoilGrids.org.

### Soil input object

The soil input for simulations is an object of class `soil` (also a data frame) that is created using a function with the same name:

```{r}
examplesoil = soil(spar)
class(examplesoil)
```

In addition to the physical soil description, this object contains soil parameters needed for soil water balance simulations:
```{r}
examplesoil
```
For example, `macro` specifies the macroporosity of each layer. The meaning of all elements in the soil object can be found in the help page for function `soil()`.

At any time, one can show the characteristics and status of the soil object using its `summary` function:
```{r}
summary(examplesoil, model = "SX")
```

Importantly, the soil object is used to store the degree of moisture of each soil layer. In particular, element `W` contains the state variable that represents moisture content - the proportion of moisture **relative to field capacity** - which is normally initialized to 1 for each layer:
```{r}
examplesoil$W
```

Advanced soil plant energy and water balance modelling requires considering the temperature of soil. Hence, `Temp` contains the temperature (in degrees) of soil layers:
```{r}
examplesoil$Temp
```
Soil layer temperatures are initialized to missing values, so that at the first time step they will be set to atmospheric temperature. While simple water balance modeling can be run using either Saxton's or Van Genuchten's equations as water retention curves, Van Genuchten's model is forced for advanced modelling.

### Water retention curves

The modelled moisture content of the soil depends on the **water retention curve** used to represent the relationship between soil volumetric water content ($\theta$; %) and soil water potential ($\Psi$; MPa). By default the Saxton (`model = "SX"`) equations are used to model the water retention curve, but the user may choose to follow Van Genuchten - Mualem equations, which will give slightly different values for the same texture:
```{r}
summary(examplesoil, model="VG")
```
While Saxton equations use texture and organic matter as inputs, the Van Genuchten-Mualem equations need other parameters, which are estimated using pedotransfer functions and their names start with `VG_` (two alternative options are provided in function `soil` to estimate Van Genuchten parameters). The following code calls function `soil_retentionCurvePlot()` to illustrate the difference between the two water retention models in this soil:

```{r, fig = TRUE, fig.width= 5, fig.height=3, fig.align= 'center', echo=TRUE}
soil_retentionCurvePlot(examplesoil, model="both")
```

Low-level functions, such as `soil_psi2thetaSX()` and `soil_psi2thetaVG()` (and their counterparts `soil_theta2psiSX()` and `soil_theta2psiVG()`), can be used to calculate volumetric soil moisture from the water potential (and viceversa) using the two models. When simulating soil water balance, the user can choose among the two models (see `control` parameters below).


## Meteorological forcing

All simulations in the package require daily weather inputs. The minimum weather variables that are required are minimum/maximum temperature, minimum/maximum relative humidity, precipitation and radiation. Other variables like wind speed are recommended but not required. Here we show an example of meteorological forcing data. 
```{r}
data(examplemeteo)
head(examplemeteo)
```
Simulation models in **medfate** have been designed to work along with data generated from package [meteoland](https://emf-creaf.github.io/meteoland/). The user is strongly recommended to resort to this package to obtain suitable weather input for medfate simulations.


## Simulation control

Apart from data inputs, the behaviour of simulation models can be controlled using a set of global parameters. The default parameterization is obtained using function `defaultControl()`:

```{r}
control = defaultControl()
names(control)
```

Control parameters should normally be left to their default value until their effect on simulations is fully understood.

## Input objects for simulation functions

Simulation functions `spwb()` and `growth()` (and similar functions) require first combining forest, soil, species-parameter and simulation control inputs into a single input object (of class `spwbInput` or `growthInput`) that is then used as input to the corresponding simulation function along with weather data. The combination of vegetation, soil and control inputs is done via functions  `forest2spwbInput()` and `forest2growthInput()`. While it complicates the code, having this additional step is handy because cohort-level parameters and state variables initialized can then be modified by the user (or an automated calibration algorithm) before calling the actual simulation functions. The input objects for functions `spwb()` and `growth()` are presented in more detail in articles [*Basic water balance*](https://emf-creaf.github.io/medfate/articles/runmodels/BasicWaterBalance.html) and [*Forest growth*](https://emf-creaf.github.io/medfate/articles/runmodels/ForestGrowth.html), respectively.

Function `fordyn()` is different from the other two simulation functions, in the sense that the user enters forest, soil, species-parameter and simulation control inputs directly into the simulation function (in fact, `fordyn()` internally calls `forest2growthInput()` to initialize the input object to function `growth()`).
