### `r name`

```{r, echo = FALSE}
dfbaiSP = dfbai[dfbai$SP==sp,]
dfdbhSP = dfdbh[dfdbh$SP==sp,]
```

Prediction ability for **diameter increase** of surviving trees:

```{r, echo = FALSE, fig.width=8, fig.height = 4, warning=FALSE, message=FALSE}
plot_scatter_dbh(dfdbhSP, points = TRUE)
```

Relationship between **diameter increase** and climatic variables (MAT, P/PET and available PAR):
```{r, echo = FALSE, fig.width=8, fig.height = 12, warning=FALSE, message=FALSE}
plot_cov_clim_dbh(dfdbhSP, points = TRUE)
```

Prediction ability for **stand-level basal area increase** of surviving trees:

```{r, echo = FALSE, fig.width=8, fig.height = 4, warning=FALSE, message=FALSE}
plot_scatter_bai(dfbaiSP)
```

Relationship between **basal area increase** and climatic variables (MAT and P/PET):
```{r, echo = FALSE, fig.width=8, fig.height = 8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfbaiSP)
```

Spatial distribution of errors in **basal area increase**: 
```{r, echo = FALSE, fig.height=7, fig.width=7}
mapdata_sp = mapdata
mapdata_sp$error = NA
mapdata_sp[dfbai$ID[dfbai$Name==name], "error"] = dfbai$error[dfbai$Name==name]
bai_error_map(mapdata_sp)
```
