### `r sp_name`

```{r, echo = FALSE}
sp_names <- sp_name
if(sp_name=="Quercus pubescens") sp_names <- c("Quercus pubescens", "Quercus pyrenaica")
dfbai_SP = dfbai_sp[dfbai_sp$Species_ini %in% sp_names,]
dfbatot_SP = dfbatot_sp[dfbatot_sp$Species %in% sp_names,]
dfbarecr_SP = dfbarecr_sp[dfbarecr_sp$Species %in% sp_names,]
dfbadead_SP = dfbadead_sp[dfbadead_sp$Species %in% sp_names,]
dfNtot_SP = dfNtot_sp[dfNtot_sp$Species %in% sp_names,]
dfNrecr_SP = dfNrecr_sp[dfNrecr_sp$Species %in% sp_names,]
dfNdead_SP = dfNdead_sp[dfNdead_sp$Species %in% sp_names,]
dfdbh_SP = dfdbh[dfdbh$Species_ini %in% sp_names,]
dfh_SP = dfh[dfh$Species_ini %in% sp_names,]
```


#### Annual diameter increment

Prediction ability for **diameter increase** (cm/yr) of surviving trees:
```{r, echo = FALSE}
dfdbh_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```


<!-- Predictive capacity plots: -->

<!-- ```{r, echo = FALSE, fig.width=8, fig.height = 4, warning=FALSE, message=FALSE} -->
<!-- dfdbhSP_bas <- dfdbh_SP |> filter(transpirationMode == "Granier", -->
<!--                              period == "IFN24") -->
<!-- dfdbhSP_adv <- dfdbh_SP |> filter(transpirationMode == "Granier", -->
<!--                              period == "IFN24") -->
<!-- plot_scatter_tree(dfdbhSP_bas, dfdbhSP_adv,  points = TRUE) -->
<!-- ``` -->

<!-- Relationship between **diameter increment** and climatic variables (MAT and P/PET): -->
<!-- ```{r, echo = FALSE, fig.width=8, fig.height = 12, warning=FALSE, message=FALSE} -->
<!-- plot_cov_clim_tree(dfdbhSP_bas, dfdbhSP_adv,  points = TRUE) -->
<!-- ``` -->




#### Annual height increment

Prediction ability for **height increase** (cm/yr) of surviving trees:

```{r, echo = FALSE}
dfh_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```

<!-- Predictive capacity plots (IFN2-IFN4): -->

<!-- ```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE} -->
<!-- dfhSP_bas <- dfh_SP |> filter(transpirationMode == "Granier", -->
<!--                              period == "IFN24") -->
<!-- dfhSP_adv <- dfh_SP |> filter(transpirationMode == "Granier", -->
<!--                              period == "IFN24") -->
<!-- plot_scatter_tree(dfhSP_bas, dfhSP_adv, points = TRUE, xylim = c(0,20), errorlim = c(-20,20), initiallim = c(0, 2000), -->
<!--                   var = "height") -->
<!-- ``` -->

<!-- Relationship between height increase and climatic variables (MAT, P/PET and available PAR; IFN2 - IFN4): -->

<!-- ```{r, echo = FALSE, fig.width=8, fig.height=12, warning=FALSE, message=FALSE} -->
<!-- plot_cov_clim_tree(dfhSP_bas, dfhSP_bas, points = TRUE, xylim = c(0,20), errorlim = c(-20,20), var = "height") -->
<!-- ``` -->

#### Growth basal area increment

Prediction ability for **basal area increase due to growth** (m2/ha/yr) of surviving trees: 

```{r, echo = FALSE}
dfbai_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```

#### Mortality

Prediction ability for **basal area decrease due to mortality** (m2/ha/yr): 

```{r, echo = FALSE}
dfbadead_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```

Prediction ability for **density decrease due to mortality** (ind/ha/yr): 

```{r, echo = FALSE}
dfNdead_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```


#### Ingrowth


Prediction ability for **basal area increase due to ingrowth** (m2/ha/yr): 

```{r, echo = FALSE, warning=FALSE}
dfbarecr_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```


Prediction ability for **density increase due to ingrowth** (ind/ha/yr): 

```{r, echo = FALSE, warning=FALSE}
dfNrecr_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```

#### Overall basal area changes

Prediction ability for **overall basal area changes** (including growth, mortality and ingrowth):

```{r, echo = FALSE, warning=FALSE}
dfbatot_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```


Predictive capacity plots (IFN2-IFN4):

```{r, echo = FALSE, fig.width=8, fig.height = 4, warning=FALSE, message=FALSE}
dfbatot_bas <- dfbatot_SP |> filter(transpirationMode == "Granier",
                             period == "IFN24")
dfbatot_adv <- dfbatot_SP |> filter(transpirationMode == "Sperry",
                             period == "IFN24") 
plot_scatter_bai(dfbatot_bas, dfbatot_adv, quantity = "basal area change", xylim = c(-0.5,2), errorlim = c(-2,2))
```

Relationship between **basal area changes** and climatic variables (MAT and P/PET; IFN2-IFN4):
```{r, echo = FALSE, fig.width=8, fig.height = 8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfbatot_bas, dfbatot_adv, quantity = "Basal area change", ylim = c(-0.5,2), errorlim = c(-2,2))
```

Spatial distribution of errors (IFN2-IFN4): 
```{r, echo = FALSE, fig.height=4, fig.width=9}
p1<-bai_error_map(dfbatot_bas)+labs(title="Basic sub-model")
p2<-bai_error_map(dfbatot_adv)+labs(title="Advanced sub-model")
plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
```


Prediction ability for **overall density changes** (including growth, mortality and ingrowth):

```{r, echo = FALSE, warning=FALSE}
dfNtot_SP |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```