---
title: "Verification of carbon decomposition (DAYCENT)"
author: "Inés Delsman Valderrama (CREAF), Miquel De Cáceres (CREAF)"
date: "`r Sys.Date()`"
description: >
   Verification of carbon decomposition implemented from DAYCENT against the output from CENTURY (4.7)
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(medfate)
library(ggplot2)
library(kableExtra)
library(knitr)
library(tidyr)
library(dplyr)
library(RColorBrewer)
library(cowplot)
```


## Introduction

```{r, include=FALSE}
#' Load data
wd <- '/home/miquel/OneDrive/mcaceres_work/model_development/medfate_evaluation/RcenturyBenchmark/'
file_cent <- paste0(wd,"Rdata/RC2m_Process.RData")
if(!file.exists(file_cent)) stop("File does not exist")
load(file_cent)
```


This document presents an exercise to verify the implementation of the carbon decomposition process of DAYCENT into MEDFATE. The verification is done against the outputs of CENTURY (ver. 4.7) corresponding to simulation of Duke and Harvard forests.

## Methods

```{r, echo = FALSE}
# HARVARD: Extract the first element from tree list
harvard_tree <- tree$`1`
# DUKE: Extract the third element from tree list
duke_tree <- tree$`3`
```

### Initial soil carbon content

+ Surface structural (`strucc(1)`) was applied as the initial carbon content for the `Leaves` pool. Similarly, belowground structural (`strucc(2)`) was applied to the `FineRoots` pool. The remaining litter pools (snags, small branches, large wood and coarse roots) are set to zero initial value.

```{r, include = FALSE}
# Initialize litter pools
nlitter <- 1
Leaves_D <- out_duke$strucc.1.[1]
FineRoots_D <- out_duke$strucc.2.[1]
litterData_DUKE <- data.frame(Species = as.character(rep(duke_tree$label, nlitter)),
                         Leaves = as.numeric(rep(Leaves_D, nlitter)),
                         Twigs = as.numeric(rep(0, nlitter)),
                         SmallBranches = as.numeric(rep(0, nlitter)),
                         LargeWood = as.numeric(rep(0, nlitter)),
                         CoarseRoots = as.numeric(rep(0, nlitter)),
                         FineRoots = as.numeric(rep(FineRoots_D, nlitter)))
Leaves_H <- out_harvard$strucc.1.[1]
FineRoots_H <- out_harvard$strucc.2.[1]
litterData_HARVARD <- data.frame(Species = as.character(rep(harvard_tree$label, nlitter)),
                         Leaves = as.numeric(rep(Leaves_H, nlitter)),
                         Twigs = as.numeric(rep(0, nlitter)),
                         SmallBranches = as.numeric(rep(0, nlitter)),
                         LargeWood = as.numeric(rep(0, nlitter)),
                         CoarseRoots = as.numeric(rep(0, nlitter)),
                         FineRoots = as.numeric(rep(FineRoots_H, nlitter)))
# Initialize (empty) snag pools
nsnag <- 0
snagData<- data.frame(Species = as.character(rep(NA, nsnag)),
                      DBH = as.numeric(rep(NA, nsnag)),
                      Height = as.numeric(rep(NA, nsnag)),
                      DeathAge = as.numeric(rep(NA, nsnag)),
                      SmallBranches = as.numeric(rep(NA, nsnag)),
                      LargeWood = as.numeric(rep(NA, nsnag)))
```

+ We took the soil organic carbon (SOC) initial values from the CENTURY `site.100` input file, and the surface and belowground metabolic data from the first row of the output file of our CENTURY simulation of the Duke site.

```{r, include = FALSE}
# Duke
SOM1C1_D <- subset(site_d$`organic matter`$df, field == "SOM1CI(1,1)")$value
SOM1C2_D <- subset(site_d$`organic matter`$df, field == "SOM1CI(2,1)")$value
SOM2C1_D <- subset(site_d$`organic matter`$df, field == "SOM2CI(1,1)")$value
SOM2C2_D <- subset(site_d$`organic matter`$df, field == "SOM2CI(2,1)")$value
SOM3C_D <- subset(site_d$`organic matter`$df, field == "SOM3CI(1)")$value

Metabc1_D <-   out_duke$metabc.1.[1]
Metabc2_D <- out_duke$metabc.2.[1]

SOCData_DUKE <- c(SurfaceMetabolic = Metabc1_D, BelowgroundMetabolic = Metabc2_D, SurfaceActive = SOM1C1_D, BelowgroundActive = SOM1C2_D, SurfaceSlow = SOM2C1_D, BelowgroundSlow = SOM2C2_D, BelowgroundPassive = SOM3C_D)

# Harvard
SOM1C1_H <- subset(site_h$`organic matter`$df, field == "SOM1CI(1,1)")$value
SOM1C2_H <- subset(site_h$`organic matter`$df, field == "SOM1CI(2,1)")$value
SOM2C1_H <- subset(site_h$`organic matter`$df, field == "SOM2CI(1,1)")$value
SOM2C2_H <- subset(site_h$`organic matter`$df, field == "SOM2CI(2,1)")$value
SOM3C_H <- subset(site_h$`organic matter`$df, field == "SOM3CI(1)")$value

Metabc1_H <- out_harvard$metabc.1.[1]
Metabc2_H <- out_harvard$metabc.2.[1]

SOCData_HARVARD <- c(SurfaceMetabolic = Metabc1_H, BelowgroundMetabolic = Metabc2_H, SurfaceActive = SOM1C1_H, BelowgroundActive = SOM1C2_H, SurfaceSlow = SOM2C1_H, BelowgroundSlow = SOM2C2_H, BelowgroundPassive = SOM3C_H)
```


|  MEDFATE pool             | CENTURY parameter | DUKE         | HARVARD      |
|---------------------------|-------------------|--------------|--------------|
| `Leaves`                  |  `strucc.1.`      | `r round(Leaves_D)` | `r round(Leaves_H)` |
| `FineRoots`               |  `strucc.2.`      | `r round(FineRoots_D)` | `r round(FineRoots_H)` |
| `SurfaceMetabolic`        |  `metabc.1.`      | `r round(Metabc1_D)` | `r round(Metabc1_H)` |
| `BelowgroundMetabolic`    |  `metabc.2.`      | `r round(Metabc2_D)` | `r round(Metabc2_H)` |
| `SurfaceActive`           |  `SOM1CI(1,1)`    | `r round(SOM1C1_D)`    | `r round(SOM1C1_H)`    |
| `BelowgroundActive`       |  `SOM1CI(2,1)`    | `r round(SOM1C2_D)`    | `r round(SOM1C2_H)`    |
| `SurfaceSlow`             |  `SOM2CI(1,1)`    | `r round(SOM2C1_D)`    | `r round(SOM2C1_H)`    |
| `BelowgroundSlow`         |  `SOM2CI(2,1)`    | `r round(SOM2C2_D)`    | `r round(SOM2C2_H)`    |
| `BelowgroundPassive`         |  `SOM3CI(1)`    | `r round(SOM3C_D)`    | `r round(SOM3C_H)`    |


### Decomposition parameters

#### Litter decomposition parameters

```{r, echo=FALSE}
#' Function to extract decomposition parameters from 'tree' objects in Rcentury
sp_decomp_params <- function(rcentury_tree) {
   # % lignin leaves tree$`3`$df$WDLIG(1)
   # % lignin fine roots tree$`3`$df$WDLIG(2)
   # % lignin branches tree$`3`$df$WDLIG(3)
   # % lignin large wood tree$`3`$df$WDLIG(4)
   # % lignin coarse roots tree$`3`$df$WDLIG(5)
   # average 3-5 for sapwood lignin
   # 
   # read lignin/N ratios from
   # tree$cerfor(3,1,1) C/N ratio
     
  wdlig1 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(1)",1]
  wdlig2 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(2)",1]
  wdlig3 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(3)",1]
  wdlig4 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(4)",1]
  wdlig5 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(5)",1]
  cn_leaves <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,1,1)",1]
  cn_fineroots <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,2,1)",1]
  cn_branches <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,3,1)",1]
  cn_largewood <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,4,1)",1]
  cn_coarseroots <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,5,1)",1]
  cn_wood <- mean(c(cn_branches, cn_largewood, cn_coarseroots))
  leaf_c <- 0.5
  wood_c <- 0.5
  fineroot_c <- 0.5
  n_leaves <- (1/cn_leaves)*leaf_c
  n_fineroot <- (1/cn_fineroots)*fineroot_c
  n_wood <- (1/cn_wood)*wood_c
  df <- data.frame(Species  = rcentury_tree$label,
                   LeafLignin = wdlig1*100, # %
                   WoodLignin = mean(wdlig3, wdlig4, wdlig5)*100,# %
                   FineRootLignin = wdlig2*100,# %
                   Nleaf =  n_leaves*1000, # mg / g
                   Nsapwood = n_wood*1000, # mg / g
                   Nfineroot = n_fineroot*1000,
                   LeafLigninN = wdlig1/n_leaves) 
  return(df)
}

# Extracts litter decomposition parameters
paramsLitterDecomposition_DUKE <- sp_decomp_params(duke_tree)
paramsLitterDecomposition_HARVARD <- sp_decomp_params(harvard_tree)
# paramsLitterDecomposition_HARVARD$LeafLignin <- 5

# Anatomy parameters
paramsAnatomy_DUKE <- data.frame(Species = "DUKE",
                            WoodDensity = 0.7)
paramsAnatomy_HARVARD <- data.frame(Species = "HARVF",
                            WoodDensity = 0.7)

knitr::kable(rbind(paramsLitterDecomposition_DUKE, paramsLitterDecomposition_HARVARD))
```

#### Maximum decomposition rates

```{r, echo = FALSE}
SurfaceMetabolic <- fix$`1`$df[fix$`1`$df[,2]=="DEC2(1)",1]
BelowgroundMetabolic <- fix$`1`$df[fix$`1`$df[,2]=="DEC2(2)",1]
Leaves <- fix$`1`$df[fix$`1`$df[,2]=="DEC1(1)",1]
FineRoots <- fix$`1`$df[fix$`1`$df[,2]=="DEC1(2)",1]
SurfaceActive <- fix$`1`$df[fix$`1`$df[,2]=="DEC3(1)",1]
BelowgroundActive <- fix$`1`$df[fix$`1`$df[,2]=="DEC3(2)",1]
SurfaceSlow <- fix$`1`$df[fix$`1`$df[,2]=="DEC5(1)",1]
BelowgroundSlow <- fix$`1`$df[fix$`1`$df[,2]=="DEC5(2)",1]
BelowgroundPassive <- fix$`1`$df[fix$`1`$df[,2]=="DEC4",1]

SmallBranches_D <- duke_tree$df[duke_tree$df[,2]=="DECW1",1]
LargeWood_D <- duke_tree$df[duke_tree$df[,2]=="DECW2",1]
CoarseRoots_D <- duke_tree$df[duke_tree$df[,2]=="DECW3",1]

SmallBranches_H <- harvard_tree$df[harvard_tree$df[,2]=="DECW1",1]
LargeWood_H <- harvard_tree$df[harvard_tree$df[,2]=="DECW2",1]
CoarseRoots_H <- harvard_tree$df[harvard_tree$df[,2]=="DECW3",1]

annualTurnoverRate_DUKE <- duke_tree$df[duke_tree$df[,2]=="TMIX",1]
annualTurnoverRate_HARVARD <- harvard_tree$df[harvard_tree$df[,2]=="TMIX",1]
annualTurnoverRate_DUKE <- 0.11
annualTurnoverRate_HARVARD <- 0.11
```

The maximum (base) decomposition rates for the different soil pools in MEDFATE (i.e. control parameter `baseAnnualRates`) were adapted to match the maximum rates from CENTURY (these can be found in the `fix.100` file). 

|  Element in `baseAnnualRates`  | CENTURY pool | CENTURY parameter | Value |
|--------------------------------|--------------|-------------------|-------|
| `Leaves`             | Surface structural | `DEC1(1)`         | `r Leaves` |
| `FineRoots`             | Belowground structural | `DEC1(2)`       | `r FineRoots`|
| `SurfaceMetabolic`   | Surface metabolic  | `DEC2(1)`         | `r SurfaceMetabolic` |
| `BelowgroundMetabolic`   | Belowground metabolic  | `DEC2(2)`         | `r BelowgroundMetabolic` |
| `SurfaceActive` | Surface active | `DEC3(1)` | `r SurfaceActive` |
| `BelowgroundActive` | Soil active | `DEC3(2)` | `r BelowgroundActive` |
| `BelowgroundPassive` | Soil slow turnover | `DEC4` | `r BelowgroundPassive` |
| `SurfaceSlow` | Surface intermediate | `DEC5(1)` | `r SurfaceSlow` |
| `BelowgroundSlow` | Soil intermediate | `DEC5(2)` | `r BelowgroundSlow` |

Maximum decomposition rates for dead wood pools (small branches, large wood and coarse roots) were taken from elements `DECW1`, `DECW2` and `DECW3` of `tree.fix`. 

|  Element in `baseAnnualRates`  | CENTURY parameter | DUKE    | Harvard |
|--------------------------------|-------------------|---------|---------|
| `SmallBranches`                |  `DECW1`          | `r SmallBranches_D` | `r SmallBranches_H` |
| `LargeWood`                |  `DECW2`          | `r LargeWood_D` | `r LargeWood_H` |
| `CoarseRoots`                |  `DECW3`          | `r CoarseRoots_D` | `r CoarseRoots_H` |


Rate of mixing between surface slow and belowground slow compartments was taken from `tree.100`:

|  Element in `control`  | CENTURY parameter | DUKE    | Harvard |
|--------------------------------|-------------------|---------|---------|
| `annualTurnoverRate`                |  `TMIX`          | `r annualTurnoverRate_DUKE` | `r annualTurnoverRate_HARVARD` |

```{r, echo = FALSE}
#Change base rates
control <- defaultControl()
baseAnnualRates <- control$decompositionAnnualBaseRates
baseAnnualRates_DUKE <- c(SurfaceMetabolic = SurfaceMetabolic, 
                          BelowgroundMetabolic = BelowgroundMetabolic, 
                          Leaves = Leaves , 
                          FineRoots = FineRoots, 
                          Twigs = 1.8,  
                          SmallBranches = SmallBranches_D,
                          LargeWood = LargeWood_D, 
                          CoarseRoots = CoarseRoots_D,
                          SurfaceActive = SurfaceActive, 
                          BelowgroundActive = BelowgroundActive, 
                          SurfaceSlow = SurfaceSlow, 
                          BelowgroundSlow = BelowgroundSlow, 
                          BelowgroundPassive = BelowgroundPassive)
baseAnnualRates_HARVARD <- c(SurfaceMetabolic = SurfaceMetabolic, 
                          BelowgroundMetabolic = BelowgroundMetabolic, 
                          Leaves = Leaves,  ## 
                          FineRoots = FineRoots, 
                          Twigs = 1.8,  
                          SmallBranches = SmallBranches_H,
                          LargeWood = LargeWood_H, 
                          CoarseRoots = CoarseRoots_H,
                          SurfaceActive = SurfaceActive, 
                          BelowgroundActive = BelowgroundActive, 
                          SurfaceSlow = SurfaceSlow, 
                          BelowgroundSlow = BelowgroundSlow, 
                          BelowgroundPassive = BelowgroundPassive)
```


#### Soil texture and pH

We took the the sand and clay values (and multiplied them by 100 to obtain the fractions) and the pH value from the CENTURY `site.100` file.

```{r, echo = FALSE}
# Retrieve soil site data
sand_DUKE <- subset(site_d$`site and control`$df, field == "SAND")$value * 100
clay_DUKE <- subset(site_d$`site and control`$df, field == "CLAY")$value * 100
soilPH_DUKE <- subset(site_d$`site and control`$df, field == "PH")$value 

sand_HARVARD <- subset(site_h$`site and control`$df, field == "SAND")$value * 100
clay_HARVARD <- subset(site_h$`site and control`$df, field == "CLAY")$value * 100
soilPH_HARVARD <- subset(site_h$`site and control`$df, field == "PH")$value 
```


|  Soil property  | CENTURY parameter | DUKE    | Harvard |
|--------------------------------|-------------------|---------|---------|
| Percent sand                |  `SAND`          | `r sand_DUKE` | `r sand_HARVARD` |
| Percent clay                |  `CLAY`          | `r clay_DUKE` | `r clay_HARVARD` |
| Soil pH                |  `PH`          | `r soilPH_DUKE` | `r soilPH_HARVARD` |

### Environmental variation

Variation of environmental conditions for carbon decomposition were drawn from CENTURY outputs. Specifically, soil temperature was taken from CENTURY variable `stemp`, whereas CENTURY variable `asmos.1.` was used to estimate soil relative moisture.

```{r, echo=FALSE,fig.width=8, fig.height=6}
#' Function to transform Rcentury output for soil temperature and soil moisture from month to day resolution
env_daily<-function(out, yearIni = 2000) {
  yearMax <- yearIni + max(out$time) - 1
  dateIni <- as.Date(paste0(yearIni, "-01-01"))
  dateFin <- as.Date(paste0(yearMax, "-12-31"))
  days <- seq(dateIni, dateFin, by="day")
  months <- cut(days, breaks="months")
  unique_months <- unique(months)
  # Trim out to the complete years
  out <- out[1:length(unique_months), , drop = FALSE]
  df <- data.frame(date = days, Step = 1:length(days), SoilMoisture = NA, SoilTemperature = NA,
                   AirTemperature = NA, AirRelativeHumidity = NA)
  rel_mois <- out$asmos.1./ max(out$asmos.1.)
  for(i in 1:length(unique_months)) {
    sel_i <- months == unique_months[i]
    df$SoilMoisture[sel_i] <- rel_mois[i]
    df$SoilTemperature[sel_i] <- out$stemp[i]
  }
  return(df)
}

# Extracts environmental conditions used in CENTURY simulations
environmentalConditions_DUKE <- env_daily(out_duke)
environmentalConditions_DUKE$forest <- "DUKE"
environmentalConditions_HARVARD <- env_daily(out_harvard)
environmentalConditions_HARVARD$forest <- "HARVARD"

g1 <- ggplot(rbind(environmentalConditions_DUKE, environmentalConditions_HARVARD), 
       aes(x = date, y = SoilTemperature, col = forest)) +
  geom_line(linewidth = 0.5) +
  labs(title = "Soil temperature",
       x = "Year", 
       y = "Soil temperature (ºC)") +
  theme_minimal() 
g2 <- ggplot(rbind(environmentalConditions_DUKE, environmentalConditions_HARVARD), 
       aes(x = date, y = SoilMoisture, col = forest)) +
  geom_line(linewidth = 0.5) +
  labs(title = "Soil moisture",
       x = "Year", 
       y = "Relative soil moisture") +
  theme_minimal() 
plot_grid(g1, g2, nrow = 2)
```

### Litter input

Litter inputs (senescence of leaves, small branches, large wood, fine roots and coarse roots) are estimated using CENTURY output corresponding to variation in live carbon pools in the forest system (i.e. `rleavc`, `fbrchc`, `rlwodc`, `frootc` and `crootc`) and the CENTURY parameters specifying monthly death rates for those live pools (i.e. leaf death rates `leafdr(x)` for each month and `wooddr(2-5)` for the remaining four pools)

```{r, include=FALSE, fig.width=8, fig.height=8}
prd_daily_new<-function(out, rcentury_tree, yearIni = 2000) {
  decid <- rcentury_tree$df[rcentury_tree$df[,2]=="DECID",1]
  species <- rcentury_tree$label
  yearMax <- yearIni + max(out$time) - 1
  dateIni <- as.Date(paste0(yearIni, "-01-01"))
  dateFin <- as.Date(paste0(yearMax, "-12-31"))
  days <- seq(dateIni, dateFin, by="day")
  months <- cut(days, breaks="months")
  unique_months <- unique(months)
  leafdr <- c(rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(1)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(2)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(3)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(4)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(5)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(6)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(7)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(8)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(9)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(10)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(11)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(12)",1])
  wooddr1 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(1)",1] #fraction of forest which is deciduous; the fraction of leaves which fall during senescence month or at the end of the growing season, used when decid = 1 or 2
  wooddr2 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(2)",1] #monthly death rate fraction for fine root component
  wooddr3 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(3)",1] #monthly death rate fraction for fine branch component
  wooddr4 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(4)",1] #monthly death rate fraction for large wood component
  wooddr5 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(5)",1] #monthly death rate fraction for coarse root component
  # Trim out to the complete years
  out <- out_duke[1:length(unique_months), , drop = FALSE]
  # fbrchc -> C in forest system fine branch component (g m-2).
  # crootc -> C in forest system coarse root component (g m-2).
  # rlwodc -> C in forest system large wood component (g m-2).
  # frootc -> C in forest system fine root component (g m-2).
  # rleavc -> C in forest system leaf component (g m-2).
  # print(wooddr1)
  df <- data.frame(date = days, Step = 1:length(days), Species = species, Leaves = NA, Twigs = NA, SmallBranches = NA, LargeWood = NA, CoarseRoots = NA, FineRoots = NA)
  for(i in 1:length(unique_months)) {
    sel_i <- months == unique_months[i]
    nmonth_days <- sum(sel_i)
    month_number <- as.numeric(format(days[sel_i], "%m")[1])
    if(decid == 0) {
      df$Leaves[sel_i] <- out$rleavc[i]*leafdr[month_number]/nmonth_days
    } else {
      if((i %% 12)==10) df$Leaves[sel_i] <- out$rleavc[i]*wooddr1/nmonth_days
      else df$Leaves[sel_i] <- 0
    }
    df$Twigs[sel_i] <- 0
    df$SmallBranches[sel_i] <- out$fbrchc[i]*wooddr3/nmonth_days
    df$LargeWood[sel_i] <- out$rlwodc[i]*wooddr4/nmonth_days
    df$CoarseRoots[sel_i] <- out$crootc[i]*wooddr5/nmonth_days
    df$FineRoots[sel_i] <- out$frootc[i]*wooddr2/nmonth_days
  }
  return(df)
}

# Extracts litter production during CENTURY simulations
litterProduction_DUKE <- prd_daily_new(out_duke, duke_tree)
litterProduction_HARVARD <- prd_daily_new(out_harvard, harvard_tree)
```

```{r, echo = FALSE}
RC2m_Litter_DUKE <- as.data.frame(litterProduction_DUKE) |>
  select(date, Leaves, SmallBranches, LargeWood, CoarseRoots, FineRoots)|>
  pivot_longer(cols = c(Leaves, SmallBranches, LargeWood, CoarseRoots, FineRoots), 
               names_to = "pool", 
               values_to = "value") 
RC2m_Litter_HARVARD <- as.data.frame(litterProduction_HARVARD) |>
  select(date, Leaves, SmallBranches, LargeWood, CoarseRoots, FineRoots)|>
  pivot_longer(cols = c(Leaves, SmallBranches, LargeWood, CoarseRoots, FineRoots), 
               names_to = "pool", 
               values_to = "value") 
g1 <- ggplot(RC2m_Litter_DUKE, 
       aes(x = date, y = value, color = pool)) +
  geom_line(linewidth = 0.5) +
  labs(title = "Litter input (DUKE)",
       x = "Year", 
       y = "Daily litter input (g C/m2/day)",
       color = "Pool") +
  theme_minimal() +
  theme(legend.position = "right")
g2 <- ggplot(RC2m_Litter_HARVARD, 
       aes(x = date, y = value, color = pool)) +
  geom_line(linewidth = 0.5) +
  labs(title = "Litter input (HARVARD)",
       x = "Year", 
       y = "Daily litter input (g C/m2/day)",
       color = "Pool") +
  theme_minimal() +
  theme(legend.position = "right")
plot_grid(g1, g2, nrow = 2)
```


### Simulations

MEDFATE simulations were run using function `decomposition_DAYCENT()`. 

```{r, echo=FALSE}
# Call DAYCENT
# IMPORTANT: Species should be the same in 'litterData', 'paramsLitterDecomposition' and 'litterProduction'
res_DUKE <-decomposition_DAYCENT(snagData, litterData_DUKE, SOCData_DUKE,
                            paramsLitterDecomposition_DUKE,
                            paramsAnatomy_DUKE,
                            baseAnnualRates_DUKE,
                            annualTurnoverRate_DUKE,
                            environmentalConditions_DUKE,
                            litterProduction_DUKE,
                            sand_DUKE,
                            clay_DUKE,
                            soilPH_DUKE)

res_HARVARD <-decomposition_DAYCENT(snagData, litterData_HARVARD, SOCData_HARVARD,
                            paramsLitterDecomposition_HARVARD,
                            paramsAnatomy_HARVARD,
                            baseAnnualRates_HARVARD,
                            annualTurnoverRate_HARVARD,
                            environmentalConditions_HARVARD,
                            litterProduction_HARVARD,
                            sand_HARVARD,
                            clay_HARVARD,
                            soilPH_HARVARD)
```

### Comparison

The monthly results from the CENTURY simulation are transformed to daily data to make it comparable to the results from MEDFATE. The results are compared in terms of total soil C stock, soil organic matter pools (SOM pools), (leaf and fine root) litter pools and dead wood pools. 

## Results

### Total soil carbon


```{r, echo=FALSE}
# Get dimensions
n_days <- nrow(res_DUKE$DecompositionPools)  
n_months <- nrow(out_duke) 

# Calculate days per month (approximately equal)
days_per_month <- rep(round(n_days / n_months), n_months)
# Adjust last month to match exactly
days_per_month[length(days_per_month)] <- n_days - sum(days_per_month[-length(days_per_month)])

Rcentury_DUKE <- cbind(
  out_duke[, c("strucc.1.", "strucc.2.", "metabc.1.", "metabc.2.", "som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c")]) |>
  mutate(Total = rowSums(across(c(strucc.1., strucc.2., metabc.1., metabc.2., som1c.1., som1c.2., som2c.1., som2c.2., som3c)), na.rm = TRUE)) |>
  select(Total) |>
  tidyr::pivot_longer(cols = c(Total), 
                      names_to = "pool", 
                      values_to = "value") |>
  mutate(
    days = days_per_month,
    monthly_value = value  
  ) |>
  rowwise() |>
  mutate(
    daily_values = list(rep(monthly_value, days))
  ) |>
  unnest(daily_values) |>
  mutate(
    day = 1:n(),
    model = "CENTURY",
    forest = "DUKE"
  ) |>
  select(day, value = daily_values, model)

Rcentury_HARVARD <- cbind(
  out_harvard[, c("strucc.1.", "strucc.2.", "metabc.1.", "metabc.2.", "som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c")]) |>
  mutate(Total = rowSums(across(c(strucc.1., strucc.2., metabc.1., metabc.2., som1c.1., som1c.2., som2c.1., som2c.2., som3c)), na.rm = TRUE)) |>
  select(Total) |>
  tidyr::pivot_longer(cols = c(Total), 
                      names_to = "pool", 
                      values_to = "value") |>
  mutate(
    days = days_per_month,
    monthly_value = value  
  ) |>
  rowwise() |>
  mutate(
    daily_values = list(rep(monthly_value, days))
  ) |>
  unnest(daily_values) |>
  mutate(
    day = 1:n(),
    model = "CENTURY",
    forest = "HARVARD"
  ) |>
  select(day, value = daily_values, model)
```



```{r message=FALSE, warning=FALSE, fig.width=8, fig.height=6, echo=FALSE}
# Create data frames with time indices
RC2m_HARVARD <- data.frame(
  value = rowSums(res_HARVARD$DecompositionPools[,c(3,8:15)], na.rm = TRUE),
  day = 1:n_days,
  model = "MEDFATE"
)
# Create data frames with time indices
RC2m_DUKE <- data.frame(
  value = rowSums(res_DUKE$DecompositionPools[,c(3,8:15)], na.rm = TRUE),
  day = 1:n_days,
  model = "MEDFATE"
)
# Plot

g1 <- ggplot(bind_rows(RC2m_DUKE, Rcentury_DUKE), aes(x = day, y = value, color = model)) +
  geom_line() +
  labs(title = "Total soil C stock (DUKE)",
       x = "Time (20 years in days)", 
       y = "Soil C stocks (g/m2)") +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = c(0.8,0.2))
  
# Plot
g2 <- ggplot(bind_rows(RC2m_HARVARD, Rcentury_HARVARD), aes(x = day, y = value, color = model)) +
  geom_line() +
  labs(title = "Total soil C stock (HARVARD)",
       x = "Time (20 years in days)", 
       y = "Soil C stocks (g/m2)") +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = c(0.8,0.8))

plot_grid(g1, g2, nrow = 2)
```

### SOM pools

```{r, echo=FALSE, fig.width=8, fig.height=8}
SOMPools <- c("SurfaceActive", "BelowgroundActive", "SurfaceSlow", "BelowgroundSlow", "BelowgroundPassive", "Total")
SOM_col <- brewer.pal(n = length(SOMPools), name = "Dark2")

reps <- round(n_days / n_months)

## DUKE
Rcentury_SOMpools_DUKE <- cbind(
  out_duke[, c("som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c")]
) |>
  mutate(Total = rowSums(across(c(som1c.1., som1c.2., som2c.1., som2c.2., som3c)), na.rm = TRUE))  |>
  tidyr::pivot_longer(cols = c(som1c.1., som1c.2., som2c.1., som2c.2., som3c,Total), 
                      names_to = "pool", 
                      values_to = "value")|>
  mutate(
    pool = case_when(
      pool == "som1c.1." ~ "SurfaceActive",
      pool == "som1c.2." ~ "BelowgroundActive",
      pool == "som2c.1." ~ "SurfaceSlow",
      pool == "som2c.2." ~ "BelowgroundSlow",
      pool == "som3c" ~ "BelowgroundPassive",
      pool == "Total" ~ "Total")) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  mutate(
    day = 1:n(),
    model = "CENTURY"
  ) |>
  select(day, value, pool, model)

RC2m_SOMpools_DUKE <- as.data.frame(res_DUKE$DecompositionPools[, 11:15]) |>
  mutate(Total = SurfaceActive+ BelowgroundActive+ SurfaceSlow+BelowgroundSlow +BelowgroundPassive,
         day = 1:n_days) |>
  pivot_longer(cols = all_of(SOMPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(model = "MEDFATE")

## HARVARD
Rcentury_SOMpools_HARVARD <- cbind(
  out_harvard[, c("som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c")]
) |>
  mutate(Total = rowSums(across(c(som1c.1., som1c.2., som2c.1., som2c.2., som3c)), na.rm = TRUE))  |>
  tidyr::pivot_longer(cols = c(som1c.1., som1c.2., som2c.1., som2c.2., som3c,Total), 
                      names_to = "pool", 
                      values_to = "value")|>
  mutate(
    pool = case_when(
      pool == "som1c.1." ~ "SurfaceActive",
      pool == "som1c.2." ~ "BelowgroundActive",
      pool == "som2c.1." ~ "SurfaceSlow",
      pool == "som2c.2." ~ "BelowgroundSlow",
      pool == "som3c" ~ "BelowgroundPassive",
      pool == "Total" ~ "Total")) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  mutate(
    day = 1:n(),
    model = "CENTURY"
  ) |>
  select(day, value, pool, model)

RC2m_SOMpools_HARVARD <- as.data.frame(res_HARVARD$DecompositionPools[, 11:15]) |>
  mutate(Total = SurfaceActive+ BelowgroundActive+ SurfaceSlow+BelowgroundSlow +BelowgroundPassive,
         day = 1:n_days) |>
  pivot_longer(cols = all_of(SOMPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(model = "MEDFATE")

# PLOTTING

g1 <- ggplot(bind_rows(RC2m_SOMpools_DUKE, Rcentury_SOMpools_DUKE)|>
         filter(pool != "Total"), 
       aes(x = day, y = value, color = pool, linetype = model)) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("MEDFATE" = 1, "CENTURY" = 2)) + 
  labs(title = "SOM Pools (DUKE)",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")

g2 <- ggplot(bind_rows(RC2m_SOMpools_HARVARD, Rcentury_SOMpools_HARVARD)|>
         filter(pool != "Total"), 
       aes(x = day, y = value, color = pool, linetype = model)) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("MEDFATE" = 1, "CENTURY" = 2)) + 
  labs(title = "SOM Pools (HARVARD)",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")
plot_grid(g1, g2, nrow = 2)
```

### Structural (leaf and fine root) and metabolic pools

```{r echo=FALSE, fig.width=8, fig.height=8}
LitterPools <- c("SurfaceLitter", "BelowgroundLitter", "SurfaceMetabolic", "BelowgroundMetabolic", "Total")
Litter_col <- brewer.pal(n = length(LitterPools), name = "Dark2")

Rcentury_LitterPools_DUKE <- cbind(
  out_duke[, c("strucc.1.","strucc.2.", "metabc.1.", "metabc.2.")]
) |>
  mutate(Total = rowSums(dplyr::across(c(strucc.1., strucc.2., metabc.1., metabc.2.)), na.rm = TRUE)) |> 
  dplyr::select(
    V1  = "strucc.1.",           # Direct renaming
    V2 = "strucc.2.",           # Direct renaming  
    V3 = "metabc.1.",
    V4 = "metabc.2.",
    V5 = "Total",
  ) |>
  tidyr::pivot_longer(cols = c(V1,V2,V3,V4,V5), 
                      names_to = "pool", 
                      values_to = "value") |>
  mutate(
    pool = case_when(
      pool == "V1" ~ "SurfaceLitter",
      pool == "V2" ~ "BelowgroundLitter",
      pool == "V3" ~ "SurfaceMetabolic",
      pool == "V4" ~ "BelowgroundMetabolic",
      pool == "V5" ~ "Total")) |>
  group_by(pool) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  ungroup() |>
  mutate(
    day = rep(1:n_days, times = n_distinct(pool)),  # Creates 1:n_days for each pool
    type = "CENTURY"
  ) |>
  select(day, value, pool, type)

RC2m_LitterPools_DUKE <- as.data.frame(res_DUKE$DecompositionPools) |>
  select(LeafLitter, FineRootLitter, SurfaceMetabolic, BelowgroundMetabolic)|>
  rename(SurfaceLitter = LeafLitter,
         BelowgroundLitter = FineRootLitter) |>
  mutate(Total = SurfaceLitter + BelowgroundLitter + SurfaceMetabolic + BelowgroundMetabolic,
         day = 1:n_days) |>
  pivot_longer(cols = all_of(LitterPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "MEDFATE")
## HARVARD
Rcentury_LitterPools_HARVARD <- cbind(
  out_harvard[, c("strucc.1.","strucc.2.", "metabc.1.", "metabc.2.")]
) |>
  mutate(Total = rowSums(dplyr::across(c(strucc.1., strucc.2., metabc.1., metabc.2.)), na.rm = TRUE)) |> 
  dplyr::select(
    V1  = "strucc.1.",           # Direct renaming
    V2 = "strucc.2.",           # Direct renaming  
    V3 = "metabc.1.",
    V4 = "metabc.2.",
    V5 = "Total",
  ) |>
  tidyr::pivot_longer(cols = c(V1,V2,V3,V4,V5), 
                      names_to = "pool", 
                      values_to = "value") |>
  mutate(
    pool = case_when(
      pool == "V1" ~ "SurfaceLitter",
      pool == "V2" ~ "BelowgroundLitter",
      pool == "V3" ~ "SurfaceMetabolic",
      pool == "V4" ~ "BelowgroundMetabolic",
      pool == "V5" ~ "Total")) |>
  group_by(pool) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  ungroup() |>
  mutate(
    day = rep(1:n_days, times = n_distinct(pool)),  # Creates 1:n_days for each pool
    type = "CENTURY"
  ) |>
  select(day, value, pool, type)

RC2m_LitterPools_HARVARD <- as.data.frame(res_HARVARD$DecompositionPools) |>
  select(LeafLitter, FineRootLitter, SurfaceMetabolic, BelowgroundMetabolic)|>
  rename(SurfaceLitter = LeafLitter,
         BelowgroundLitter = FineRootLitter) |>
  mutate(Total = SurfaceLitter + BelowgroundLitter + SurfaceMetabolic + BelowgroundMetabolic,
         day = 1:n_days) |>
  pivot_longer(cols = all_of(LitterPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "MEDFATE")


## PLOTTING
g1 <- ggplot(bind_rows(RC2m_LitterPools_DUKE, Rcentury_LitterPools_DUKE) |>
         filter(pool != "Total"), 
       aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("MEDFATE" = 1, "CENTURY" = 2)) + 
  labs(title = "Litter Pools (DUKE)",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  # ylim(0, 500) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")

g2 <- ggplot(bind_rows(RC2m_LitterPools_HARVARD, Rcentury_LitterPools_HARVARD)|>
         filter(pool != "Total"), 
       aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("MEDFATE" = 1, "CENTURY" = 2)) + 
  labs(title = "Litter Pools (HARVARD)",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  # ylim(0, 500) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")
plot_grid(g1, g2, nrow = 2)
```


### Dead wood pools

```{r echo=FALSE, fig.width=8, fig.height=8}
DeadWoodPools <- c("SmallBranches", "LargeWood", "CoarseRoots", "Total")
DeadWood_col <- brewer.pal(n = length(DeadWoodPools), name = "Dark2")

## DUKE
Rcentury_DeadWoodPools_DUKE <- cbind(
  out_duke[, c("wood1c","wood2c", "wood3c")]
) |>
  mutate(Total = rowSums(dplyr::across(c(wood1c, wood2c, wood3c)), na.rm = TRUE)) |> 
  dplyr::select(
    V1  = "wood1c",           # Direct renaming
    V2 = "wood2c",           # Direct renaming  
    V3 = "wood3c",
    V4 = "Total",
  ) |>
  tidyr::pivot_longer(cols = c(V1,V2,V3,V4), 
                      names_to = "pool", 
                      values_to = "value") |>
  mutate(
    pool = case_when(
      pool == "V1" ~ "SmallBranches",
      pool == "V2" ~ "LargeWood",
      pool == "V3" ~ "CoarseRoots",
      pool == "V4" ~ "Total")) |>
  group_by(pool) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  ungroup() |>
  mutate(
    day = rep(1:n_days, times = n_distinct(pool)),  # Creates 1:n_days for each pool
    type = "CENTURY"
  ) |>
  select(day, value, pool, type)

RC2m_DeadWoodPools_DUKE <- as.data.frame(res_DUKE$DecompositionPools) |>
  select(SmallBranchLitter, LargeWoodLitter, CoarseRootLitter)|>
  rename(SmallBranches = SmallBranchLitter,
         LargeWood = LargeWoodLitter,
         CoarseRoots =  CoarseRootLitter) |>
  mutate(Total = SmallBranches + LargeWood + CoarseRoots,
         day = 1:n_days) |>
  pivot_longer(cols = all_of(DeadWoodPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "MEDFATE")

## HARVARD
Rcentury_DeadWoodPools_HARVARD <- cbind(
  out_harvard[, c("wood1c","wood2c", "wood3c")]
) |>
  mutate(Total = rowSums(dplyr::across(c(wood1c, wood2c, wood3c)), na.rm = TRUE)) |> 
  dplyr::select(
    V1  = "wood1c",           # Direct renaming
    V2 = "wood2c",           # Direct renaming  
    V3 = "wood3c",
    V4 = "Total",
  ) |>
  tidyr::pivot_longer(cols = c(V1,V2,V3,V4), 
                      names_to = "pool", 
                      values_to = "value") |>
  mutate(
    pool = case_when(
      pool == "V1" ~ "SmallBranches",
      pool == "V2" ~ "LargeWood",
      pool == "V3" ~ "CoarseRoots",
      pool == "V4" ~ "Total")) |>
  group_by(pool) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  ungroup() |>
  mutate(
    day = rep(1:n_days, times = n_distinct(pool)),  # Creates 1:n_days for each pool
    type = "CENTURY"
  ) |>
  select(day, value, pool, type)

RC2m_DeadWoodPools_HARVARD <- as.data.frame(res_HARVARD$DecompositionPools) |>
  select(SmallBranchLitter, LargeWoodLitter, CoarseRootLitter)|>
  rename(SmallBranches = SmallBranchLitter,
         LargeWood = LargeWoodLitter,
         CoarseRoots =  CoarseRootLitter) |>
  mutate(Total = SmallBranches + LargeWood + CoarseRoots,
         day = 1:n_days) |>
  pivot_longer(cols = all_of(DeadWoodPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "MEDFATE")

## PLOTTING
g1 <- ggplot(bind_rows(RC2m_DeadWoodPools_DUKE, Rcentury_DeadWoodPools_DUKE) |>
         filter(pool != "Total"), 
       aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("MEDFATE" = 1, "CENTURY" = 2)) + 
  labs(title = "Dead wood Pools (DUKE)",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")

g2 <- ggplot(bind_rows(RC2m_DeadWoodPools_HARVARD, Rcentury_DeadWoodPools_HARVARD) |>
         filter(pool != "Total"), 
       aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.5) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("MEDFATE" = 1, "CENTURY" = 2)) + 
  labs(title = "Dead wood Pools (HARVARD)",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  # ylim(0, 500) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")

plot_grid(g1, g2, nrow = 2)
```