---
title: "Verification of carbon decomposition (DAYCENT)"
author: "Inés Delsman Valderrama (CREAF), Miquel De Cáceres (CREAF)"
date: "`r Sys.Date()`"
description: >
   Verification of carbon decomposition implemented from DAYCENT against the output from CENTURY (4.7)
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(medfate)
library(ggplot2)
library(kableExtra)
library(knitr)
library(tidyr)
library(dplyr)
library(RColorBrewer)
```

```{r, include=FALSE}
wd='/home/miquel/OneDrive/mcaceres_work/model_development/medfate_evaluation/RcenturyBenchmark/'
load(paste0(wd,"Rdata/RC2m_Process.RData"))
```



```{r, include=FALSE}
#' Function to transform Rcentury output for soil temperature and soil moisture from month to day resolution
env_daily<-function(out, yearIni = 2000) {
  yearMax <- yearIni + max(out_duke$time) - 1
  dateIni <- as.Date(paste0(yearIni, "-01-01"))
  dateFin <- as.Date(paste0(yearMax, "-12-31"))
  days <- seq(dateIni, dateFin, by="day")
  months <- cut(days, breaks="months")
  unique_months <- unique(months)
  # Trim out to the complete years
  out <- out_duke[1:length(unique_months), , drop = FALSE]
  df <- data.frame(date = days, Step = 1:length(days), SoilMoisture = NA, SoilTemperature = NA,
                   AirTemperature = NA, AirRelativeHumidity = NA)
  rel_mois <- out$asmos.1./ max(out$asmos.1.)
  for(i in 1:length(unique_months)) {
    sel_i <- months == unique_months[i]
    df$SoilMoisture[sel_i] <- rel_mois[i]
    df$SoilTemperature[sel_i] <- out$stemp[i]
  }
  return(df)
}

# Extract the third element from tree list
rcentury_tree <- tree$`3`

prd_daily_new<-function(out, rcentury_tree, yearIni = 2000) {
  species <- rcentury_tree$label
  yearMax <- yearIni + max(out_duke$time) - 1
  dateIni <- as.Date(paste0(yearIni, "-01-01"))
  dateFin <- as.Date(paste0(yearMax, "-12-31"))
  days <- seq(dateIni, dateFin, by="day")
  months <- cut(days, breaks="months")
  unique_months <- unique(months)
  leafdr <- c(rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(1)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(2)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(3)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(4)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(5)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(6)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(7)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(8)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(9)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(10)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(11)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(12)",1])
  wooddr2 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(2)",1] #monthly death rate fraction for fine root component
  wooddr3 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(3)",1] #monthly death rate fraction for fine branch component
  wooddr4 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(4)",1] #monthly death rate fraction for large wood component
  wooddr5 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(5)",1] #monthly death rate fraction for coarse root component
  # Trim out to the complete years
  out <- out_duke[1:length(unique_months), , drop = FALSE]
  # fbrchc -> C in forest system fine branch component (g m-2).
  # crootc -> C in forest system coarse root component (g m-2).
  # rlwodc -> C in forest system large wood component (g m-2).
  # frootc -> C in forest system fine root component (g m-2).
  # rleavc -> C in forest system leaf component (g m-2).
  
  df <- data.frame(date = days, Step = 1:length(days), Species = species, Leaves = NA, Twigs = NA, SmallBranches = NA, LargeWood = NA, CoarseRoots = NA, FineRoots = NA)
  for(i in 1:length(unique_months)) {
    sel_i <- months == unique_months[i]
    nmonth_days <- sum(sel_i)
    month_number <- as.numeric(format(days[sel_i], "%m")[1])
    df$Leaves[sel_i] <- out$rleavc[i]*leafdr[month_number]/nmonth_days
    df$Twigs[sel_i] <- 0
    df$SmallBranches[sel_i] <- out$fbrchc[i]*wooddr3/nmonth_days
    df$LargeWood[sel_i] <- out$rlwodc[i]*wooddr4/nmonth_days
    df$CoarseRoots[sel_i] <- out$crootc[i]*wooddr5/nmonth_days
    df$FineRoots[sel_i] <- out$frootc[i]*wooddr2/nmonth_days
  }
  return(df)
}

#' Function to extract decomposition parameters from 'tree' objects in Rcentury
sp_decomp_params <- function(rcentury_tree) {
   # % lignin leaves tree$`3`$df$WDLIG(1)
   # % lignin fine roots tree$`3`$df$WDLIG(2)
   # % lignin branches tree$`3`$df$WDLIG(3)
   # % lignin large wood tree$`3`$df$WDLIG(4)
   # % lignin coarse roots tree$`3`$df$WDLIG(5)
   # average 3-5 for sapwood lignin
   # 
   # read lignin/N ratios from
   # tree$cerfor(3,1,1) C/
     
  wdlig1 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(1)",1]
  wdlig2 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(2)",1]
  wdlig3 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(3)",1]
  wdlig4 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(4)",1]
  wdlig5 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(5)",1]
  cn_leaves <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,1,1)",1]
  cn_fineroots <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,2,1)",1]
  cn_branches <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,3,1)",1]
  cn_largewood <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,4,1)",1]
  cn_coarseroots <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,5,1)",1]
  cn_wood <- mean(c(cn_branches, cn_largewood, cn_coarseroots))
  leaf_c <- 0.5
  wood_c <- 0.5
  fineroot_c <- 0.5
  n_leaves <- (1/cn_leaves)*leaf_c
  n_fineroot <- (1/cn_fineroots)*fineroot_c
  n_wood <- (1/cn_wood)*wood_c
  df <- data.frame(Species  = rcentury_tree$label,
                   LeafLignin = wdlig1*100, # %
                   WoodLignin = mean(wdlig3, wdlig4, wdlig5)*100,# %
                   FineRootLignin = wdlig2*100,# %
                   Nleaf = n_leaves*1000, # mg / g
                   Nsapwood = n_wood*1000, # mg / g
                   Nfineroot = n_fineroot*1000) # mg / g
  return(df)
}

litterProduction <- prd_daily_new(out_duke, rcentury_tree)


# Extracts litter decomposition parameters
environmentalConditions <- env_daily(out_duke)
paramsLitterDecomposition <- sp_decomp_params(rcentury_tree)
```


## Initial surface/soil plant residue

In the Duke example, surface structural = 72 g m2. This value is applied to the leaves pool (strucc(1)).

Additionaly, soil structural = 57.75 g m2. This value is applied to the fine roots pools (strucc(2)).

```{r}
# Initialize litter pools
nlitter <- 1

strucc1 <- out_duke$strucc.1.[1]
strucc2 <- out_duke$strucc.2.[1]

litterData <- data.frame(Species = as.character(rep(paramsLitterDecomposition$Species[1], nlitter)),
                         Leaves = as.numeric(rep(strucc1, nlitter)),
                         Twigs = as.numeric(rep(0, nlitter)),
                         SmallBranches = as.numeric(rep(0, nlitter)),
                         LargeWood = as.numeric(rep(0, nlitter)),
                         CoarseRoots = as.numeric(rep(0, nlitter)),
                         FineRoots = as.numeric(rep(strucc2, nlitter)))
```

Snags are not relevant at this point in time. No changes will be made. 

```{r, include=FALSE}
# Initialize (empty) snag pools
nsnag <- 0
snagData<- data.frame(Species = as.character(rep(NA, nsnag)),
                      DBH = as.numeric(rep(NA, nsnag)),
                      Height = as.numeric(rep(NA, nsnag)),
                      DeathAge = as.numeric(rep(NA, nsnag)),
                      SmallBranches = as.numeric(rep(NA, nsnag)),
                      LargeWood = as.numeric(rep(NA, nsnag)))
```
## SOC pools

We took the SOC data from the site.100 file and the surface and soil metabolic data from the output file of our Rcentury simulation of the Duke site.

```{r}
# Initialize SOC pools
SOM1C1 <- subset(site_d$`organic matter`$df, field == "SOM1CI(1,1)")$value
SOM1C2 <- subset(site_d$`organic matter`$df, field == "SOM1CI(2,1)")$value
SOM2C1 <- subset(site_d$`organic matter`$df, field == "SOM2CI(1,1)")$value
SOM2C2 <- subset(site_d$`organic matter`$df, field == "SOM2CI(2,1)")$value
SOM3C <- subset(site_d$`organic matter`$df, field == "SOM3CI(1)")$value

Metabc1 <-   out_duke$metabc.1.[1]
Metabc2 <- out_duke$metabc.2.[1]

SOCData <- c(SurfaceMetabolic = Metabc1, BelowgroundMetabolic = Metabc2, SurfaceActive = SOM1C1, BelowgroundActive = SOM1C2, SurfaceSlow = SOM2C1, BelowgroundSlow = SOM2C2, BelowgroundPassive = SOM3C)
```

```{r, include=FALSE}
# Retrieve overall control variables (base annual decomposition rates and annual turnover rate)
control <- defaultControl()
baseAnnualRates <- control$decompositionAnnualBaseRates
annualTurnoverRate <- control$decompositionAnnualTurnoverRate
```

The base rates from RC2m have been changed to match the base rates from Rcentury. These can be found in the fix.100 file  under the name 'decx(x)'. As Rcentury does not distinguish between leaves, small branches etc., the base rate for surface structural has been applied to leaves while Twigs, Smallbranches and LargeWood remain unchanges. The same counts for the soil structural. Fineroots has been changed and Coarse roots remain the same.

```{r}
#Change base rates
baseAnnualRates <- c(SurfaceMetabolic = 14.8, BelowgroundMetabolic = 18.5, Leaves = 3.9 , FineRoots = 4.9, Twigs = 1.8,  SmallBranches = 1.5,LargeWood = 0.02 , CoarseRoots = 0.1,SurfaceActive = 6, BelowgroundActive = 7.3, SurfaceSlow = 0.03, BelowgroundSlow = 0.07, BelowgroundPassive = 0.0008)
# baseAnnualRates <- c(SurfaceMetabolic = 14.8, BelowgroundMetabolic = 18.5, Leaves = 3.9 , FineRoots = 4.9, Twigs = 3.9,  SmallBranches = 3.9,LargeWood = 3.9 , CoarseRoots = 3.9,SurfaceActive = 6, BelowgroundActive = 7.3, SurfaceSlow = 0.03, BelowgroundSlow = 0.07, BelowgroundPassive = 0.0008)

```

## Texture & pH

We took the the sand and clay values (and multiplied them by 100 to obtain the fractions) and the pH value from the site.100 file.

```{r}
# Retrieve soil site data
sand <- subset(site_d$`site and control`$df, field == "SAND")$value * 100
clay <- subset(site_d$`site and control`$df, field == "CLAY")$value * 100
soilPH <- subset(site_d$`site and control`$df, field == "PH")$value 
```

```{r, include=FALSE}
# Call DAYCENT
# IMPORTANT: Species should be the same in 'litterData', 'paramsLitterDecomposition' and 'litterProduction'
paramsAnatomy <- data.frame(Species = "DUKE",
                            WoodDensity = 0.7)
res <-decomposition_DAYCENT(snagData, litterData, SOCData,
                            paramsLitterDecomposition,
                            paramsAnatomy,
                            baseAnnualRates,
                            annualTurnoverRate,
                            environmentalConditions,
                            litterProduction,
                            sand,
                            clay,
                            soilPH)
```

## Results

The monthly results from the Rcentury simulation are transformed to daily data to make it comparable to the results from medfate. The results are further divided into total C stock, soil organic matter pools (SOM pools) and litter pools. 

### Total

The SOM and Litter pools are summed manually to calculate the total C stock for both models. For Rcentury, there is also the option to use the parameter 'tomres', which should already include the total C stock. In this case, the manual option was preferable to main.

```{r include=FALSE}
RC_total <- cbind(
  out_duke[, c("strucc.1.", "strucc.2.", "metabc.1.", "metabc.2.", "som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c")]
) |>
  mutate(Total = rowSums(across(c(strucc.1., strucc.2., metabc.1., metabc.2., som1c.1., som1c.2., som2c.1., som2c.2., som3c)), na.rm = TRUE)) |>
  select(Total)

RC_total <-RC_total |>
  tidyr::pivot_longer(cols = c(Total), 
                      names_to = "pool", 
                      values_to = "value")
```

```{r message=FALSE, warning=FALSE}
# Get dimensions
n_days <- nrow(res$DecompositionPools)  
n_months <- nrow(out_duke) 

# Create data frames with time indices
RC2m_duke <- data.frame(
  value = rowSums(res$DecompositionPools[,c(3,8:15)], na.rm = TRUE),
  day = 1:n_days,
  type = "RC2m"
)

# Calculate days per month (approximately equal)
days_per_month <- rep(round(n_days / n_months), n_months)
# Adjust last month to match exactly
days_per_month[length(days_per_month)] <- n_days - sum(days_per_month[-length(days_per_month)])

# Expand monthly data to daily
Rcentury_duke <- RC_total |>
  mutate(
    days = days_per_month,
    monthly_value = value  
  ) |>
  rowwise() |>
  mutate(
    daily_values = list(rep(monthly_value, days))
  ) |>
  unnest(daily_values) |>
  mutate(
    day = 1:n(),
    type = "Rcentury"
  ) |>
  select(day, value = daily_values, type)

# Combine
combined <- bind_rows(RC2m_duke, Rcentury_duke)

# Plot
ggplot(combined, aes(x = day, y = value, color = type)) +
  geom_line() +
  labs(title = "Total C stock",
       x = "Time (20 years in days)", 
       y = "C stocks (g/m2)") +
  ylim(1400,4500) +
  xlim(0,7671) +
  theme_minimal() 
  
```
This figure shows a similar pattern for both models: a small decrease in the beginning and then a an increase later on. However, RC2m increases faster than Rcentury and there is a ~500 g/m2 difference in total C stock.

### SOM pool analysis

To further analyze the results, all SOM pools are selected from the models results and compared.

```{r, include=FALSE}
# Get dimensions
RC_SOM <- cbind(
  out_duke[, c("som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c")]
) |>
  mutate(Total = rowSums(across(c(som1c.1., som1c.2., som2c.1., som2c.2., som3c)), na.rm = TRUE)) 

RC_SOM <-RC_SOM |>
  tidyr::pivot_longer(cols = c(som1c.1., som1c.2., som2c.1., som2c.2., som3c,Total), 
                      names_to = "pool", 
                      values_to = "value")
```

Then we rename the Century pool codes to the names used in Bonan. 
```{r, include=FALSE}
SOMPools <- c("SurfaceActive", "BelowgroundActive", "SurfaceSlow", "BelowgroundSlow", "BelowgroundPassive", "Total")

Rcentury_SOMpools <- RC_SOM |>
  mutate(
    pool = case_when(
      pool == "som1c.1." ~ "SurfaceActive",
      pool == "som1c.2." ~ "BelowgroundActive",
      pool == "som2c.1." ~ "SurfaceSlow",
      pool == "som2c.2." ~ "BelowgroundSlow",
      pool == "som3c" ~ "BelowgroundPassive",
      pool == "Total" ~ "Total"))

SOM_col <- brewer.pal(n = length(SOMPools), name = "Dark2")
```


```{r, echo=FALSE}
res_SOM <- res
res_SOM$DecompositionPools <- cbind(
  res_SOM$DecompositionPools,
  Total = rowSums(res$DecompositionPools[, 11:15], na.rm = TRUE)
  )

RC2m_SOMpools <- as.data.frame(res_SOM$DecompositionPools[, SOMPools]) |>
  mutate(day = 1:n_days, type = "RC2m")


RC2m_SOMpools <- RC2m_SOMpools |>
  pivot_longer(cols = all_of(SOMPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "RC2m")

reps <- round(n_days / n_months)

Rcentury_SOMpools_daily <- Rcentury_SOMpools |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  mutate(
    day = 1:n(),
    type = "Rcentury"
  ) |>
  select(day, value, pool, type)

# Combine
combined <- bind_rows(RC2m_SOMpools, Rcentury_SOMpools_daily)

ggplot(combined, aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("RC2m" = 1, "Rcentury" = 2)) + 
  labs(title = "SOM Pools",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  ylim(0, 4000) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")
```

This figure shows that the results of both the Rcentury model and RC2m model are very similar, which indicates that the simulation of the SOM pools in RC2m appears to function well.

### Litter pool analysis

```{r include=FALSE}
LitterPools <- c("SurfaceLitter", "BelowgroundLitter", "SurfaceMetabolic", "BelowgroundMetabolic", "Total")

Leaf_Litter <- cbind(
  out_duke[, c("strucc.1.","strucc.2.", "metabc.1.", "metabc.2.")]
) |>
  mutate(Total = rowSums(dplyr::across(c(strucc.1., strucc.2., metabc.1., metabc.2.)), na.rm = TRUE))

Rcentury_Litter <- Leaf_Litter|> 
  dplyr::select(
    V1  = "strucc.1.",           # Direct renaming
    V2 = "strucc.2.",           # Direct renaming  
    V3 = "metabc.1.",
    V4 = "metabc.2.",
    V5 = "Total",
  )

Rcentury_Litter <-Rcentury_Litter |>
  tidyr::pivot_longer(cols = c(V1,V2,V3,V4,V5), 
                      names_to = "pool", 
                      values_to = "value")

Rcentury_Litter <-  Rcentury_Litter |>
  mutate(
    pool = case_when(
      pool == "V1" ~ "SurfaceLitter",
      pool == "V2" ~ "BelowgroundLitter",
      pool == "V3" ~ "SurfaceMetabolic",
      pool == "V4" ~ "BelowgroundMetabolic",
      pool == "V5" ~ "Total"))
 
Litter_col <- brewer.pal(n = length(SOMPools), name = "Dark2")
```

```{r echo=FALSE}
res_all <- res
res_all$DecompositionPools <- cbind(
  res$DecompositionPools,
  SurfaceLitter = rowSums(res$DecompositionPools[, 3, drop=FALSE], na.rm = TRUE),
  BelowgroundLitter = rowSums(res$DecompositionPools[, 8, drop=FALSE], na.rm = TRUE),
  Total = rowSums(res$DecompositionPools[, c(3, 8:10), drop=FALSE], na.rm = TRUE)
  )

RC2m_LitterPools <- as.data.frame(res_all$DecompositionPools[, LitterPools]) |>
  mutate(day = 1:n_days, type = "RC2m")


RC2m_LitterPools <- RC2m_LitterPools |>
  pivot_longer(cols = all_of(LitterPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "RC2m")

Rcentury_LitterPools <- Rcentury_Litter |>
  group_by(pool) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  ungroup() |>
  mutate(
    day = rep(1:n_days, times = n_distinct(pool)),  # Creates 1:n_days for each pool
    type = "Rcentury"
  ) |>
  select(day, value, pool, type)

# Combine
combined <- bind_rows(RC2m_LitterPools, Rcentury_LitterPools)

ggplot(combined, aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("RC2m" = 1, "Rcentury" = 2)) + 
  labs(title = "Litter Pools",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  # ylim(0, 500) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")
```

Contrary to the SOM pools, RC2m shows a rapid increase in total and Surface litter after 2000 days. Soil litter increases as well, while the other pools remain near 0. In Rcentury, all pools decrease slowly and remain near 0. 


## Structural litter

The Century model only includes leaves and fine roots in the structural pools, while RC2m also includes fine branches, large wood litter and coarse roots. These are included in the Century model, but in the dead wood compartments. Two new simulations of the models are created to compare both models with all structural litter and with only leaves and fine roots.

```{r, echo =FALSE}
# Initialize litter pools
litterData <- data.frame(Species = as.character(rep(paramsLitterDecomposition$Species[1], nlitter)),
                         Leaves = as.numeric(rep(72, nlitter)),
                         Twigs = as.numeric(rep(0, nlitter)),
                         SmallBranches = as.numeric(rep(0, nlitter)),
                         LargeWood = as.numeric(rep(0, nlitter)),
                         CoarseRoots = as.numeric(rep(0, nlitter)),
                         FineRoots = as.numeric(rep(57.75, nlitter)))

# Initialize SOC pools
SOM1C1 <- subset(site_d$`organic matter`$df, field == "SOM1CI(1,1)")$value
SOM1C2 <- subset(site_d$`organic matter`$df, field == "SOM1CI(2,1)")$value
SOM2C1 <- subset(site_d$`organic matter`$df, field == "SOM2CI(1,1)")$value
SOM2C2 <- subset(site_d$`organic matter`$df, field == "SOM2CI(2,1)")$value
SOM3C <- subset(site_d$`organic matter`$df, field == "SOM3CI(1)")$value
Metabc1 <-   out_duke$metabc.1.[1]
Metabc2 <- out_duke$metabc.2.[1]

SOCData <- c(SurfaceMetabolic = Metabc1, BelowgroundMetabolic = Metabc2, SurfaceActive = SOM1C1, BelowgroundActive = SOM1C2,SurfaceSlow = SOM2C1, BelowgroundSlow = SOM2C2, BelowgroundPassive = SOM3C)

res <-decomposition_DAYCENT(snagData, litterData, SOCData,
                            paramsLitterDecomposition,
                            paramsAnatomy,
                            baseAnnualRates,
                            annualTurnoverRate,
                            environmentalConditions,
                            litterProduction,
                            sand,
                            clay,
                            soilPH)

```

### All structural litter

A new selection of litter data has been made, which adapts the SurfaceStructural pool to include C in dead fine branches and large wood, and the SoilStructural to include coarse roots. 

```{r include=FALSE}
Litter <- cbind(
  out_duke[, c("strucc.1.", "wood1c", "wood2c", "wood3c", "strucc.2.", "metabc.1.", "metabc.2.")]
) |>
  mutate(SurfaceStructural = rowSums(across(c(strucc.1.,wood1c,wood2c)), na.rm = TRUE)) |>
  mutate(SoilStructural = rowSums(across(c(strucc.2.,wood3c)), na.rm = TRUE)) |>
  mutate(Total = rowSums(across(c(strucc.1., wood1c, wood2c, wood3c, strucc.2., metabc.1., metabc.2.)), na.rm = TRUE))
```


```{r include=FALSE}
# If you want to add this back to original dataframe
Rcentury_Litter_all <- Litter|> 
  dplyr::select(
    V1  = "SurfaceStructural",           # Direct renaming
    V2 = "SoilStructural",           # Direct renaming  
    V3 = "metabc.1.",
    V4 = "metabc.2.",
    V5 = "Total",
  )

Rcentury_Litter_all <-Rcentury_Litter_all |>
  tidyr::pivot_longer(cols = c(V1,V2,V3,V4,V5), 
                      names_to = "pool", 
                      values_to = "value")

Rcentury_Litter_all <-  Rcentury_Litter_all |>
  mutate(
    pool = case_when(
      pool == "V1" ~ "SurfaceLitter",
      pool == "V2" ~ "BelowgroundLitter",
      pool == "V3" ~ "SurfaceMetabolic",
      pool == "V4" ~ "BelowgroundMetabolic",
      pool == "V5" ~ "Total"))
```

```{r echo=FALSE}
# Expand monthly data to daily
Rcentury_Litter_all <- Rcentury_Litter_all |>
  group_by(pool) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  ungroup() |>
  mutate(
    day = rep(1:n_days, times = n_distinct(pool)),  # Creates 1:n_days for each pool
    type = "Rcentury"
  ) |>
  select(day, value, pool, type)

combined <- bind_rows(RC2m_LitterPools, Rcentury_Litter_all)

ggplot(combined, aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = Litter_col) +
  scale_linetype_manual(values = c("RC2m" = 1, "Rcentury" = 2 )) + 
  labs(title = "Litter Pools",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  ylim(0, 850) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")
```

The total litter has improved on the previous version. There are still some differences between the Surface litter and Soil litter pools, but overall less severe than before. Both now also include a total of all pools.

#### Total

```{r include=FALSE}
RC_total_all <- cbind(
  out_duke[, c("strucc.1.", "strucc.2.", "metabc.1.", "metabc.2.", "som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c", "wood1c", "wood2c", "wood3c")]
) |>
  mutate(Total = rowSums(across(c(strucc.1., strucc.2., metabc.1., metabc.2., som1c.1., som1c.2., som2c.1., som2c.2., som3c,wood1c, wood2c, wood3c)), na.rm = TRUE)) |>
  select(Total)

RC_total_all <-RC_total_all |>
  tidyr::pivot_longer(cols = c(Total), 
                      names_to = "pool", 
                      values_to = "value")


Rcentury_all <- RC_total_all |>
  mutate(
    days = days_per_month,
    monthly_value = value  
  ) |>
  rowwise() |>
  mutate(
    daily_values = list(rep(monthly_value, days))
  ) |>
  unnest(daily_values) |>
  mutate(
    day = 1:n(),
    type = "Rcentury_all"
  ) |>
  select(day, value = daily_values, type)

# Combine
combined <- bind_rows(RC2m_duke, Rcentury_duke, Rcentury_all)

# Plot
ggplot(combined, aes(x = day, y = value, color = type)) +
  geom_line() +
  labs(title = "Total C stock",
       x = "Time (20 years in days)", 
       y = "C stocks (g/m2)") +
  ylim(0,4200) +
  xlim(0,7671) +
  theme_minimal() 
  
```
If we include the dead wood in Rcentury, we can see an improvement in the total C stock as well.

### Only leaves

In this simulation, only the leaf litter and !Coarse root! litter pools are selected and shown with the regular Rcentury run.

```{r include=FALSE}
LitterPools <- c("LeafLitter", "FineRootLitter", "SurfaceMetabolic", "BelowgroundMetabolic", "Total")
Litter_col <- brewer.pal(n = length(LitterPools), name = "Dark2")


# Get dimensions
RC_Litter <- cbind(
  out_duke[, c("strucc.1.","strucc.2.", "metabc.1.", "metabc.2.")]
) |>
  mutate(Total = rowSums(across(c(strucc.1., strucc.2., metabc.1., metabc.2.)), na.rm = TRUE))|>
  select(strucc.1., strucc.2., metabc.1., metabc.2., Total)

RC_Litter <-RC_Litter |>
  tidyr::pivot_longer(cols = c(strucc.1., strucc.2., metabc.1., metabc.2., Total), 
                      names_to = "pool", 
                      values_to = "value")

Rcentury_LeafLitter <-  RC_Litter |>
  mutate(
    pool = case_when(
      pool == "strucc.1." ~ "LeafLitter",
      pool == "strucc.2." ~ "FineRootLitter",
      pool == "metabc.1." ~ "SurfaceMetabolic",
      pool == "metabc.2." ~ "BelowgroundMetabolic",
      pool == "Total" ~ "Total"))

Rcentury_LeafLitter <- Rcentury_LeafLitter |>
  group_by(pool) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  ungroup() |>
  mutate(
    day = rep(1:n_days, times = n_distinct(pool)),  # Creates 1:n_days for each pool
    type = "Rcentury"
  ) |>
  select(day, value, pool, type)
```

```{r echo=FALSE}
res_leaves <- res
res_leaves$DecompositionPools <- cbind(
  res$DecompositionPools,
  Total = rowSums(res$DecompositionPools[, c(3,8:10)], na.rm = TRUE)
  )

RC2m_LeafLitter <- as.data.frame(res_leaves$DecompositionPools[, LitterPools]) |>
  mutate(day = 1:n_days, type = "RC2m")

RC2m_LeafLitter <- RC2m_LeafLitter |>
  pivot_longer(cols = all_of(LitterPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "RC2m")


combined <- bind_rows(RC2m_LeafLitter, Rcentury_LeafLitter)

ggplot(combined, aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = Litter_col) +
  scale_linetype_manual(values = c("RC2m" = 1, "Rcentury" = 2 )) + 
  labs(title = "Litter Pools",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  ylim(0, 250) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")

```

The difference between RC2m and Rcentury is much less in the structural litter pools (leaf litter) compared to the original litter pool analysis. The Rcentury results are slightly higher, but do show similar patternes.

# Harvard

```{r, include=FALSE}
#' Function to transform Rcentury output for soil temperature and soil moisture from month to day resolution
env_daily<-function(out, yearIni = 2000) {
  yearMax <- yearIni + max(out_harvard$time) - 1
  dateIni <- as.Date(paste0(yearIni, "-01-01"))
  dateFin <- as.Date(paste0(yearMax, "-12-31"))
  days <- seq(dateIni, dateFin, by="day")
  months <- cut(days, breaks="months")
  unique_months <- unique(months)
  # Trim out to the complete years
  out <- out_harvard[1:length(unique_months), , drop = FALSE]
  df <- data.frame(date = days, Step = 1:length(days), SoilMoisture = NA, SoilTemperature = NA,
                   AirTemperature = NA, AirRelativeHumidity = NA)
  rel_mois <- out$asmos.1./ max(out$asmos.1.)
  for(i in 1:length(unique_months)) {
    sel_i <- months == unique_months[i]
    df$SoilMoisture[sel_i] <- rel_mois[i]
    df$SoilTemperature[sel_i] <- out$stemp[i]
  }
  return(df)
}

# Extract the third element from tree list
rcentury_tree <- tree$`1`

prd_daily_new<-function(out, rcentury_tree, yearIni = 2000) {
  species <- rcentury_tree$label
  yearMax <- yearIni + max(out_duke$time) - 1
  dateIni <- as.Date(paste0(yearIni, "-01-01"))
  dateFin <- as.Date(paste0(yearMax, "-12-31"))
  days <- seq(dateIni, dateFin, by="day")
  months <- cut(days, breaks="months")
  unique_months <- unique(months)
  leafdr <- c(rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(1)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(2)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(3)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(4)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(5)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(6)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(7)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(8)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(9)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(10)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(11)",1],
              rcentury_tree$df[rcentury_tree$df[,2]=="LEAFDR(12)",1])
  wooddr2 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(2)",1] #monthly death rate fraction for fine root component
  wooddr3 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(3)",1] #monthly death rate fraction for fine branch component
  wooddr4 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(4)",1] #monthly death rate fraction for large wood component
  wooddr5 <- rcentury_tree$df[rcentury_tree$df[,2]=="WOODDR(5)",1] #monthly death rate fraction for coarse root component
  # Trim out to the complete years
  out <- out_harvard[1:length(unique_months), , drop = FALSE]
  # fbrchc -> C in forest system fine branch component (g m-2).
  # crootc -> C in forest system coarse root component (g m-2).
  # rlwodc -> C in forest system large wood component (g m-2).
  # frootc -> C in forest system fine root component (g m-2).
  # rleavc -> C in forest system leaf component (g m-2).
  
  df <- data.frame(date = days, Step = 1:length(days), Species = species, Leaves = NA, Twigs = NA, SmallBranches = NA, LargeWood = NA, CoarseRoots = NA, FineRoots = NA)
  for(i in 1:length(unique_months)) {
    sel_i <- months == unique_months[i]
    nmonth_days <- sum(sel_i)
    month_number <- as.numeric(format(days[sel_i], "%m")[1])
    df$Leaves[sel_i] <- out$rleavc[i]*leafdr[month_number]/nmonth_days
    df$Twigs[sel_i] <- 0
    df$SmallBranches[sel_i] <- out$fbrchc[i]*wooddr3/nmonth_days
    df$LargeWood[sel_i] <- out$rlwodc[i]*wooddr4/nmonth_days
    df$CoarseRoots[sel_i] <- out$crootc[i]*wooddr5/nmonth_days
    df$FineRoots[sel_i] <- out$frootc[i]*wooddr2/nmonth_days
  }
  return(df)
}

#' Function to extract decomposition parameters from 'tree' objects in Rcentury
sp_decomp_params <- function(rcentury_tree) {
   # % lignin leaves tree$`3`$df$WDLIG(1)
   # % lignin fine roots tree$`3`$df$WDLIG(2)
   # % lignin branches tree$`3`$df$WDLIG(3)
   # % lignin large wood tree$`3`$df$WDLIG(4)
   # % lignin coarse roots tree$`3`$df$WDLIG(5)
   # average 3-5 for sapwood lignin
   # 
   # read lignin/N ratios from
   # tree$cerfor(3,1,1) C/
     
  wdlig1 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(1)",1]
  wdlig2 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(2)",1]
  wdlig3 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(3)",1]
  wdlig4 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(4)",1]
  wdlig5 <- rcentury_tree$df[rcentury_tree$df[,2]=="WDLIG(5)",1]
  cn_leaves <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,1,1)",1]
  cn_fineroots <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,2,1)",1]
  cn_branches <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,3,1)",1]
  cn_largewood <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,4,1)",1]
  cn_coarseroots <- rcentury_tree$df[rcentury_tree$df[,2]=="CERFOR(3,5,1)",1]
  cn_wood <- mean(c(cn_branches, cn_largewood, cn_coarseroots))
  leaf_c <- 0.5
  wood_c <- 0.5
  fineroot_c <- 0.5
  n_leaves <- (1/cn_leaves)*leaf_c
  n_fineroot <- (1/cn_fineroots)*fineroot_c
  n_wood <- (1/cn_wood)*wood_c
  df <- data.frame(Species  = rcentury_tree$label,
                   LeafLignin = wdlig1*100, # %
                   WoodLignin = mean(wdlig3, wdlig4, wdlig5)*100,# %
                   FineRootLignin = wdlig2*100,# %
                   Nleaf = n_leaves*1000, # mg / g
                   Nsapwood = n_wood*1000, # mg / g
                   Nfineroot = n_fineroot*1000) # mg / g
  return(df)
}

litterProduction <- prd_daily_new(out_duke, rcentury_tree)


# Extracts litter decomposition parameters
environmentalConditions <- env_daily(out_duke)
paramsLitterDecomposition <- sp_decomp_params(rcentury_tree)
```

```{r include=FALSE}
# Initialize litter pools
nlitter <- 1

strucc1 <- out_harvard$strucc.1.[1]
strucc2 <- out_harvard$strucc.2.[1]

litterData <- data.frame(Species = as.character(rep(paramsLitterDecomposition$Species[1], nlitter)),
                         Leaves = as.numeric(rep(strucc1, nlitter)),
                         Twigs = as.numeric(rep(0, nlitter)),
                         SmallBranches = as.numeric(rep(0, nlitter)),
                         LargeWood = as.numeric(rep(0, nlitter)),
                         CoarseRoots = as.numeric(rep(0, nlitter)),
                         FineRoots = as.numeric(rep(strucc2, nlitter)))
```

```{r, include=FALSE}
# Initialize (empty) snag pools
nsnag <- 0
snagData<- data.frame(Species = as.character(rep(NA, nsnag)),
                      DBH = as.numeric(rep(NA, nsnag)),
                      Height = as.numeric(rep(NA, nsnag)),
                      DeathAge = as.numeric(rep(NA, nsnag)),
                      SmallBranches = as.numeric(rep(NA, nsnag)),
                      LargeWood = as.numeric(rep(NA, nsnag)))
```

```{r include=FALSE}
# Initialize SOC pools
SOM1C1 <- subset(site_h$`organic matter`$df, field == "SOM1CI(1,1)")$value
SOM1C2 <- subset(site_h$`organic matter`$df, field == "SOM1CI(2,1)")$value
SOM2C1 <- subset(site_h$`organic matter`$df, field == "SOM2CI(1,1)")$value
SOM2C2 <- subset(site_h$`organic matter`$df, field == "SOM2CI(2,1)")$value
SOM3C <- subset(site_h$`organic matter`$df, field == "SOM3CI(1)")$value

Metabc1 <- out_harvard$metabc.1.[1]
Metabc2 <- out_harvard$metabc.2.[1]

SOCData <- c(SurfaceMetabolic = Metabc1, BelowgroundMetabolic = Metabc2, SurfaceActive = SOM1C1, BelowgroundActive = SOM1C2, SurfaceSlow = SOM2C1, BelowgroundSlow = SOM2C2, BelowgroundPassive = SOM3C)
```

```{r, include=FALSE}
# Retrieve overall control variables (base annual decomposition rates and annual turnover rate)
control <- defaultControl()
baseAnnualRates <- control$decompositionAnnualBaseRates
annualTurnoverRate <- control$decompositionAnnualTurnoverRate
```

```{r include=FALSE}
#Change base rates
baseAnnualRates <- c(SurfaceMetabolic = 14.8, BelowgroundMetabolic = 18.5, Leaves = 3.9 , FineRoots = 4.9, Twigs = 1.8,  SmallBranches = 1.5,LargeWood = 0.02 , CoarseRoots = 0.1,SurfaceActive = 6, BelowgroundActive = 7.3, SurfaceSlow = 0.03, BelowgroundSlow = 0.07, BelowgroundPassive = 0.0008)

```

```{r include=FALSE}
# Retrieve soil site data
sand <- subset(site_h$`site and control`$df, field == "SAND")$value * 100
clay <- subset(site_h$`site and control`$df, field == "CLAY")$value * 100
soilPH <- subset(site_h$`site and control`$df, field == "PH")$value 
```

```{r, include=FALSE}
# Call DAYCENT
# IMPORTANT: Species should be the same in 'litterData', 'paramsLitterDecomposition' and 'litterProduction'
paramsAnatomy <- data.frame(Species = "HARVARD",
                            WoodDensity = 0.7)
res <-decomposition_DAYCENT(snagData, litterData, SOCData,
                            paramsLitterDecomposition,
                            paramsAnatomy,
                            baseAnnualRates,
                            annualTurnoverRate,
                            environmentalConditions,
                            litterProduction,
                            sand,
                            clay,
                            soilPH)
```

## Total

For this simulation, all structural litter has been taken into consideration by default. So for the total C stock, wood1c, wood2c, etc. have been included.

```{r echo=FALSE}
RC_total <- cbind(
  out_harvard[, c("strucc.1.", "strucc.2.", "metabc.1.", "metabc.2.", "som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c", "wood1c", "wood2c", "wood3c")]
) |>
  mutate(Total = rowSums(across(c(strucc.1., strucc.2., metabc.1., metabc.2., som1c.1., som1c.2., som2c.1., som2c.2., som3c,wood1c, wood2c, wood3c)), na.rm = TRUE)) |>
  select(Total)

RC_total <-RC_total |>
  tidyr::pivot_longer(cols = c(Total), 
                      names_to = "pool", 
                      values_to = "value")

n_days <- nrow(res$DecompositionPools)  
n_months <- nrow(out_harvard) 

# Create data frames with time indices
RC2m <- data.frame(
  value = rowSums(res$DecompositionPools, na.rm = TRUE),
  day = 1:n_days,
  type = "RC2m"
)

# Calculate days per month (approximately equal)
days_per_month <- rep(round(n_days / n_months), n_months)
# Adjust last month to match exactly
days_per_month[length(days_per_month)] <- n_days - sum(days_per_month[-length(days_per_month)])


Rcentury <- RC_total |>
  mutate(
    days = days_per_month,
    monthly_value = value  
  ) |>
  rowwise() |>
  mutate(
    daily_values = list(rep(monthly_value, days))
  ) |>
  unnest(daily_values) |>
  mutate(
    day = 1:n(),
    type = "Rcentury"
  ) |>
  select(day, value = daily_values, type)

# Combine
combined <- bind_rows(RC2m, Rcentury)

# Plot
ggplot(combined, aes(x = day, y = value, color = type)) +
  geom_line() +
  labs(title = "Total C stock",
       x = "Time (20 years in days)", 
       y = "C stocks (g/m2)") +
  ylim(0,6000) +
  xlim(0,7671) +
  theme_minimal() 
  
```

Here you can see 

## SOM pool analysis

```{r, include=FALSE}
# Get dimensions
RC_SOM <- cbind(
  out_harvard[, c("som1c.1.", "som1c.2.", "som2c.1.", "som2c.2.", "som3c")]
) |>
  mutate(Total = rowSums(across(c(som1c.1., som1c.2., som2c.1., som2c.2., som3c)), na.rm = TRUE)) 

RC_SOM <-RC_SOM |>
  tidyr::pivot_longer(cols = c(som1c.1., som1c.2., som2c.1., som2c.2., som3c,Total), 
                      names_to = "pool", 
                      values_to = "value")
```


```{r, include=FALSE}
SOMPools <- c("SurfaceActive", "BelowgroundActive", "SurfaceSlow", "BelowgroundSlow", "BelowgroundPassive", "Total")

Rcentury_SOMpools <- RC_SOM |>
  mutate(
    pool = case_when(
      pool == "som1c.1." ~ "SurfaceActive",
      pool == "som1c.2." ~ "BelowgroundActive",
      pool == "som2c.1." ~ "SurfaceSlow",
      pool == "som2c.2." ~ "BelowgroundSlow",
      pool == "som3c" ~ "BelowgroundPassive",
      pool == "Total" ~ "Total"))

SOM_col <- brewer.pal(n = length(SOMPools), name = "Dark2")
```


```{r, echo=FALSE}
res_SOM <- res
res_SOM$DecompositionPools <- cbind(
  res_SOM$DecompositionPools,
  Total = rowSums(res$DecompositionPools[, 11:15], na.rm = TRUE)
  )

RC2m_SOMpools <- as.data.frame(res_SOM$DecompositionPools[, SOMPools]) |>
  mutate(day = 1:n_days, type = "RC2m")


RC2m_SOMpools <- RC2m_SOMpools |>
  pivot_longer(cols = all_of(SOMPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "RC2m")

reps <- round(n_days / n_months)

Rcentury_SOMpools_daily <- Rcentury_SOMpools |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  mutate(
    day = 1:n(),
    type = "Rcentury"
  ) |>
  select(day, value, pool, type)

# Combine
combined <- bind_rows(RC2m_SOMpools, Rcentury_SOMpools_daily)

ggplot(combined, aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = SOM_col) +
  scale_linetype_manual(values = c("RC2m" = 1, "Rcentury" = 2)) + 
  labs(title = "SOM Pools",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  ylim(0, 5000) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")
```

## Litter

Again, all structural components are included here (so not just leaves and fine roots).

```{r include=FALSE}
Litter <- cbind(
  out_harvard[, c("strucc.1.", "wood1c", "wood2c", "wood3c", "strucc.2.", "metabc.1.", "metabc.2.")]
) |>
  mutate(SurfaceStructural = rowSums(across(c(strucc.1.,wood1c,wood2c)), na.rm = TRUE)) |>
  mutate(SoilStructural = rowSums(across(c(strucc.2.,wood3c)), na.rm = TRUE)) |>
  mutate(Total = rowSums(across(c(strucc.1., wood1c, wood2c, wood3c, strucc.2., metabc.1., metabc.2.)), na.rm = TRUE))

res_Litter <- res
res_Litter$DecompositionPools <- cbind(
res$DecompositionPools,
  SurfaceLitter = rowSums(res$DecompositionPools[, 3:6], na.rm = TRUE),
  BelowgroundLitter = rowSums(res$DecompositionPools[, 7:8], na.rm = TRUE),
  Total = rowSums(res$DecompositionPools[, 3:10], na.rm = TRUE)
  )
LitterPools <- c("SurfaceLitter", "BelowgroundLitter", "SurfaceMetabolic", "BelowgroundMetabolic", "Total")
Litter_col <- brewer.pal(n = length(LitterPools), name = "Dark2")
```


```{r include=FALSE}
# If you want to add this back to original dataframe
Rcentury_Litter <- Litter|> 
  dplyr::select(
    V1  = "SurfaceStructural",           # Direct renaming
    V2 = "SoilStructural",           # Direct renaming  
    V3 = "metabc.1.",
    V4 = "metabc.2.",
    V5 = "Total",
  )

Rcentury_Litter <-Rcentury_Litter |>
  tidyr::pivot_longer(cols = c(V1,V2,V3,V4,V5), 
                      names_to = "pool", 
                      values_to = "value")

Rcentury_Litter <-  Rcentury_Litter |>
  mutate(
    pool = case_when(
      pool == "V1" ~ "SurfaceLitter",
      pool == "V2" ~ "BelowgroundLitter",
      pool == "V3" ~ "SurfaceMetabolic",
      pool == "V4" ~ "BelowgroundMetabolic",
      pool == "V5" ~ "Total"))
```

```{r echo=FALSE}
# Expand monthly data to daily
Rcentury_Litter <- Rcentury_Litter |>
  group_by(pool) |>
  slice(rep(1:n(), each = reps, length.out = n_days)) |>
  ungroup() |>
  mutate(
    day = rep(1:n_days, times = n_distinct(pool)),  # Creates 1:n_days for each pool
    type = "Rcentury"
  ) |>
  select(day, value, pool, type)

RC2m_LitterPools <- as.data.frame(res_Litter$DecompositionPools[, LitterPools]) |>
  mutate(day = 1:n_days, type = "RC2m")


RC2m_LitterPools <- RC2m_LitterPools |>
  pivot_longer(cols = all_of(LitterPools), 
               names_to = "pool", 
               values_to = "value") |>
  mutate(type = "RC2m")

combined <- bind_rows(RC2m_LitterPools, Rcentury_Litter)

ggplot(combined, aes(x = day, y = value, color = pool, linetype = type)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = Litter_col) +
  scale_linetype_manual(values = c("RC2m" = 1, "Rcentury" = 2 )) + 
  labs(title = "Litter Pools",
       x = "Time (20 years in days)", 
       y = "Carbon Stock (g/m2)",
       color = "Pool",
       linetype = "Model") +
  ylim(0, 1100) +
  xlim(0,7671) +
  theme_minimal() +
  theme(legend.position = "right")
```

