---
title: "Evaluation of forest dynamics with IFN data"
author: "Miquel De Cáceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: TRUE
---

```{r setup, include=FALSE}
library(kableExtra)
library(sf)
library(knitr)
library(ggplot2)
library(cowplot)
library(tidyverse)
```

```{r, echo = FALSE}
plot_scatter_dbh<-function(dfdbh, points = TRUE){
g1<-ggplot(dfdbh, aes(x = pred, y= obs))
if(points) g1 <- g1 +  geom_point(size=0.3, col="dark gray", alpha = 0.3)
g1 <- g1 +  geom_abline(intercept = 0, slope = 1, col ="black")+
  xlab("Predicted diameter increment (cm/yr)")+
  ylab("Observed diameter increment (cm/yr)")+
  xlim(c(0,1.5))+ ylim(c(0,1.5))+
  theme_bw()
g2<-ggplot(dfdbh, aes(x = DBH3, y= error))
if(points) g2 <- g2 +  geom_point(size=0.3, col="dark gray", alpha = 0.3)
g2 <- g2 +  geom_abline(intercept = 0, slope = 0, col ="black")+
  geom_smooth(method="gam", formula = y ~ s(x, bs = "cs"), col="red")+
  xlab("Initial diameter (cm)")+
  ylab("Error diameter increment (cm/yr)")+
  ylim(c(-1.5,1.5))+xlim(c(0,50))+
  theme_bw()
  plot_grid(g1, g2, nrow=1)
}

plot_scatter_bai<-function(dfbai, quantity = "basal area increment", xylim = c(0,20), errorlim = c(-20,20)) {
g3<-ggplot(dfbai, aes(x = pred, y= obs))+
  geom_point(size=0.3, col="dark gray", alpha = 0.3)+
  geom_abline(intercept = 0, slope = 1, col ="black")+
  xlab(paste0("Predicted ", quantity," (m2/ha)"))+
  ylab(paste0("Observed ", quantity," (m2/ha)"))+
  xlim(xylim)+ ylim(xylim)+
  theme_bw()
g4<-ggplot(dfbai, aes(x = ini, y= error))+
  geom_point(size=0.3, col="dark gray", alpha = 0.3)+
  geom_abline(intercept = 0, slope = 0, col ="black")+
  geom_smooth(method=NULL, formula = y ~ x, col="red")+
  xlab("Initial basal area (m2/ha)")+
  ylab(paste0("Error ", quantity," (m2/ha)"))+
  xlim(c(0,30))+ylim(errorlim)+
  theme_bw()
  plot_grid(g3, g4, nrow=1)
}

plot_cov_clim_dbh<-function(dfdbh, points = TRUE){
  dfdbh$PPET[dfdbh$PPET>2] = 2
  df = bind_rows(data.frame(MAT = dfdbh$MAT, PPET = dfdbh$PPET, PAR = dfdbh$light, value = dfdbh$obs, type="Observed"),
                 data.frame(MAT = dfdbh$MAT, PPET = dfdbh$PPET, PAR = dfdbh$light, value = dfdbh$pred, type="Predicted"))
  g1<-ggplot(df, aes(x = PPET, y=value))
  if(points) g1 <- g1 + geom_point(aes(col = type), size=0.3, alpha = 0.3)
  g1 <- g1 + geom_smooth(aes(col = type), method="gam", formula = y ~ s(x, bs = "cs"))+
    ylab("Annual diameter increment (cm/yr)")+ ylim(c(0,1.5))+
    xlab("Moisture index (P/PET)")+
    scale_color_discrete("")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), legend.background = element_blank())
  g2<-ggplot(dfdbh, aes(x = PPET, y=error))
  if(points) g2 <- g2 + geom_point(col = "dark gray", size=0.3, alpha = 0.3)
  g2 <- g2 + geom_smooth(aes(x = PPET, y=error), formula = y ~ s(x, bs = "cs"), method="gam", col= "red")+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab("Annual diameter increment error (cm)")+ylim(c(-1.5,1.5))+
    xlab("Moisture index (P/PET)")+
    theme_bw()
  g3<-ggplot(df, aes(x = MAT, y=value))
  if(points) g3 <- g3 + geom_point(aes(col = type), size=0.3, alpha = 0.3)
  g3 <- g3 + geom_smooth(aes(col = type), method="gam", formula = y ~ s(x, bs = "cs"))+
    ylab("Annual diameter increment (cm/yr)")+ylim(c(0,1.5))+
    xlab("Mean annual temperature (ºC)")+
    scale_color_discrete("")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), legend.background = element_blank())
  g4<-ggplot(dfdbh, aes(x = MAT, y=error))
  if(points) g4 <- g4 + geom_point(col= "dark gray", size=0.3, alpha = 0.3)
  g4 <- g4 + geom_smooth(formula = y ~ s(x, bs = "cs"), method="gam", col= "red")+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab("Annual diameter increment error (cm)")+ylim(c(-1.5,1.5))+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()
  g5<-ggplot(df, aes(x = PAR, y=value))
  if(points) g5 <- g5 + geom_point(aes(col = type), size=0.3, alpha = 0.3)
  g5 <- g5 + geom_smooth(aes(col = type), method="gam", formula = y ~ s(x, bs = "cs"))+
    ylab("Annual diameter increment (cm/yr)")+ylim(c(0,1.5))+
    xlab("Available PAR (%)")+
    scale_color_discrete("")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), legend.background = element_blank())
  g6<-ggplot(dfdbh, aes(x = light, y=error))
  if(points) g6 <- g6 + geom_point(col= "dark gray", size=0.3, alpha = 0.3)
  g6 <- g6 + geom_smooth(formula = y ~ s(x, bs = "cs"), method="gam", col= "red")+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab("Annual diameter increment error (cm)")+ylim(c(-1.5,1.5))+
    xlab("Available PAR (%)")+
    theme_bw()
  plot_grid(g1, g2, g3, g4, g5,g6, nrow=3)
}

plot_cov_clim_bai<-function(dfbai, quantity = "Basal area increment", ylim = c(0,20), errorlim = c(-20,20)){
  dfbai$PPET[dfbai$PPET>2] = 2
  df = bind_rows(data.frame(MAT = dfbai$MAT, PPET = dfbai$PPET, value = dfbai$obs, type="Observed"),
                 data.frame(MAT = dfbai$MAT, PPET = dfbai$PPET, value = dfbai$pred, type="Predicted"))
  g1<-ggplot(df)+
    geom_point(aes(x = PPET, y=value, col = type), size=0.3, alpha = 0.3)+
    geom_smooth(aes(x = PPET, y=value, col = type), method="loess", formula = y ~ x)+
    ylab(paste0(quantity," (m2/ha)"))+ ylim(ylim)+
    xlab("Moisture index (P/PET)")+
    scale_color_discrete("")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), legend.background = element_blank())
  g2<-ggplot(dfbai)+
    geom_point(aes(x = PPET, y=error), col= "dark gray", size=0.3, alpha = 0.3)+
    geom_smooth(aes(x = PPET, y=error), method="loess", formula = y ~ x, col= "red")+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0(quantity," error (m2/ha)"))+ylim(errorlim)+
    xlab("Moisture index (P/PET)")+
    theme_bw()
  g3<-ggplot(df)+
    geom_point(aes(x = MAT, y=value, col = type), size=0.3, alpha = 0.3)+
    geom_smooth(aes(x = MAT, y=value, col = type), method="loess", formula = y ~ x)+
    ylab(paste0(quantity," (m2/ha)"))+ ylim(ylim)+
    xlab("Mean annual temperature (ºC)")+
    scale_color_discrete("")+theme_bw()+
    theme(legend.position = c(0.75,0.85), legend.background = element_blank())
    
  g4<-ggplot(dfbai)+
    geom_point(aes(x = MAT, y=error), col= "dark gray", size=0.3, alpha = 0.3)+
    geom_smooth(aes(x = MAT, y=error), method="loess", formula = y ~ x, col= "red")+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0(quantity," error (m2/ha)"))+ ylim(errorlim)+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()
  plot_grid(g1, g2, g3, g4, nrow=2)
}

bai_error_map <-function(data_bai){
  map_upper = 10
  map_lower = -10
  data_bai = data_bai[!is.na(data_bai$error),]
  data_bai$error[data_bai$error > map_upper] = map_upper
  data_bai$error[data_bai$error < map_lower] = map_lower
  ggplot(data_bai)+
    geom_sf(aes(col=error), size=0.5)+
    scale_color_gradient2("m2/ha",low="blue", mid="yellow", high="red")+
    theme_bw()+
    theme(panel.grid = element_blank())
}
```

## Data processing

```{r, echo = FALSE}
eval_res <- readRDS("~/OneDrive/Professional/MedfateWorks/IFNevaluation/Rdata/IFNeval_results.rds")

SpParams = eval_res$SpParams
target_spp = eval_res$target_spp
target_IDs = eval_res$target_IDs

dfdbh = eval_res$dfdbh
dfh = eval_res$dfh
dfbadead = eval_res$dfbadead
dfbai = eval_res$dfbai
dfbai_all = eval_res$dfbai_all
mapdata = eval_res$mapdata
medfate_ver = eval_res$package.version
```


We processed **`r length(target_IDs)`** repeated plots (A1 between IFN3 and IFN4) dominated by the target species (> 80% basal area corresponding to the target species) and without signs of management (cut trees).

Package version was **`r medfate_ver`**

Simulations were conducted between 2000/2001 (IFN3) and 2014/2016 (IFN4) depending on the target plot. Target species and key parameters values used in simulations:

```{r, echo = FALSE}

df = SpParams[SpParams$SpIndex %in% target_spp, c("Name", "Tmax_LAI", "Tmax_LAIsq", "WUE","Al2As","SLA", "RGRsapwoodmax", "SRsapwood", "fHDmin", "fHDmax")]
row.names(df) = NULL
df %>%
  kbl() %>%
  kable_styling()
```




## Growth of surviving trees

Comparison of growth of trees that survived in both observed and simulated data.

### Annual diameter increment (cm/yr)

Predictive capacity table:


```{r, echo = FALSE}
res_sp <- dfdbh %>% group_by(Name) %>%
  summarize(n = n(),
            Obs = mean(obs, na.rm=TRUE),
            Pred = mean(pred, na.rm=TRUE),
            Bias = mean(error, na.rm=TRUE),
            Biasrel = 100*Bias/Obs,
            MAE = mean(abs(error), na.rm=TRUE),
            MAErel = 100*MAE/Obs,
            R2 = cor(obs,pred)^2)
res_all <- dfdbh %>%
  summarize(Name = "All", 
            n = n(),
            Obs = mean(obs, na.rm=TRUE),
            Pred = mean(pred, na.rm=TRUE),
            Bias = mean(error, na.rm=TRUE),
            Biasrel = 100*Bias/Obs,
            MAE = mean(abs(error), na.rm=TRUE),
            MAErel = 100*MAE/Obs,
            R2 = cor(obs,pred)^2)

bind_rows(res_all,res_sp) %>%
  kbl() %>%
  kable_styling()
```

Predictive capacity plot:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
plot_scatter_dbh(dfdbh, points = TRUE)
```

Relationship between diameter increase and climatic variables (MAT, P/PET and available PAR):

```{r, echo = FALSE, fig.width=8, fig.height=12, warning=FALSE, message=FALSE}
plot_cov_clim_dbh(dfdbh, points = TRUE)
```


### Annual height increment (cm/yr)


Predictive capacity table:

```{r, echo = FALSE}
res_sp <- dfh %>% group_by(Name) %>%
  summarize(n = n(),
            Obs = mean(obs, na.rm=TRUE),
            Pred = mean(pred, na.rm=TRUE),
            Bias = mean(error, na.rm=TRUE),
            Biasrel = 100*Bias/Obs,
            MAE = mean(abs(error), na.rm=TRUE),
            MAErel = 100*MAE/Obs,
            R2 = cor(obs,pred)^2)
res_all <- dfh %>%
  summarize(Name = "All",
            n = n(),
            Obs = mean(obs, na.rm=TRUE),
            Pred = mean(pred, na.rm=TRUE),
            Bias = mean(error, na.rm=TRUE),
            Biasrel = 100*Bias/Obs,
            MAE = mean(abs(error), na.rm=TRUE),
            MAErel = 100*MAE/Obs,
            R2 = cor(obs,pred)^2)
bind_rows(res_all,res_sp) %>%
  kbl() %>%
  kable_styling()
```


### Stand-level basal area increment (m2/ha)

This comparison does not take into account changes in density. In other words, densities from IFN3 are used to calculate stand-level basal area of surviving trees. It is meant to evaluate the effect of diameter increment of surviving trees in terms of stand basal area increments.


Predictive capacity table:
```{r, echo = FALSE}
dfbai %>% group_by(Name) %>%
  summarize(n = n(),
            Obs = mean(obs, na.rm=TRUE),
            Pred = mean(pred, na.rm=TRUE),
            Bias = mean(error, na.rm=TRUE),
            Biasrel = 100*Bias/Obs,
            MAE = mean(abs(error), na.rm=TRUE),
            MAErel = 100*MAE/Obs,
            R2 = cor(obs,pred)^2) %>%
  kbl() %>%
  kable_styling()
```

Predictive capacity plot:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
plot_scatter_bai(dfbai[dfbai$Name=="0-All",])
```

Relationship between basal area increase and climatic variables (MAT and P/PET):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfbai[dfbai$Name=="0-All",])
```


Spatial error distribution:

```{r, echo = FALSE, fig.width=7, fig.height=7}
map_data_bai = mapdata
map_data_bai$error = dfbai$error[dfbai$Name=="0-All"]
bai_error_map(map_data_bai)
```


## Mortality (m2/ha)

Reduction of basal area due to trees (DBH > 5) that died during the IFN3-IFN4 period against model's mortality prediction.


Predictive capacity table:

```{r, echo = FALSE, warning=FALSE}
dfbadead %>% group_by(Name) %>%
  summarize(n = n(),
            Obs = mean(obs, na.rm=TRUE),
            Pred = mean(pred, na.rm=TRUE),
            Bias = mean(error, na.rm=TRUE),
            Biasrel = 100*Bias/Obs,
            MAE = mean(abs(error), na.rm=TRUE),
            MAErel = 100*MAE/Obs,
            R2 = cor(obs,pred)^2) %>%
  kbl() %>%
  kable_styling()
```

Predictive capacity plot:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message = FALSE}
plot_scatter_bai(dfbadead[dfbadead$Name=="0-All", ], 
                 quantity = "dead basal area", xylim = c(0,10), errorlim = c(-10,10)) 
```

Relationship between dead basal area and climatic variables (MAT and P/PET):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfbadead[dfbadead$Name=="0-All",], 
                  quantity = "Dead basal area", ylim = c(0,10), errorlim = c(-10,10))
```

Spatial distribution of errors:

```{r, echo = FALSE, fig.width=7, fig.height=7}
mapdata_badead = mapdata
mapdata_badead$error = dfbadead$error[dfbadead$Name=="0-All"]
bai_error_map(mapdata_badead)
```

## Overall stand-level basal area change (m2/ha)

This includes growth of surviving trees, mortality reductions and ingrowth (increase in trees with DBH>5) due to sapling growth (in the observed data, this includes also incorporation of trees into large diameter classes due to sampling design).



Predictive capacity table:

```{r, echo = FALSE}
dfbai_all %>% group_by(Name) %>%
  summarize(n = n(),
            Obs = mean(obs, na.rm=TRUE),
            Pred = mean(pred, na.rm=TRUE),
            Bias = mean(error, na.rm=TRUE),
            Biasrel = 100*Bias/Obs,
            MAE = mean(abs(error), na.rm=TRUE),
            MAErel = 100*MAE/Obs,
            R2 = cor(obs,pred)^2) %>%
  kbl() %>%
  kable_styling()
```


Predictive capacity plot:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message = FALSE}
plot_scatter_bai(dfbai_all[dfbai_all$Name=="0-All", ], 
                 quantity = "basal area change", xylim = c(-5,20), errorlim = c(-20,20)) 
```

Relationship between overall basal area change and climatic variables (MAT and P/PET):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfbai_all[dfbai_all$Name=="0-All",], 
                  quantity = "Basal area change", ylim = c(-5,20), errorlim = c(-20,20))
```


Spatial distribution of errors:

```{r, echo = FALSE, fig.width=7, fig.height=7, warning=FALSE, message=FALSE}

mapdata_bac = mapdata
mapdata_bac$error = dfbai_all$error[dfbai_all$Name=="0-All"]
bai_error_map(mapdata_bac)
```

## Detailed growth evaluation by species

```{r, echo=FALSE, results='asis', warning = FALSE, message=FALSE}
for(i in 1:length(target_spp)) {
  
  sp = target_spp[i]
  name = names(target_spp)[i]
  res <- knitr::knit_child('_IFNEvaluation_child.Rmd', quiet = TRUE)
  cat(res, sep = '\n')
}
```


