---
title: "Evaluation of regional-level forest dynamics with forest inventory data"
author: "Miquel De Cáceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: TRUE
params:
  minTreeRecords: 200
  minPlotRecords: 100
  minPlotRecordsDetailed: 100
  IPM: FALSE
---

```{r setup, include=FALSE}
library(kableExtra)
library(sf)
library(knitr)
library(medfate)
library(ggplot2)
library(cowplot)
library(tidyverse)
evalDIR = "~/OneDrive/Professional/MEDFATE_development/RegionalLevelEvaluation"
ipmDIR = "~/OneDrive/Professional/KnowledgeTransfer/CREAF/FORESFuture/Rdata/ipm/"
```

## Introduction

```{r, echo = FALSE}
ver <- packageVersion("medfate")
files <- c(paste0(evalDIR, "/Rdata/IFNeval_results_granier_IFN23_v", ver, ".rds"),
           paste0(evalDIR, "/Rdata/IFNeval_results_granier_IFN34_v", ver, ".rds"),
           paste0(evalDIR, "/Rdata/IFNeval_results_granier_IFN24_v", ver, ".rds"))
n <- length(files)
dfdbh_vec <- vector("list", n)
dfh_vec <- vector("list", n)
dfbai_sp_vec <- vector("list", n)
dfbai_plot_vec <- vector("list", n)
dfbadead_sp_vec <- vector("list", n)
dfbadead_plot_vec <- vector("list", n)
dfbarecr_sp_vec <- vector("list", n)
dfbarecr_plot_vec <- vector("list", n)
dfbatot_sp_vec <- vector("list", n)
dfbatot_plot_vec <- vector("list", n)
for(i in 1:n) {
  post <- readRDS(files[i])
  dfdbh_vec[[i]] <- post$dfdbh
  dfh_vec[[i]] <- post$dfh
  dfbai_sp_vec[[i]] <- post$dfbai_sp
  dfbai_plot_vec[[i]] <- post$dfbai_plot
  dfbadead_sp_vec[[i]] <- post$dfbadead_sp
  dfbadead_plot_vec[[i]] <- post$dfbadead_plot
  dfbarecr_sp_vec[[i]] <- post$dfbarecr_sp
  dfbarecr_plot_vec[[i]] <- post$dfbarecr_plot
  dfbatot_sp_vec[[i]] <- post$dfbatot_sp
  dfbatot_plot_vec[[i]] <- post$dfbatot_plot
}
dfdbh <- bind_rows(dfdbh_vec)
dfh <- bind_rows(dfh_vec)
dfbai_sp <- bind_rows(dfbai_sp_vec) |> sf::st_as_sf()
dfbai_plot<- bind_rows(dfbai_plot_vec) |> sf::st_as_sf()
dfbadead_sp <- bind_rows(dfbadead_sp_vec) |> sf::st_as_sf()
dfbadead_plot<- bind_rows(dfbadead_plot_vec) |> sf::st_as_sf()
dfbarecr_sp <- bind_rows(dfbarecr_sp_vec) |> sf::st_as_sf()
dfbarecr_plot<- bind_rows(dfbarecr_plot_vec) |> sf::st_as_sf()
dfbatot_sp <- bind_rows(dfbatot_sp_vec) |> sf::st_as_sf()
dfbatot_plot<- bind_rows(dfbatot_plot_vec) |> sf::st_as_sf()

load("~/OneDrive/EMFDatasets/PoliticalBoundaries/Products/Catalunya/catalonia.rdata")
cat_sfc = sf::st_as_sfc(cat.contour)
```

```{r, echo = FALSE}
evaluation_stats<-function(df) {
  df$version <- as.character(df$version)
  df_stats <- sf::st_drop_geometry(df) |>
    group_by(version, period, transpirationMode) |>
    summarize(n = n(),
              Obs = mean(obs, na.rm=TRUE),
              Pred = mean(pred, na.rm=TRUE),
              Bias = mean(error, na.rm=TRUE),
              Biasrel = 100*Bias/abs(Obs),
              MAE = mean(abs(error), na.rm=TRUE),
              MAErel = 100*MAE/abs(Obs),
              RMSQ = sqrt(mean(error^2, na.rm=TRUE)),
              RMSQrel = 100*RMSQ/abs(Obs),
              R2 = cor(obs,pred)^2, .groups="drop")
  return(as.data.frame(df_stats))
}
```

### Goal

The aim of this article is to provide an assessment of the performance of `fordyn` (using either the basic and advanced sub-model) for the prediction of forest dynamics in Catalonia (NE of Spain). To this aim, we simulate forest dynamics between surveys of the Spanish National Forest Inventory and compare the model predictions of forest growth against inventory data for a set of permanent plots. The evaluation focuses first on the growth (in diameter and height) of surviving trees, then turning the attention to the basal area of dead trees and overall changes in basal area.


```{r, echo = FALSE}
plot_scatter_tree<-function(df_bas, df_adv, points = TRUE, 
                            xylim = c(0,1.5), errorlim = c(-1,1),
                            initiallim = c(0,50), var = "diameter"){
  if(var=="diameter") {
    df_bas$ini <- df_bas$DBH_ini
    df_adv$ini <- df_adv$DBH_ini
  } else {
    df_bas$ini <- df_bas$H_ini
    df_adv$ini <- df_adv$H_ini
  }
g1<-ggplot(df_bas, aes(x = pred, y= obs))
if(points) {
  g1 <- g1 +  geom_point(aes(x = pred, y= obs), size=0.1, col="orange", alpha = 0.05, data = df_bas)
  g1 <- g1 +  geom_point(aes(x = pred, y= obs), size=0.1, col="light blue", alpha = 0.05, data = df_adv)
}
g1 <- g1 +  geom_abline(intercept = 0, slope = 1, col ="black")+
  xlab(paste0("Predicted ", var, " increment (cm/yr)"))+
  ylab(paste0("Observed ", var, " increment (cm/yr)"))+
  xlim(xylim)+ ylim(xylim)+
  theme_bw()+
    theme(panel.grid = element_blank())

g2<-ggplot(df_bas, aes(x = ini, y= error))
if(points) {
  g2 <- g2 +  geom_point( aes(x = ini, y= error), size=0.1, col="orange", alpha = 0.05, data = df_bas)
  g2 <- g2 +  geom_point( aes(x = ini, y= error), size=0.1, col="light blue", alpha = 0.05, data = df_adv)
}
g2 <- g2 +  geom_abline(intercept = 0, slope = 0, col ="black")+
  geom_smooth(method="gam", formula = y ~ s(x, bs = "cs"), 
              col="red", data = df_bas)+
  geom_smooth(method="gam", formula = y ~ s(x, bs = "cs"), 
              col="blue", data = df_adv)+
  xlab(paste0("Initial ", var, " (cm)"))+
  ylab(paste0("Error ", var, " increment (cm/yr)"))+
  ylim(errorlim)+xlim(initiallim)+
  theme_bw()+
    theme(panel.grid = element_blank())
plot_grid(g1, g2,nrow=1)
}
```

```{r, echo = FALSE}
plot_scatter_bai<-function(dfbai_bas, dfbai_adv, quantity = "basal area increment", xylim = c(0,2), errorlim = c(-1,1)) {
g1<-ggplot(dfbai_bas, aes(x = pred, y= obs))+
  geom_point(aes(x = pred, y= obs), size=0.3, col="orange", alpha = 0.3, data = dfbai_bas)+
  geom_point(aes(x = pred, y= obs), size=0.3, col="light blue", alpha = 0.3, data = dfbai_adv)+
  geom_abline(intercept = 0, slope = 1, col ="black")+
  xlab(paste0("Predicted ", quantity," (m2/ha/yr)"))+
  ylab(paste0("Observed ", quantity," (m2/ha/yr)"))+
  xlim(xylim)+ ylim(xylim)+
  theme_bw()+
    theme(panel.grid = element_blank())

g2<-ggplot(dfbai_bas, aes(x = BA_ini, y= error))+
  geom_point( aes(x = BA_ini, y= error), size=0.3, col="orange", alpha = 0.3, data = dfbai_bas)+
  geom_point( aes(x = BA_ini, y= error), size=0.3, col="light blue", alpha = 0.3, data = dfbai_adv)+
  geom_abline(intercept = 0, slope = 0, col ="black")+
  geom_smooth(method="loess", formula = y ~ x, 
              col="red", data = dfbai_bas)+
  geom_smooth(method="loess", formula = y ~ x, 
              col="blue", data = dfbai_adv)+
  xlab("Initial basal area (m2/ha)")+
  ylab(paste0("Error ", quantity," (m2/ha/yr)"))+
  xlim(c(0,30))+ylim(errorlim)+
  theme_bw()+
    theme(panel.grid = element_blank())
plot_grid(g1, g2, nrow=1)
}
```

```{r, echo = FALSE}
plot_cov_clim_tree<-function(df_bas, df_adv, points = TRUE, xylim = c(0,1), errorlim = c(-1,1), var = "diameter"){
  df_bas$PPET[df_bas$PPET>2] = 2
  df_adv$PPET[df_adv$PPET>2] = 2
  dfbas = bind_rows(data.frame(MAT = df_bas$MAT, PPET = df_bas$PPET, PAR = df_bas$light_ini, value = df_bas$obs, type="Observed"),
                    data.frame(MAT = df_bas$MAT, PPET = df_bas$PPET, PAR = df_bas$light_ini, value = df_bas$pred, type="Predicted"))
  dfadv = bind_rows(data.frame(MAT = df_adv$MAT, PPET = df_adv$PPET, PAR = df_adv$light_ini, value = df_adv$obs, type="Observed"),
                    data.frame(MAT = df_adv$MAT, PPET = df_adv$PPET, PAR = df_adv$light_ini, value = df_adv$pred, type="Predicted"))
  g1<-ggplot(dfbas, aes(x = PPET, y=value))
  if(points) {
    g1 <- g1 +  geom_point(aes(x = PPET, y= value), size=0.1, col="gray", alpha = 0.05, data = dfbas[dfbas$type=="Observed", ])
    g1 <- g1 +  geom_point(aes(x = PPET, y= value), size=0.1, col="orange", alpha = 0.05, data = dfbas[dfbas$type=="Predicted", ])
    g1 <- g1 +  geom_point(aes(x = PPET, y= value), size=0.1, col="light blue", alpha = 0.05, data = dfadv[dfadv$type=="Predicted", ])
  }
  g1 <- g1 + 
    geom_smooth(col = "black", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfadv[dfadv$type=="Predicted", ])+
    ylab(paste0("Annual ", var, " increment (cm/yr)"))+ ylim(xylim)+
    xlab("Moisture index (P/PET)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
  
  g2<-ggplot(df_bas, aes(x = PPET, y=error))
  if(points) {
    g2 <- g2 +  geom_point(aes(x = PPET, y= error), size=0.1, col="orange", alpha = 0.05, data = df_bas)
    g2 <- g2 +  geom_point(aes(x = PPET, y= error), size=0.1, col="light blue", alpha = 0.05, data = df_adv)
  }
  g2 <- g2 + geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = df_bas)+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = df_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0("Annual ", var, " increment error (cm)"))+ylim(errorlim)+
    xlab("Moisture index (P/PET)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  
  g3<-ggplot(dfbas, aes(x = MAT, y=value))
  if(points) {
    g3 <- g3 +  geom_point(aes(x = MAT, y= value), size=0.1, col="gray", alpha = 0.05, data = dfbas[dfbas$type=="Observed", ])
    g3 <- g3 +  geom_point(aes(x = MAT, y= value), size=0.1, col="orange", alpha = 0.05, data = dfbas[dfbas$type=="Predicted", ])
    g3 <- g3 +  geom_point(aes(x = MAT, y= value), size=0.1, col="light blue", alpha = 0.05, data = dfadv[dfadv$type=="Predicted", ])
  }
  g3 <- g3 + 
    geom_smooth(col = "black", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfadv[dfadv$type=="Predicted", ])+
    ylab(paste0("Annual ", var, " increment (cm/yr)"))+ylim(xylim)+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
  
  g4<-ggplot(df_bas, aes(x = MAT, y=error))
  if(points) {
    g4 <- g4 +  geom_point(aes(x = MAT, y= error), size=0.1, col="orange", alpha = 0.05, data = df_bas)
    g4 <- g4 +  geom_point(aes(x = MAT, y= error), size=0.1, col="light blue", alpha = 0.05, data = df_adv)
  }
  g4 <- g4 + geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = df_bas)+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = df_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0("Annual ", var, " increment error (cm)"))+ylim(errorlim)+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  
  g5<-ggplot(dfbas, aes(x = PAR, y=value))
  if(points) {
    g5 <- g5 +  geom_point(aes(x = PAR, y= value), size=0.1, col="gray", alpha = 0.05, data = dfbas[dfbas$type=="Observed", ])
    g5 <- g5 +  geom_point(aes(x = PAR, y= value), size=0.1, col="orange", alpha = 0.05, data = dfbas[dfbas$type=="Predicted", ])
    g5 <- g5 +  geom_point(aes(x = PAR, y= value), size=0.1, col="light blue", alpha = 0.05, data = dfadv[dfadv$type=="Predicted", ])
  }
  g5 <- g5 + 
    geom_smooth(col = "black", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfadv[dfadv$type=="Predicted", ])+
    ylab(paste0("Annual ", var," increment (cm/yr)"))+ylim(xylim)+
    xlab("Available PAR (%)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
  
  g6<-ggplot(df_bas, aes(x = light_ini, y=error))
  if(points) {
    g6 <- g6 +  geom_point(aes(x = light_ini, y= error), size=0.1, col="orange", alpha = 0.05, data = df_bas)
    g6 <- g6 +  geom_point(aes(x = light_ini, y= error), size=0.1, col="light blue", alpha = 0.05, data = df_adv)
  }
  g6 <- g6 + geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = df_bas)+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = df_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0("Annual ", var, " increment error (cm)"))+ylim(errorlim)+
    xlab("Available PAR (%)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  plot_grid(g1, g2, g3, g4, g5,g6, nrow=3)
}
```

```{r, echo = FALSE}
plot_cov_clim_bai<-function(dfbai_bas, dfbai_adv, quantity = "Basal area increment", ylim = c(0,2), errorlim = c(-2,2)){
  dfbai_bas$PPET[dfbai_bas$PPET>2] = 2
  dfbai_adv$PPET[dfbai_adv$PPET>2] = 2
  dfbas = bind_rows(data.frame(MAT = dfbai_bas$MAT, PPET = dfbai_bas$PPET, value = dfbai_bas$obs, type="Observed"),
                    data.frame(MAT = dfbai_bas$MAT, PPET = dfbai_bas$PPET, value = dfbai_bas$pred, type="Predicted"))
  dfadv = bind_rows(data.frame(MAT = dfbai_adv$MAT, PPET = dfbai_adv$PPET, value = dfbai_adv$obs, type="Observed"),
                    data.frame(MAT = dfbai_adv$MAT, PPET = dfbai_adv$PPET, value = dfbai_adv$pred, type="Predicted"))
  
  g1<-ggplot(dfbai_bas, aes(x = PPET, y=value))+
    geom_point(aes(x = PPET, y= value), size=0.3, col="gray", alpha = 0.3, data = dfbas[dfbas$type=="Observed", ])+
    geom_point(aes(x = PPET, y= value), size=0.3, col="orange", alpha = 0.3, data = dfbas[dfbas$type=="Predicted", ])+
    geom_point(aes(x = PPET, y= value), size=0.3, col="light blue", alpha = 0.3, data = dfadv[dfadv$type=="Predicted", ])+
    geom_smooth(col = "black", method="loess", formula = y ~ x, data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="loess", formula = y ~ x, data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="loess", formula = y ~ x, data = dfadv[dfadv$type=="Predicted", ])+
    ylab(paste0(quantity," (m2/ha/yr)"))+ ylim(ylim)+
    xlab("Moisture index (P/PET)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
  
  g2<-ggplot(dfbai_bas, aes(x = PPET, y=error))+
    geom_point(aes(x = PPET, y= error), size=0.3, col="orange", alpha = 0.3, data = dfbai_bas)+
    geom_point(aes(x = PPET, y= error), size=0.3, col="light blue", alpha = 0.3, data = dfbai_adv)+
    geom_smooth(aes(x = PPET, y=error), method="loess", formula = y ~ x, col= "red", data = dfbai_bas)+
    geom_smooth(aes(x = PPET, y=error), method="loess", formula = y ~ x, col= "blue", data = dfbai_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0(quantity," error (m2/ha/yr)"))+ylim(errorlim)+
    xlab("Moisture index (P/PET)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  
  g3<-ggplot(dfbai_bas, aes(x = MAT, y=value))+
    geom_point(aes(x = MAT, y= value), size=0.3, col="gray", alpha = 0.3, data = dfbas[dfbas$type=="Observed", ])+
    geom_point(aes(x = MAT, y= value), size=0.3, col="orange", alpha = 0.3, data = dfbas[dfbas$type=="Predicted", ])+
    geom_point(aes(x = MAT, y= value), size=0.3, col="light blue", alpha = 0.3, data = dfadv[dfadv$type=="Predicted", ])+
    geom_smooth(col = "black", method="loess", formula = y ~ x, data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="loess", formula = y ~ x, data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="loess", formula = y ~ x, data = dfadv[dfadv$type=="Predicted", ])+
    ylab(paste0(quantity," (m2/ha)"))+ ylim(ylim)+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
    
  g4<-ggplot(dfbai_bas, aes(x = MAT, y=error))+
    geom_point(aes(x = MAT, y= error), size=0.3, col="orange", alpha = 0.3, data = dfbai_bas)+
    geom_point(aes(x = MAT, y= error), size=0.3, col="light blue", alpha = 0.3, data = dfbai_adv)+
    geom_smooth(aes(x = MAT, y=error), method="loess", formula = y ~ x, col= "red", data = dfbai_bas)+
    geom_smooth(aes(x = MAT, y=error), method="loess", formula = y ~ x, col= "blue", data = dfbai_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0(quantity," error (m2/ha)"))+ ylim(errorlim)+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  plot_grid(g1, g2, g3, g4, nrow=2)
}
```

```{r, echo = FALSE}
bai_error_map <-function(data_bai, upper = 1, lower = -1){
  data_bai = data_bai[!is.na(data_bai$error),]
  data_bai$error[data_bai$error > upper] = upper
  data_bai$error[data_bai$error < lower] = lower
  ggplot(data_bai)+
    geom_sf(aes(col=error), size=0.5)+
    geom_sf(data = cat_sfc)+
    scale_color_fermenter("m2/ha/yr",palette = 1, breaks = c(lower,lower*0.5,lower*0.25,lower*0.1, 0, upper*0.1,upper*0.25,upper*0.5,upper), 
                       type = 'div', direction = 2, na.value = 'transparent')+
    # scale_color_gradient2("m2/ha",low="blue", mid="yellow", high="red")+
    theme_bw()+
    theme(panel.grid = element_blank())
    # theme(panel.grid = element_blank(), legend.position = c(0.85,0.25))
}
```

### Simulation procedure

We selected permanent plots between the second (IFN2) and the fourth (IFN4) without signs of management (i.e. the presence of stumps) and whose basal area did not decrease more than 10% between inventory surveys (to avoid the effect of disturbances).

Soil physical properties were drawn from SoilGrids (Hengl 2016), complemented by rock fragment content estimates derived from surface stoniness measurements in forest plots. 

`fordyn` simulations were conducted for different periods:

  + IFN2 - IFN3 (~ 10 years)
  + IFN3 - IFN4 (~ 15 years)
  + IFN2 - IFN4 (~ 25 years)

The actual simulated period varied depending on the sampling years of the target plot. Daily weather data were obtained via interpolation on plot's coordinates using package **meteoland**. 

Default species-specific parameters were modified using the results of the [meta-modelling exercise](https://emf-creaf.github.io/medfate/articles/parametrization/Metamodelling.html) and the [growth calibration exercise](https://emf-creaf.github.io/medfate/articles/parametrization/GrowthCalibration.html). These two exercises do not provide values for all the main species included here, so it is expected that evaluation results are worse for those species not included in those exercises.

Simulations were done for both the **basic** (i.e. Granier) and **advanced** (i.e. Sperry) transpiration/photosynthesis sub-models. On a server with 20 parallel threads, computational times for the longest simulation period (25 years) are around **4 hours** (i.e. 2.5 min/plot) for the **basic** sub-model, versus around **6 days** (i.e. 1.5 hr/plot) for the **advanced** submodel. 

In the following sections, we provide the **bias**, **mean absolute error** (in absolute and relative terms), **root mean squared error** (in absolute and relative terms) and **explained variance (R-squared)** of growth and mortality predictions at the tree-level and stand-level obtained by simulations with both sub-models. Scatter plots are provided for the IFN2-IFN4 simulation to represent the relationship between predicted and observed values, as well as the factors that may influence the direction and magnitude of prediction error (i.e. initial values, environmental conditions, ...).  

<!-- Detailed results of growth evaluation by species are provided in the last section. -->

## Growth of surviving trees

Comparison of diameter/height growth of trees (DBH >= 7.5) that survived between surveys.

### Annual diameter increment

Overall predictive capacity to predict diameter increase (cm/yr):
```{r, echo = FALSE}
dfdbh |> 
  evaluation_stats() |> 
  kbl() |>
  kable_styling()
```



Predictive capacity plots (IFN2-IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
dfdbh_bas <- dfdbh |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_tree(dfdbh_bas, dfdbh_bas, points = TRUE)
```


Relationship between diameter increase and climatic variables (MAT, P/PET and available PAR; IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=12, warning=FALSE, message=FALSE}
dfdbh_bas <- dfdbh |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_cov_clim_tree(dfdbh_bas, dfdbh_bas, points = TRUE)
```


### Annual height increment

Overall predictive capacity to predict height increase (cm/yr):
```{r, echo = FALSE, warning=FALSE, message=FALSE}
dfh %>% 
  evaluation_stats() %>%
  kbl() %>%
  kable_styling()
```
Predictive capacity plots (IFN2-IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
dfh_bas <- dfh |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_tree(dfh_bas, dfh_bas, points = TRUE, xylim = c(0,20), errorlim = c(-20,20), initiallim = c(0, 2000),
                  var = "height")
```


Relationship between height increase and climatic variables (MAT, P/PET and available PAR; IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=12, warning=FALSE, message=FALSE}
plot_cov_clim_tree(dfh_bas, dfh_bas, points = TRUE, xylim = c(0,20), errorlim = c(-20,20), var = "height")
```

### Stand-level basal area increment

Comparison of basal area increment of surviving trees does not take into account changes in density. In other words, densities from the first inventory are used to calculate stand-level basal area of surviving trees. Hence, the comparison is meant to evaluate the effect of diameter increment of surviving trees in terms of annual stand basal area increments (m2/ha/yr) for the period evaluated.

Predictive capacity table:
```{r, echo = FALSE, warning=FALSE}
dfbai_plot %>% evaluation_stats() %>% 
  kbl() %>%
  kable_styling()
```

Predictive capacity plots (IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
dfbai_bas <- dfbai_plot |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_bai(dfbai_bas, dfbai_bas)
```

Relationship between basal area increase and climatic variables (MAT and P/PET; IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfbai_bas, dfbai_bas, errorlim = c(-1,1), ylim = c(0,1))
```


Spatial error distribution (IFN2 - IFN4):

```{r, echo = FALSE, fig.width=9, fig.height=4}
p1<-bai_error_map(dfbai_bas)+labs(title="Basic")
p2<-bai_error_map(dfbai_bas)+labs(title="Advanced")
plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
```



## Mortality

Annual reduction of basal area (m2/ha/yr) due to trees (DBH >= 7.5) that died during the evaluation period against model's mortality prediction. In both cases, basal area is calculated using the initial diameter of the trees, so that density reductions are the only prediction that is actually evaluated (and not the possible growth of those trees during the simulation).


Overall predictive capacity:

```{r, echo = FALSE, warning=FALSE}
dfbadead_plot %>% evaluation_stats() %>% 
  kbl() %>%
  kable_styling()
```

Predictive capacity plots (IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message = FALSE}
dfdead_bas <- dfbadead_plot |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_bai(dfdead_bas, dfdead_bas,
                  quantity = "dead basal area", xylim = c(0,0.5), errorlim = c(-0.5,0.5))
```

Relationship between dead basal area and climatic variables (MAT and P/PET; IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfdead_bas, dfdead_bas,
                   quantity = "Dead basal area", ylim = c(0,0.5), errorlim = c(-0.5,0.5))
```

Spatial distribution of errors (IFN2 - IFN4):

```{r, echo = FALSE, fig.width=9, fig.height=4}
p1<-bai_error_map(dfdead_bas, upper = 0.5, lower = -0.5)+labs(title="Basic")
p2<-bai_error_map(dfdead_bas, upper = 0.5, lower = -0.5)+labs(title="Advanced")
plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
```



## Ingrowth

Annual increase of basal area (m2/ha/yr) due to ingrowth of trees with diameters between 7.5 cm and 12.5 cm during the evaluated period.


Predictive capacity:

```{r, echo = FALSE, warning=FALSE}
dfbarecr_plot %>% evaluation_stats()  %>% 
  kbl() %>%
  kable_styling()
```

Predictive capacity plots (IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message = FALSE}
dfrecr_bas <- dfbarecr_plot |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_bai(dfrecr_bas, dfrecr_bas,
                 quantity = "ingrowth basal area", xylim = c(0,0.15), errorlim = c(-0.15,0.15))
```

Relationship between ingrowth basal area and climatic variables (MAT and P/PET; IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfrecr_bas, dfrecr_bas,
                  quantity = "Ingrowth basal area", ylim = c(0,0.15), errorlim = c(-0.15,0.15))
```

Spatial distribution of errors (IFN2 - IFN4):

```{r, echo = FALSE, fig.width=9, fig.height=4}
p1<-bai_error_map(dfrecr_bas, upper = 0.5, lower = -0.5)+labs(title="Basic")
p2<-bai_error_map(dfrecr_bas, upper = 0.5, lower = -0.5)+labs(title="Advanced")
plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
```



## Overall stand-level basal area change

This includes annual changes in basal area (m2/ha/yr) due to growth of surviving trees, mortality reductions and ingrowth derived from sapling growth. In the observed data, basal area changes include also the incorporation of trees into large diameter classes that results from the variable-radius sampling design. Since it takes into account all processes together, this evaluation is the most rellevant of all.

Overall predictive capacity:

```{r, echo = FALSE, warning=FALSE}
dfbatot_plot %>% evaluation_stats() %>% 
  kbl() %>%
  kable_styling()
```


Predictive capacity plot (IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message = FALSE}
dftot_bas <- dfbatot_plot |> filter(transpirationMode == "Granier",
                             period == "IFN24")
p1<-plot_scatter_bai(dftot_bas, dftot_bas,
                 quantity = "basal area change", xylim = c(-0.5,2), errorlim = c(-2,2))
p1
```

Relationship between overall basal area change and climatic variables (MAT and P/PET; IFN2 - IFN4):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
p2<-plot_cov_clim_bai(dftot_bas, dftot_bas,
                  quantity = "Basal area change", ylim = c(-0.5,2), errorlim = c(-2,2))
p2
```

```{r, echo = FALSE}
p0<-cowplot::plot_grid(p1,p2, nrow=2, rel_heights = c(0.33,0.66))
ggsave(paste0(evalDIR,"/Plots/BAtot_eval.png"),p0, width = 8, height=10)
```

Spatial distribution of errors (IFN2 - IFN4):

```{r, echo = FALSE, fig.width=9, fig.height=4, warning=FALSE, message=FALSE}
p1<-bai_error_map(dftot_bas)+labs(title="Basic sub-model")
p2<-bai_error_map(dftot_bas)+labs(title="Advanced sub-model")
pm<-plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
ggsave(paste0(evalDIR,"/Plots/BAtot_error_map.png"),pm, width = 13, height=6, bg="white")
pm
```


<!-- ## Detailed evaluation by species -->

<!-- In the following we provide detailed evaluation results for the most important species. -->


<!-- ```{r, echo=FALSE, results='asis', warning = FALSE, message=FALSE} -->
<!-- # sp_n <- dfbai_all_bas[dfbai_all_bas$Name!="0-All",] %>% group_by(Name) %>% -->
<!-- #     summarize(n = n()) -->
<!-- # sp100 = sp_n$Name[sp_n$n>params$minPlotRecordsDetailed] -->
<!-- sp100 = c("Abies alba", "Fagus sylvatica", "Pinus halepensis", "Pinus nigra", -->
<!--           "Pinus sylvestris", "Pinus uncinata", "Quercus faginea", "Quercus ilex", -->
<!--           "Quercus pubescens", "Quercus suber") -->
<!-- target_spp = target_spp_bas[names(target_spp_bas) %in% sp100] -->
<!-- for(i in 1:length(target_spp)) { -->

<!--   sp = target_spp[i] -->
<!--   name = names(target_spp)[i] -->
<!--   res <- knitr::knit_child('_RLEvaluation_child.Rmd', quiet = TRUE) -->
<!--   cat(res, sep = '\n') -->
<!-- } -->
<!-- ``` -->


