---
title: "Evaluation of regional-level forest dynamics with forest inventory data"
author: "Miquel De Cáceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: TRUE
params:
  minTreeRecords: 200
  minPlotRecords: 100
  minPlotRecordsDetailed: 100
  IPM: FALSE
---

```{r setup, include=FALSE}
library(kableExtra)
library(sf)
library(knitr)
library(medfate)
library(ggplot2)
library(cowplot)
library(tidyverse)
evalDIR = "~/OneDrive/Professional/MedfateWorks/RegionalLevelEvaluation"
ipmDIR = "~/OneDrive/Professional/KnowledgeTransfer/CREAF/FORESFuture/Rdata/ipm/"
```

## Introduction

```{r, echo = FALSE}
ver <- packageVersion("medfate")
files <- c(paste0(evalDIR, "/Rdata/IFNeval_results_granier_IFN23_v", ver, ".rds"),
           paste0(evalDIR, "/Rdata/IFNeval_results_granier_IFN34_v", ver, ".rds"),
           paste0(evalDIR, "/Rdata/IFNeval_results_granier_IFN24_v", ver, ".rds"))
n <- length(files)
dfdbh_vec <- vector("list", n)
dfh_vec <- vector("list", n)
dfbai_sp_vec <- vector("list", n)
dfbai_plot_vec <- vector("list", n)
dfbadead_sp_vec <- vector("list", n)
dfbadead_plot_vec <- vector("list", n)
dfbarecr_sp_vec <- vector("list", n)
dfbarecr_plot_vec <- vector("list", n)
dfbatot_sp_vec <- vector("list", n)
dfbatot_plot_vec <- vector("list", n)
for(i in 1:n) {
  post <- readRDS(files[i])
  dfdbh_vec[[i]] <- post$dfdbh
  dfh_vec[[i]] <- post$dfh
  dfbai_sp_vec[[i]] <- post$dfbai_sp
  dfbai_plot_vec[[i]] <- post$dfbai_plot
  dfbadead_sp_vec[[i]] <- post$dfbadead_sp
  dfbadead_plot_vec[[i]] <- post$dfbadead_plot
  dfbarecr_sp_vec[[i]] <- post$dfbarecr_sp
  dfbarecr_plot_vec[[i]] <- post$dfbarecr_plot
  dfbatot_sp_vec[[i]] <- post$dfbatot_sp
  dfbatot_plot_vec[[i]] <- post$dfbatot_plot
}
dfdbh <- bind_rows(dfdbh_vec)
dfh <- bind_rows(dfh_vec)
dfbai_sp <- bind_rows(dfbai_sp_vec) |> sf::st_as_sf()
dfbai_plot<- bind_rows(dfbai_plot_vec) |> sf::st_as_sf()
dfbadead_sp <- bind_rows(dfbadead_sp_vec) |> sf::st_as_sf()
dfbadead_plot<- bind_rows(dfbadead_plot_vec) |> sf::st_as_sf()
dfbarecr_sp <- bind_rows(dfbarecr_sp_vec) |> sf::st_as_sf()
dfbarecr_plot<- bind_rows(dfbarecr_plot_vec) |> sf::st_as_sf()
dfbatot_sp <- bind_rows(dfbatot_sp_vec) |> sf::st_as_sf()
dfbatot_plot<- bind_rows(dfbatot_plot_vec) |> sf::st_as_sf()

# if(params$IPM) {
#   df_IPM = readRDS(paste0(ipmDIR, "IPM_evaluation_table.rds"))
#   IDs_IPM = row.names(df_IPM)
#   dfbai_all_bas <-dfbai_all_bas[dfbai_all_bas$ID %in% IDs_IPM,]
#   dfbai_all_adv <-dfbai_all_adv[dfbai_all_adv$ID %in% IDs_IPM,]
#   dfbai_bas <-dfbai_bas[dfbai_bas$ID %in% IDs_IPM,]
#   dfbai_adv <-dfbai_adv[dfbai_adv$ID %in% IDs_IPM,]
#   dfbarecr_bas <-dfbarecr_bas[dfbarecr_bas$ID %in% IDs_IPM,]
#   dfbarecr_adv <-dfbarecr_adv[dfbarecr_adv$ID %in% IDs_IPM,]
#   dfbadead_bas <-dfbadead_bas[dfbadead_bas$ID %in% IDs_IPM,]
#   dfbadead_adv <-dfbadead_adv[dfbadead_adv$ID %in% IDs_IPM,]
# }

load("~/OneDrive/EMFDatasets/PoliticalBoundaries/Products/Catalunya/catalonia.rdata")
cat_sfc = sf::st_as_sfc(cat.contour)
```

```{r, echo = FALSE}
  evaluation_stats<-function(df) {
    df_stats <- as.data.frame(sf::st_drop_geometry(df)) |>
      group_by(version, period,transpirationMode) |>
      summarize(n = n(),
                Obs = mean(obs, na.rm=TRUE),
                Pred = mean(pred, na.rm=TRUE),
                Bias = mean(error, na.rm=TRUE),
                Biasrel = 100*Bias/abs(Obs),
                MAE = mean(abs(error), na.rm=TRUE),
                MAErel = 100*MAE/abs(Obs),
                RMSQ = sqrt(mean(error^2, na.rm=TRUE)),
                RMSQrel = 100*RMSQ/abs(Obs),
                R2 = cor(obs,pred)^2, .groups="drop")
    return(df_stats)
  }
```

### Goal

The aim of this article is to provide an assessment of the performance of `fordyn` for the prediction of forest dynamics in Catalonia (NE of Spain). To this aim, we simulate forest dynamics between two surveys of the Spanish National Forest Inventory and compare the model predictions of forest growth against inventory data for a large set of permanent plots. The evaluation focuses first on the growth (in diameter and height) of surviving trees, then turning the attention to the basal area of dead trees and overall changes in basal area.



```{r, echo = FALSE}
plot_scatter_dbh<-function(dfdbh_bas, dfdbh_adv, points = TRUE){
g1<-ggplot(dfdbh_bas, aes(x = pred, y= obs))
if(points) {
  g1 <- g1 +  geom_point(aes(x = pred, y= obs), size=0.1, col="orange", alpha = 0.05, data = dfdbh_bas)
  g1 <- g1 +  geom_point(aes(x = pred, y= obs), size=0.1, col="light blue", alpha = 0.05, data = dfdbh_adv)
}
g1 <- g1 +  geom_abline(intercept = 0, slope = 1, col ="black")+
  xlab("Predicted diameter increment (cm/yr)")+
  ylab("Observed diameter increment (cm/yr)")+
  xlim(c(0,1.5))+ ylim(c(0,1.5))+
  theme_bw()+
    theme(panel.grid = element_blank())

g2<-ggplot(dfdbh_bas, aes(x = DBH_ini, y= error))
if(points) {
  g2 <- g2 +  geom_point( aes(x = DBH_ini, y= error), size=0.1, col="orange", alpha = 0.05, data = dfdbh_bas)
  g2 <- g2 +  geom_point( aes(x = DBH_ini, y= error), size=0.1, col="light blue", alpha = 0.05, data = dfdbh_adv)
}
g2 <- g2 +  geom_abline(intercept = 0, slope = 0, col ="black")+
  geom_smooth(method="gam", formula = y ~ s(x, bs = "cs"), 
              col="red", data = dfdbh_bas)+
  geom_smooth(method="gam", formula = y ~ s(x, bs = "cs"), 
              col="blue", data = dfdbh_adv)+
  xlab("Initial diameter (cm)")+
  ylab("Error diameter increment (cm/yr)")+
  ylim(c(-1.0,1.0))+xlim(c(0,50))+
  theme_bw()+
    theme(panel.grid = element_blank())
plot_grid(g1, g2,nrow=1)
}

plot_scatter_bai<-function(dfbai_bas, dfbai_adv, quantity = "basal area increment", xylim = c(0,2), errorlim = c(-1,1)) {
g1<-ggplot(dfbai_bas, aes(x = pred, y= obs))+
  geom_point(aes(x = pred, y= obs), size=0.3, col="orange", alpha = 0.3, data = dfbai_bas)+
  geom_point(aes(x = pred, y= obs), size=0.3, col="light blue", alpha = 0.3, data = dfbai_adv)+
  geom_abline(intercept = 0, slope = 1, col ="black")+
  xlab(paste0("Predicted ", quantity," (m2/ha/yr)"))+
  ylab(paste0("Observed ", quantity," (m2/ha/yr)"))+
  xlim(xylim)+ ylim(xylim)+
  theme_bw()+
    theme(panel.grid = element_blank())

g2<-ggplot(dfbai_bas, aes(x = BA_ini, y= error))+
  geom_point( aes(x = BA_ini, y= error), size=0.3, col="orange", alpha = 0.3, data = dfbai_bas)+
  geom_point( aes(x = BA_ini, y= error), size=0.3, col="light blue", alpha = 0.3, data = dfbai_adv)+
  geom_abline(intercept = 0, slope = 0, col ="black")+
  geom_smooth(method="loess", formula = y ~ x, 
              col="red", data = dfbai_bas)+
  geom_smooth(method="loess", formula = y ~ x, 
              col="blue", data = dfbai_adv)+
  xlab("Initial basal area (m2/ha)")+
  ylab(paste0("Error ", quantity," (m2/ha/yr)"))+
  xlim(c(0,30))+ylim(errorlim)+
  theme_bw()+
    theme(panel.grid = element_blank())
plot_grid(g1, g2, nrow=1)
}

plot_cov_clim_dbh<-function(dfdbh_bas, dfdbh_adv, points = TRUE){
  dfdbh_bas$PPET[dfdbh_bas$PPET>2] = 2
  dfdbh_adv$PPET[dfdbh_adv$PPET>2] = 2
  dfbas = bind_rows(data.frame(MAT = dfdbh_bas$MAT, PPET = dfdbh_bas$PPET, PAR = dfdbh_bas$light_ini, value = dfdbh_bas$obs, type="Observed"),
                    data.frame(MAT = dfdbh_bas$MAT, PPET = dfdbh_bas$PPET, PAR = dfdbh_bas$light_ini, value = dfdbh_bas$pred, type="Predicted"))
  dfadv = bind_rows(data.frame(MAT = dfdbh_adv$MAT, PPET = dfdbh_adv$PPET, PAR = dfdbh_adv$light_ini, value = dfdbh_adv$obs, type="Observed"),
                    data.frame(MAT = dfdbh_adv$MAT, PPET = dfdbh_adv$PPET, PAR = dfdbh_adv$light_ini, value = dfdbh_adv$pred, type="Predicted"))
  g1<-ggplot(dfbas, aes(x = PPET, y=value))
  if(points) {
    g1 <- g1 +  geom_point(aes(x = PPET, y= value), size=0.1, col="gray", alpha = 0.05, data = dfbas[dfbas$type=="Observed", ])
    g1 <- g1 +  geom_point(aes(x = PPET, y= value), size=0.1, col="orange", alpha = 0.05, data = dfbas[dfbas$type=="Predicted", ])
    g1 <- g1 +  geom_point(aes(x = PPET, y= value), size=0.1, col="light blue", alpha = 0.05, data = dfadv[dfadv$type=="Predicted", ])
  }
  g1 <- g1 + 
    geom_smooth(col = "black", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfadv[dfadv$type=="Predicted", ])+
    ylab("Annual diameter increment (cm/yr)")+ ylim(c(0,1.0))+
    xlab("Moisture index (P/PET)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
  
  g2<-ggplot(dfdbh_bas, aes(x = PPET, y=error))
  if(points) {
    g2 <- g2 +  geom_point(aes(x = PPET, y= error), size=0.1, col="orange", alpha = 0.05, data = dfdbh_bas)
    g2 <- g2 +  geom_point(aes(x = PPET, y= error), size=0.1, col="light blue", alpha = 0.05, data = dfdbh_adv)
  }
  g2 <- g2 + geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfdbh_bas)+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfdbh_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab("Annual diameter increment error (cm)")+ylim(c(-1.0,1.0))+
    xlab("Moisture index (P/PET)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  
  g3<-ggplot(dfbas, aes(x = MAT, y=value))
  if(points) {
    g3 <- g3 +  geom_point(aes(x = MAT, y= value), size=0.1, col="gray", alpha = 0.05, data = dfbas[dfbas$type=="Observed", ])
    g3 <- g3 +  geom_point(aes(x = MAT, y= value), size=0.1, col="orange", alpha = 0.05, data = dfbas[dfbas$type=="Predicted", ])
    g3 <- g3 +  geom_point(aes(x = MAT, y= value), size=0.1, col="light blue", alpha = 0.05, data = dfadv[dfadv$type=="Predicted", ])
  }
  g3 <- g3 + 
    geom_smooth(col = "black", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfadv[dfadv$type=="Predicted", ])+
    ylab("Annual diameter increment (cm/yr)")+ylim(c(0,1.0))+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
  
  g4<-ggplot(dfdbh_bas, aes(x = MAT, y=error))
  if(points) {
    g4 <- g4 +  geom_point(aes(x = MAT, y= error), size=0.1, col="orange", alpha = 0.05, data = dfdbh_bas)
    g4 <- g4 +  geom_point(aes(x = MAT, y= error), size=0.1, col="light blue", alpha = 0.05, data = dfdbh_adv)
  }
  g4 <- g4 + geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfdbh_bas)+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfdbh_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab("Annual diameter increment error (cm)")+ylim(c(-1.0,1.0))+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  
  g5<-ggplot(dfbas, aes(x = PAR, y=value))
  if(points) {
    g5 <- g5 +  geom_point(aes(x = PAR, y= value), size=0.1, col="gray", alpha = 0.05, data = dfbas[dfbas$type=="Observed", ])
    g5 <- g5 +  geom_point(aes(x = PAR, y= value), size=0.1, col="orange", alpha = 0.05, data = dfbas[dfbas$type=="Predicted", ])
    g5 <- g5 +  geom_point(aes(x = PAR, y= value), size=0.1, col="light blue", alpha = 0.05, data = dfadv[dfadv$type=="Predicted", ])
  }
  g5 <- g5 + 
    geom_smooth(col = "black", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfadv[dfadv$type=="Predicted", ])+
    ylab("Annual diameter increment (cm/yr)")+ylim(c(0,1.0))+
    xlab("Available PAR (%)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
  
  g6<-ggplot(dfdbh_bas, aes(x = light_ini, y=error))
  if(points) {
    g6 <- g6 +  geom_point(aes(x = light_ini, y= error), size=0.1, col="orange", alpha = 0.05, data = dfdbh_bas)
    g6 <- g6 +  geom_point(aes(x = light_ini, y= error), size=0.1, col="light blue", alpha = 0.05, data = dfdbh_adv)
  }
  g6 <- g6 + geom_smooth(col = "red", method="gam", formula = y ~ s(x, bs = "cs"), data = dfdbh_bas)+
    geom_smooth(col = "blue", method="gam", formula = y ~ s(x, bs = "cs"), data = dfdbh_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab("Annual diameter increment error (cm)")+ylim(c(-1.0,1.0))+
    xlab("Available PAR (%)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  plot_grid(g1, g2, g3, g4, g5,g6, nrow=3)
}

plot_cov_clim_bai<-function(dfbai_bas, dfbai_adv, quantity = "Basal area increment", ylim = c(0,2), errorlim = c(-2,2)){
  dfbai_bas$PPET[dfbai_bas$PPET>2] = 2
  dfbai_adv$PPET[dfbai_adv$PPET>2] = 2
  dfbas = bind_rows(data.frame(MAT = dfbai_bas$MAT, PPET = dfbai_bas$PPET, value = dfbai_bas$obs, type="Observed"),
                    data.frame(MAT = dfbai_bas$MAT, PPET = dfbai_bas$PPET, value = dfbai_bas$pred, type="Predicted"))
  dfadv = bind_rows(data.frame(MAT = dfbai_adv$MAT, PPET = dfbai_adv$PPET, value = dfbai_adv$obs, type="Observed"),
                    data.frame(MAT = dfbai_adv$MAT, PPET = dfbai_adv$PPET, value = dfbai_adv$pred, type="Predicted"))
  
  g1<-ggplot(dfbai_bas, aes(x = PPET, y=value))+
    geom_point(aes(x = PPET, y= value), size=0.3, col="gray", alpha = 0.3, data = dfbas[dfbas$type=="Observed", ])+
    geom_point(aes(x = PPET, y= value), size=0.3, col="orange", alpha = 0.3, data = dfbas[dfbas$type=="Predicted", ])+
    geom_point(aes(x = PPET, y= value), size=0.3, col="light blue", alpha = 0.3, data = dfadv[dfadv$type=="Predicted", ])+
    geom_smooth(col = "black", method="loess", formula = y ~ x, data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="loess", formula = y ~ x, data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="loess", formula = y ~ x, data = dfadv[dfadv$type=="Predicted", ])+
    ylab(paste0(quantity," (m2/ha/yr)"))+ ylim(ylim)+
    xlab("Moisture index (P/PET)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
  
  g2<-ggplot(dfbai_bas, aes(x = PPET, y=error))+
    geom_point(aes(x = PPET, y= error), size=0.3, col="orange", alpha = 0.3, data = dfbai_bas)+
    geom_point(aes(x = PPET, y= error), size=0.3, col="light blue", alpha = 0.3, data = dfbai_adv)+
    geom_smooth(aes(x = PPET, y=error), method="loess", formula = y ~ x, col= "red", data = dfbai_bas)+
    geom_smooth(aes(x = PPET, y=error), method="loess", formula = y ~ x, col= "blue", data = dfbai_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0(quantity," error (m2/ha/yr)"))+ylim(errorlim)+
    xlab("Moisture index (P/PET)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  
  g3<-ggplot(dfbai_bas, aes(x = MAT, y=value))+
    geom_point(aes(x = MAT, y= value), size=0.3, col="gray", alpha = 0.3, data = dfbas[dfbas$type=="Observed", ])+
    geom_point(aes(x = MAT, y= value), size=0.3, col="orange", alpha = 0.3, data = dfbas[dfbas$type=="Predicted", ])+
    geom_point(aes(x = MAT, y= value), size=0.3, col="light blue", alpha = 0.3, data = dfadv[dfadv$type=="Predicted", ])+
    geom_smooth(col = "black", method="loess", formula = y ~ x, data = dfbas[dfbas$type=="Observed", ])+
    geom_smooth(col = "red", method="loess", formula = y ~ x, data = dfbas[dfbas$type=="Predicted", ])+
    geom_smooth(col = "blue", method="loess", formula = y ~ x, data = dfadv[dfadv$type=="Predicted", ])+
    ylab(paste0(quantity," (m2/ha)"))+ ylim(ylim)+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()+
    theme(legend.position = c(0.75,0.85), 
          legend.background = element_blank(),
          panel.grid = element_blank())
    
  g4<-ggplot(dfbai_bas, aes(x = MAT, y=error))+
    geom_point(aes(x = MAT, y= error), size=0.3, col="orange", alpha = 0.3, data = dfbai_bas)+
    geom_point(aes(x = MAT, y= error), size=0.3, col="light blue", alpha = 0.3, data = dfbai_adv)+
    geom_smooth(aes(x = MAT, y=error), method="loess", formula = y ~ x, col= "red", data = dfbai_bas)+
    geom_smooth(aes(x = MAT, y=error), method="loess", formula = y ~ x, col= "blue", data = dfbai_adv)+
    geom_abline(intercept = 0, slope = 0, col ="black")+
    ylab(paste0(quantity," error (m2/ha)"))+ ylim(errorlim)+
    xlab("Mean annual temperature (ºC)")+
    theme_bw()+
    theme(panel.grid = element_blank())
  plot_grid(g1, g2, g3, g4, nrow=2)
}

bai_error_map <-function(data_bai, upper = 1, lower = -1){
  data_bai = data_bai[!is.na(data_bai$error),]
  data_bai$error[data_bai$error > upper] = upper
  data_bai$error[data_bai$error < lower] = lower
  ggplot(data_bai)+
    geom_sf(aes(col=error), size=0.5)+
    geom_sf(data = cat_sfc)+
    scale_color_fermenter("m2/ha",palette = 1, breaks = c(lower,lower*0.5,lower*0.25,lower*0.1, 0, upper*0.1,upper*0.25,upper*0.5,upper), 
                       type = 'div', direction = 2, na.value = 'transparent')+
    # scale_color_gradient2("m2/ha",low="blue", mid="yellow", high="red")+
    theme_bw()+
    theme(panel.grid = element_blank())
    # theme(panel.grid = element_blank(), legend.position = c(0.85,0.25))
}
```

### Simulation procedure


We selected repeated plots (A1 between IFN2 and IFN4) without signs of management (cut trees) and whose basal area did not decrease between inventory surveys.

Soil physical properties were drawn from SoilGrids (Hengl 2016), complemented by rock fragment content estimates derived from surface stoniness measurements in forest plots. Simulations were conducted for different periods:

  + IFN2 - IFN3 (~ 10 years)
  + IFN3 - IFN4 (~ 15 years)
  + IFN2 - IFN4 (~ 25 years)

Actual simulated years depending on the sampling years of the target plot. Daily weather data were obtained via interpolation on plot's coordinates using package **meteoland**. 

Default species-specific parameters were modified using the results of the [meta-modelling exercise](https://emf-creaf.github.io/medfate/articles/parametrization/Metamodelling.html) and the [growth calibration exercise](https://emf-creaf.github.io/medfate/articles/parametrization/GrowthCalibration.html). These two exercises do not provide values for all the main species included here, so it is expected that evaluation results are worse for those species not included in those exercises.

Simulations were done for both the **basic** and **advanced** transpiration/photosynthesis sub-models. On a server with 20 parallel threads, computational times for the longest simulation (IFN2 - IFN4) are around **5 hours** (i.e. 2 min/plot) for the **basic** sub-model, versus around **6 days** (i.e. 1 hr/plot) for the **advanced** submodel. 

In the following sections, we provide the **bias**, **mean absolute error** (in absolute and relative terms), **root mean squared error** (in absolute and relative terms) and **explained variance (R-squared)** of growth and mortality predictions at the tree-level and stand-level obtained by simulations with both sub-models. Scatter plots are provided to represent the relationship between predicted and observed values, as well as the factors that may influence the direction and magnitude of prediction error (i.e. initial values, environmental conditions, ...).  

<!-- Detailed results of growth evaluation by species are provided in the last section. -->

## Growth of surviving trees

Comparison of growth of trees (DBH >= 7.5) that survived in both observed and simulated data.

### Annual diameter increment

Overall predictive capacity to predict diameter increase (cm/yr):
```{r, echo = FALSE}
dfdbh %>% 
  evaluation_stats() %>% 
  kbl() %>%
  kable_styling()
```



Predictive capacity plots:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
dfdbh_bas <- dfdbh |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_dbh(dfdbh_bas, dfdbh_bas, points = TRUE)
```


Relationship between diameter increase and climatic variables (MAT, P/PET and available PAR):

```{r, echo = FALSE, fig.width=8, fig.height=12, warning=FALSE, message=FALSE}
dfdbh_bas <- dfdbh |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_cov_clim_dbh(dfdbh_bas, dfdbh_bas, points = TRUE)
```


### Annual height increment

Overall predictive capacity to predict height increase (cm/yr):
```{r, echo = FALSE, warning=FALSE, message=FALSE}
dfh %>% 
  evaluation_stats() %>%
  kbl() %>%
  kable_styling()
```


### Stand-level basal area increment

This comparison does not take into account changes in density. In other words, densities from the first inventory are used to calculate stand-level basal area of surviving trees. It is meant to evaluate the effect of diameter increment of surviving trees in terms of annual stand basal area increments (m2/ha/yr) for the period evaluated.

Predictive capacity table:
```{r, echo = FALSE, warning=FALSE}
dfbai_plot %>% evaluation_stats() %>% 
  kbl() %>%
  kable_styling()
```

Predictive capacity plots:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
dfbai_bas <- dfbai_plot |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_bai(dfbai_bas, dfbai_bas)
```

Relationship between basal area increase and climatic variables (MAT and P/PET):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfbai_bas, dfbai_bas, errorlim = c(-1,1), ylim = c(0,1))
```


Spatial error distribution:

```{r, echo = FALSE, fig.width=9, fig.height=4}
p1<-bai_error_map(dfbai_bas)+labs(title="Basic")
p2<-bai_error_map(dfbai_bas)+labs(title="Advanced")
plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
```



## Mortality

Annual reduction of basal area (m2/ha/yr) due to trees (DBH >= 7.5) that died during the evaluation period against model's mortality prediction.


Overall predictive capacity:

```{r, echo = FALSE, warning=FALSE}
dfbadead_plot %>% evaluation_stats() %>% 
  kbl() %>%
  kable_styling()
```

Predictive capacity plots:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message = FALSE}
dfdead_bas <- dfbadead_plot |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_bai(dfdead_bas, dfdead_bas,
                  quantity = "dead basal area", xylim = c(0,0.5), errorlim = c(-0.5,0.5))
```

Relationship between dead basal area and climatic variables (MAT and P/PET):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfdead_bas, dfdead_bas,
                   quantity = "Dead basal area", ylim = c(0,0.5), errorlim = c(-0.5,0.5))
```

Spatial distribution of errors:

```{r, echo = FALSE, fig.width=9, fig.height=4}
p1<-bai_error_map(dfdead_bas, upper = 0.5, lower = -0.5)+labs(title="Basic")
p2<-bai_error_map(dfdead_bas, upper = 0.5, lower = -0.5)+labs(title="Advanced")
plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
```



## Ingrowth

Annual increase of basal area (m2/ha/yr) due to ingrowth of trees with diameters between 7.5 cm and 12.5 cm during the evaluated period.


Predictive capacity:

```{r, echo = FALSE, warning=FALSE}
dfbarecr_plot %>% evaluation_stats()  %>% 
  kbl() %>%
  kable_styling()
```

Predictive capacity plots:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message = FALSE}
dfrecr_bas <- dfbarecr_plot |> filter(transpirationMode == "Granier",
                             period == "IFN24")
plot_scatter_bai(dfrecr_bas, dfrecr_bas,
                 quantity = "ingrowth basal area", xylim = c(0,0.15), errorlim = c(-0.15,0.15))
```

Relationship between ingrowth basal area and climatic variables (MAT and P/PET):

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
plot_cov_clim_bai(dfrecr_bas, dfrecr_bas,
                  quantity = "Ingrowth basal area", ylim = c(0,0.15), errorlim = c(-0.15,0.15))
```

Spatial distribution of errors:

```{r, echo = FALSE, fig.width=9, fig.height=4}
p1<-bai_error_map(dfrecr_bas, upper = 0.5, lower = -0.5)+labs(title="Basic")
p2<-bai_error_map(dfrecr_bas, upper = 0.5, lower = -0.5)+labs(title="Advanced")
plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
```



## Overall stand-level basal area change

This includes annual changes in basal area (m2/ha/yr) due to growth of surviving trees, mortality reductions and ingrowth (increase in trees with 7.5 < DBH < 12.5) derived from sapling growth. In the observed data, basal area changes include also the incorporation of trees into large diameter classes that results from the variable-radius sampling design.

Overall predictive capacity:

```{r, echo = FALSE, warning=FALSE}
dfbatot_plot %>% evaluation_stats() %>% 
  kbl() %>%
  kable_styling()
```


Predictive capacity plot:

```{r, echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message = FALSE}
dftot_bas <- dfbatot_plot |> filter(transpirationMode == "Granier",
                             period == "IFN24")
p1<-plot_scatter_bai(dftot_bas, dftot_bas,
                 quantity = "basal area change", xylim = c(-0.5,2), errorlim = c(-2,2))
p1
```

Relationship between overall basal area change and climatic variables (MAT and P/PET) for the **basic** sub-model:

```{r, echo = FALSE, fig.width=8, fig.height=8, warning=FALSE, message=FALSE}
p2<-plot_cov_clim_bai(dftot_bas, dftot_bas,
                  quantity = "Basal area change", ylim = c(-0.5,2), errorlim = c(-2,2))
p2
```

```{r, echo = FALSE}
p0<-cowplot::plot_grid(p1,p2, nrow=2, rel_heights = c(0.33,0.66))
ggsave(paste0(evalDIR,"/Plots/BAtot_eval.png"),p0, width = 8, height=10)
```

Spatial distribution of errors:

```{r, echo = FALSE, fig.width=9, fig.height=4, warning=FALSE, message=FALSE}
p1<-bai_error_map(dftot_bas)+labs(title="Basic sub-model")
p2<-bai_error_map(dftot_bas)+labs(title="Advanced sub-model")
pm<-plot_grid(p1+theme(legend.position = "none"),p2+theme(legend.position = "none"),
          get_legend(p1),nrow=1, rel_widths = c(1,1,0.25))
ggsave(paste0(evalDIR,"/Plots/BAtot_error_map.png"),pm, width = 13, height=6, bg="white")
pm
```


<!-- ## Detailed evaluation by species -->

<!-- In the following we provide detailed evaluation results for the most important species. -->


<!-- ```{r, echo=FALSE, results='asis', warning = FALSE, message=FALSE} -->
<!-- # sp_n <- dfbai_all_bas[dfbai_all_bas$Name!="0-All",] %>% group_by(Name) %>% -->
<!-- #     summarize(n = n()) -->
<!-- # sp100 = sp_n$Name[sp_n$n>params$minPlotRecordsDetailed] -->
<!-- sp100 = c("Abies alba", "Fagus sylvatica", "Pinus halepensis", "Pinus nigra", -->
<!--           "Pinus sylvestris", "Pinus uncinata", "Quercus faginea", "Quercus ilex", -->
<!--           "Quercus pubescens", "Quercus suber") -->
<!-- target_spp = target_spp_bas[names(target_spp_bas) %in% sp100] -->
<!-- for(i in 1:length(target_spp)) { -->

<!--   sp = target_spp[i] -->
<!--   name = names(target_spp)[i] -->
<!--   res <- knitr::knit_child('_RLEvaluation_child.Rmd', quiet = TRUE) -->
<!--   cat(res, sep = '\n') -->
<!-- } -->
<!-- ``` -->


