---
title: "Model evaluation in experimental plots"
author: "Miquel De Cáceres (CREAF), Víctor Granda (CREAF),  Rafael Poyatos (CREAF), Nicolas Martin-StPaul (INRAE)"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: TRUE
---



```{r setup, include=FALSE}
params = list(wd='/home/miquel/OneDrive/mcaceres_work/model_development/medfate_development/StandLevelEvaluation/',
              model='spwb',
              names = c("Hesse"),
              codes = c("FRAHES"),
              # names=c("Puéchabon", "Mitra","Hesse", "Yatir", "Rinconada", "Vallcebre (Barrol)", "Vallcebre (Sort)",
              #         "Fontainbleau-Barbeau", "Davos Seehornwald", "EucFACE", "Prades", "Wombat","Font-Blanche",
              #         "Can Balasc", "Armallones", "Ronda", "Morgan-Monroe"),
              # codes=c("FRAPUE", "MITRA","FRAHES", "ISRYAT", "ESPRIN", "QVALLCEBRE", "PVALLCEBRE",
              #         "FONTAINEBLEAU", "DAVOS", "EUCFACE","PRADES", "WOMBAT", "FONBLA",
              #         "CANBALASC", "ESPALTARM", "RONDA", "USAMORSF"),
              confs=c('granier','sperry', 'cochard'),
              save_plots = FALSE)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(knitr)
library(kableExtra)
library(medfate)
library(medfateutils)
library(ggplot2)

```

```{r message=FALSE, warning=FALSE}
this_path <- getwd()
knitr::opts_knit$set(root.dir = params$wd)
n_sites <- length(params$codes)
n_confs <- length(params$confs)
n_rows <- n_sites*n_confs
df <- data.frame(Site = rep(NA, n_rows), 
                 Mode = rep(NA, n_rows),
                 Path = rep(NA, n_rows))
df$SiteData = vector("list", n_rows)
df$InputObject = vector("list", n_rows)
df$Result = vector("list",n_rows)
cnt <- 0
df <- tibble::as_tibble(df)
df_general_sites <- data.frame(Plot = rep("", n_sites),
                               Country = rep("", n_sites),
                               Stand = rep("", n_sites),
                               SAPFLUXNET = rep("", n_sites),
                               FLUXNET = rep("", n_sites))

for(i in 1:n_sites) {
  site <- params$codes[i]
  for(j in 1:n_confs) {
    conf <- params$confs[j]
    cnt <- cnt+1
    df$Site[cnt] <- site
    df$Mode[cnt] <- conf
    df$Path[cnt] <- file.path(params$wd, 'Output', packageVersion('medfate')[[1]], params$model, conf, site)
    file_name_input <- file.path(df$Path[cnt],
                       paste0('simulation_input.rds'))
    file_name_output <- file.path(df$Path[cnt],
                       paste0('simulation_result.rds'))
    df$InputObject[[cnt]] <- readRDS(file_name_input)
    df$Result[[cnt]] <-   readRDS(file_name_output)
    site_list  <- medfatereports::load_list(site)
    df$SiteData[[cnt]] <- site_list
    df_general_sites$Plot[i] <- site_list$siteData$Value[1]
    col_country <- which(site_list$siteData$Attribute=="Country")
    df_general_sites$Country[i] <- site_list$siteData$Value[col_country]
    col_sapfluxnet_code <- which(site_list$siteData$Attribute=="SAPFLUXNET code")
    if(length(col_sapfluxnet_code)==1) df_general_sites$SAPFLUXNET[i] <- site_list$siteData$Value[col_sapfluxnet_code]
    col_fluxnet_code <- which(site_list$siteData$Attribute=="FLUXNET/ICOS code")
    if(length(col_fluxnet_code)==1) df_general_sites$FLUXNET[i] <- site_list$siteData$Value[col_fluxnet_code]
    col_stand <- which(site_list$siteData$Attribute=="Stand description")
    if(length(col_stand)==1) df_general_sites$Stand[i] <- site_list$siteData$Value[col_stand]
  }
}
names(df_general_sites)[5] <- "FLUXNET/ICOS"
knitr::opts_knit$set(root.dir = this_path)
```

# Introduction

This document presents **medfate** (**ver. `r packageVersion('medfate')[[1]]`**) model evaluation results at stand-level, using data from a set of **`r length(params$codes)` experimental forest plots**. The main source of observed data are SAPFLUXNET database ([Poyatos et al. 2021](https://essd.copernicus.org/articles/13/2607/2021/)) and FLUXNET 2015 dataset ([Pastorello et al. 2020](https://doi.org/10.1038/s41597-020-0534-3)).

Forest water balance simulations (i.e. function `spwb()`) have been conducted using the three transpiration modes (i.e. `Granier`, `Sperry` or `Cochard`). Only soil characteristics have been tuned, but calibration exercises have not been conducted. When available, however, leaf area to sapwood area ratios have been used. Thus, the evaluation is meant to be more or less representative of simulations with default species-level trait data.

## List of sites

```{r}
df_general_sites |> 
  kbl() |>
  kable_styling()
```

## Evaluation variables


| Variable                            | Level  | Observation source | Units |
|-------------------------------------|--------|--------------------|-------|
| Total evapotranspiration            | Stand  | FLUXNET / ICOS    | mm    |
| Latent heat turbulent flux          | Stand  | FLUXNET / ICOS    | MJ/m2 |
| Gross primary productivity          | Stand  | FLUXNET / ICOS    | gC/m2 |
| Soil moisture content (topsoil)     | Stand  | SAPFLUXNET / FLUXNET / ICOS         | % vol.|
| Transpiration per leaf area         | Plant  | SAPFLUXNET         | l/m2  |
| Predawn/midday leaf water potential | Plant  | SAPFLUXNET (addition)    | MPa   |


## Structure of site reports

The following contains as many sections as forest stands included in the evaluation. The following sub-sections are reported for each stand:

 1. **General information**: General information about the site, topography, soil and climate, as well as data sources used.
 2. **Model inputs**: Description of model inputs (vegetation, soil, custom species parameters and parameterization remarks). 
 3. **Evaluation results**: Evaluation results are presented for variables with available measurements.

```{r, echo=FALSE, results='asis', warning = FALSE, message=FALSE}
 for(i in 1:length(params$codes)) {
   site_code <- params$codes[i]
   site_name <- params$names[i]
   res <- knitr::knit_child(file.path(this_path,'_Stand_Evaluation_child.Rmd'), quiet = TRUE)
   cat(res, sep = '\n')
 }
```


