---
title: "Model evaluation in experimental plots"
author: "Miquel De Cáceres (CREAF), Víctor Granda (CREAF),  Rafael Poyatos (CREAF), Teresa Gimeno (CREAF), Nicolas Martin-StPaul (INRAE), Antoine Cabon (WSL), Vincenzo Saponaro (U. Tuscia)"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: TRUE
---



```{r setup, include=FALSE}
params = list(wd='/home/miquel/OneDrive/mcaceres_work/model_development/medfate_development/StandLevelEvaluation/',
              model='spwb',
              # names = c("Collelongo"),
              # codes = c("COLLELONGO"),
              names=c("Wombat","Euc-FACE",
                      "Puéchabon", "Hesse","Fontainebleau-Barbeau",  "Font-Blanche",
                      "Collelongo",
                      "Yatir",
                      "Mitra",
                      "Rinconada", "Vallcebre (Barrol)", "Vallcebre (Sort)", "Prades", "Can Balasc", "Alto-Tajo Armallones", "Ronda",
                      "Davos Seehornwald",
                      "Morgan-Monroe", "Sevilleta"),
              codes=c("WOMBAT", "EUCFACE",
                      "FRAPUE", "FRAHES", "FONTAINEBLEAU","FONBLA",
                      "COLLELONGO",
                      "ISRYAT",
                      "MITRA",
                      "ESPRIN", "QVALLCEBRE", "PVALLCEBRE", "PRADES", "CANBALASC", "ESPALTARM", "RONDA",
                      "DAVOS",
                      "USAMORSF", "SEVILLETA"),
              confs=c('granier','sperry', 'sureau'),
              save_plots = FALSE)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(knitr)
library(kableExtra)
library(medfate)
library(medfateutils)
library(ggplot2)

```

```{r message=FALSE, warning=FALSE}
this_path <- getwd()
knitr::opts_knit$set(root.dir = params$wd)
n_sites <- length(params$codes)
n_confs <- length(params$confs)
n_rows <- n_sites*n_confs
df_all <- data.frame(Site = rep(NA, n_rows), 
                 Mode = rep(NA, n_rows),
                 Path = rep(NA, n_rows))
df_all$SiteData = vector("list", n_rows)
df_all$InputObject = vector("list", n_rows)
df_all$Result = vector("list",n_rows)
cnt <- 0
df_all <- tibble::as_tibble(df_all)
df_general_sites <- data.frame(Country = rep("", n_sites),
                               Plot = rep("", n_sites),
                               Stand = rep("", n_sites),
                               SAPFLUXNET = rep("", n_sites),
                               FLUXNET = rep("", n_sites))

for(i in 1:n_sites) {
  site <- params$codes[i]
  for(j in 1:n_confs) {
    conf <- params$confs[j]
    cnt <- cnt+1
    df_all$Site[cnt] <- site
    df_all$Mode[cnt] <- conf
    df_all$Path[cnt] <- file.path(params$wd, 'Output', packageVersion('medfate')[[1]], params$model, conf, site)
    file_name_input <- file.path(df_all$Path[cnt],
                       paste0('simulation_input.rds'))
    file_name_output <- file.path(df_all$Path[cnt],
                       paste0('simulation_result.rds'))
    df_all$InputObject[[cnt]] <- readRDS(file_name_input)
    df_all$Result[[cnt]] <-   readRDS(file_name_output)
    site_list  <- medfatereports::load_list(site)
    df_all$SiteData[[cnt]] <- site_list
    df_general_sites$Plot[i] <- site_list$siteData$Value[1]
    col_country <- which(site_list$siteData$Attribute=="Country")
    df_general_sites$Country[i] <- site_list$siteData$Value[col_country]
    col_sapfluxnet_code <- which(site_list$siteData$Attribute=="SAPFLUXNET code")
    if(length(col_sapfluxnet_code)==1) df_general_sites$SAPFLUXNET[i] <- site_list$siteData$Value[col_sapfluxnet_code]
    col_fluxnet_code <- which(site_list$siteData$Attribute=="FLUXNET/ICOS code")
    if(length(col_fluxnet_code)==1) df_general_sites$FLUXNET[i] <- site_list$siteData$Value[col_fluxnet_code]
    col_stand <- which(site_list$siteData$Attribute=="Forest stand")
    if(length(col_stand)==1) df_general_sites$Stand[i] <- site_list$siteData$Value[col_stand]
  }
}
names(df_general_sites)[5] <- "FLUXNET/ICOS"
knitr::opts_knit$set(root.dir = this_path)
```

# Introduction

This document presents **medfate** (**ver. `r packageVersion('medfate')[[1]]`**) model evaluation results at stand-level, using data from a set of **`r length(params$codes)` experimental forest plots**. The main source of observed data are SAPFLUXNET database ([Poyatos et al. 2021](https://essd.copernicus.org/articles/13/2607/2021/)) and FLUXNET 2015 dataset ([Pastorello et al. 2020](https://doi.org/10.1038/s41597-020-0534-3)).

## List of sites

The table below lists the experimental forest plots used in the report and the data sources available.

```{r}
df_general_sites |> 
  kbl() |>
  kable_styling()
```

## Parametrization and simulations

Forest water balance simulations (i.e. function `spwb()`) have been conducted using the three transpiration modes (i.e. `Granier`, `Sperry` or `Sureau`). 

The set of control parameters modified from defaults in simulations are the following:

```{r}
control_granier <- yaml::yaml.load_file(system.file("yaml_configurations", "granier.yaml",
                                          package = "medfatereports"))
control_sperry <- yaml::yaml.load_file(system.file("yaml_configurations", "sperry.yaml",
                                          package = "medfatereports"))
control_sureau <- yaml::yaml.load_file(system.file("yaml_configurations", "sureau.yaml",
                                          package = "medfatereports"))
dplyr::bind_rows(as.data.frame(control_granier),
                 as.data.frame(control_sperry),
                 as.data.frame(control_sureau))|> 
  kbl() |>
  kable_styling()
```

Soil characteristics have been tuned to modulate total available water and fit observed saturation and residual moisture values, but calibration exercises have not been conducted. When available, however, local leaf area to sapwood area ratios have been used. Thus, the evaluation exercise is meant to be more or less representative of simulations with default species-level trait data.


## Evaluation variables

The table below lists the set of predicted variables that are evaluated and the data sources used:

| Variable                            | Level  | Observation source | Units |
|-------------------------------------|--------|--------------------|-------|
| Sensible heat turbulent flux        | Stand  | FLUXNET / ICOS    | MJ/m2 |
| Latent heat turbulent flux          | Stand  | FLUXNET / ICOS    | MJ/m2 |
| Gross primary productivity          | Stand  | FLUXNET / ICOS    | gC/m2 |
| Soil moisture content (topsoil)     | Stand  | SAPFLUXNET / FLUXNET / ICOS         | % vol.|
| Transpiration per leaf area         | Plant  | SAPFLUXNET         | l/m2  |
| Predawn/midday leaf water potential | Plant  | SAPFLUXNET (addition)    | MPa   |


## Structure of site reports

The following contains as many sections as forest stands included in the evaluation. The following sub-sections are reported for each stand:

 1. **General information**: General information about the site, topography, soil and climate, as well as data sources used.
 2. **Model inputs**: Description of model inputs (vegetation, soil, custom species parameters and parameterization remarks). 
 3. **Climate**: Graphical description of climate inputs and predicted soil/canopy temperatures (under Sperry).
 4. **Evaluation results**: Evaluation results are presented for variables with available measurements.

```{r, echo=FALSE, results='asis', warning = FALSE, message=FALSE}
 for(i_code in 1:length(params$codes)) {
   site_code <- params$codes[i_code]
   site_name <- params$names[i_code]
   df_site <- df_all[df_all$Site==site_code,, drop = FALSE]
   res <- knitr::knit_child(file.path(this_path,'_Stand_Evaluation_child.Rmd'), quiet = TRUE)
   cat(res, sep = '\n')
 }
```


