---
title: "Initial species parameter table for Spain"
author: "Miquel De Cáceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: TRUE
bibliography: ../medfate.bib
params:
  spparams_rebuild: FALSE
  IFN_rebuild: FALSE
---

## Introduction

This vignette describes the procedures used to obtain the initial species parameter data table that serves to obtain the data frame `SpParamsMED` included in package **medfate**. The taxon entities chosen are intended to cover most of the woody plant taxa in Spain, as they are reported in the Spanish National Forest Inventory surveys (*Inventario Forestal Nacional* or IFN). In addition, a list of `forest` objects is also created for each of three IFN surveys, where species codification follows the parameter data table.

>**IMPORTANT**: The vignette is not self-contained, in the sense that it cannot be reproduced without access to data sets that are not included. Nevertheless, it is intended to serve as example of species parameterization for other regions.

### Required packages`
Installed version of **medfate** should be **ver 2.9** or higher. 
```{r, eval = FALSE}
install.packages("medfate")
```

If the version at CRAN is older, the package should be installed from GitHub via `remotes::install_github()`:
```{r, eval = FALSE}
remotes::install_github("emf-creaf/medfate")
```

Package **medfateutils** is required for the species parameterization process, as it includes several functions that facilitate parsing data sources into parameter values. Package **medfateutils** is installed only from GitHub:

```{r, eval = FALSE}
remotes::install_github("emf-creaf/medfateutils")
```

Package **medfuels** is also needed because it includes allometric coefficients for Mediterranean shrub species. As before, package **medfuels** is installed only from GitHub:

```{r, eval = FALSE}
remotes::install_github("spif-ctfc/medfuels")
```

Finally, we need **IFNread** is also needed in this vignette to perform some operations on IFN data. As before, package **IFNread** is installed only from GitHub:

```{r, eval = FALSE}
remotes::install_github("emf-creaf/IFNread")
```

Once we have **medfate**, **medfateutils**,  **medfuels** and **IFNread**, we load them and other common packages that we will employ in this vignette:
```{r message=FALSE}
library(medfate)
library(medfateutils)
library(medfuels)
# library(IFNread)
library(tidyverse)
library(openxlsx)
```

### Target parameters

The set of species parameters needed to run the models included in **medfate** are described in a dataframe called `SpParamsDefinition`: 


```{r}
data("SpParamsDefinition")
SpParamsDefinition$ParameterName
```
It is obvious that such a sheer number of parameters cannot be searched for manually, even for a small number of target species. Hence, it is important to draw values from allometric and plant trait data bases. Not all model parameters have their counterpart in publicly-available databases, and even when using data bases it is unlikely that one will find appropriate parameter values for all species. To help with this situation, **medfate** has *inbuilt imputation procedures* to estimate missing values for many parameters before starting model simulations. However, imputed parameter estimates will normally be worse than estimates coming from data bases. Hence, we should put a strong effort in finding source parameter data before relying on inbuilt imputation.


### Steps to build a species parameter table

The following sections describe different steps that we used to obtain the species parameter table, which can be grouped into five main tasks:

 1. Initialize the parameter table with target taxonomic entities.
 2. Draw species parameters from forest inventory (IFN in our case) data.
 3. Populate tree and shrub allometric coefficients from suitable allometric databases.
 4. Populate plant functional traits from available data bases
 
 
### Source data paths

Before starting, we define file paths (in my computer) to access source data for parameter filling. Source data includes forest inventory (i.e. IFN) data as well as plant trait and allometric data bases:
```{r paths}
IFN_path = "/home/miquel/OneDrive/EMF_datasets/ForestInventories/IFN/"
IFN_out_path = "/home/miquel/OneDrive/mcaceres_work/model_initialisation/MEDFATE_Initialisation/IFN/"
DB_path = "/home/miquel/OneDrive/mcaceres_work/model_development/medfate_development/MedfateSpeciesParametrization/"
```

## Initialize species parameter table for IFN

The first task is to define the target entities that will be considered functionally distinct in **medfate** simulations. There is no *a priori* need to limit such entities to biological taxa, nor to particular taxonomic ranks. In our case, we take the species mapping defined in package **medfateutils** for Spanish IFN:

```{r}
data(IFN_species_mapping)
head(IFN_species_mapping)
```


Once we have IFN species names and codes, as well as the definition of the IFN biological entities to merge, we can call function `initSpParams()` from package **medfateutils**. This function also takes the data frame `SpParamsDefinition` as argument, from which it takes the set of parameter names to be filled:
```{r initialization, eval = params$spparams_rebuild}
SpParams <-initSpParams(sp_names = IFN_species_mapping$Name,
                        SpParamsDefinition,
                        fill_taxonomic = TRUE, verbose = FALSE)
```
Option `fill_taxonomic` tells function `initSpParams()` to use package **taxize** to find taxonomic classification of the species/genus entities. This provides very useful information for parameterization, but makes the process of initialization a bit slower.
```{r include = FALSE, echo = FALSE}
if(params$spparams_rebuild) {
  saveRDS(SpParams, "SpParamsSPAIN_init.rds")
} else {
  SpParams = readRDS("SpParamsSPAIN_init.rds")
}
```
The result of calling `initSpParams()` is a data frame with `nrow(SpParams)` rows corresponding to target entities and columns corresponding to the parameters defined in `SpParamsDefinition`. Most parameters are missing, obviously, but `initSpParams()` already filled taxonomic data:
```{r}
head(SpParams[,1:7])
```

<!-- ## Parameters drawn from IFN data -->

<!-- Some plant size parameters, such as maximum height or minimum/maximum values of the height-to-diameter ratio, can be derived from forest inventory data. In addition, it is important that medfate's parameter `GrowthForm` (either "Tree", "Tree/Shrub" or "Shrub") matches the treatment given in forest inventory surveys. Diameters are normally measured in "Tree" entities, whereas shrubs are often measured using cover indices, and some species may be recorded in either way. Thus, the `GrowthForm` parameter should also be derived from forest inventory. -->

<!-- ### Load IFN plot data -->

<!-- We first load IFN tree and shrub data corresponding to the three surveys (tree data is in two different tables in the case of IFN2):  -->
<!-- ```{r load_ifn} -->
<!-- shrubDataIFN2 <-readRDS(paste0(IFN_path,"Products/IFN2/Rdata/shrubDataIFN2_Spain.rds")) -->
<!-- shrubDataIFN2$ID<-as.character(shrubDataIFN2$ID) -->
<!-- shrubDataIFN3 <-readRDS(paste0(IFN_path,"Products/IFN3/Rdata/shrubDataIFN3_Spain.rds")) -->
<!-- shrubDataIFN4 <-readRDS(paste0(IFN_path,"Products/IFN4/Rdata/shrubDataIFN4_Spain.rds")) -->

<!-- piesMayoresDataIFN2<-readRDS(paste0(IFN_path,"Products/IFN2/Rdata/piesMayoresDataIFN2_Spain.rds")) -->
<!-- piesMenoresDataIFN2<-readRDS(paste0(IFN_path,"Products/IFN2/Rdata/piesMenoresDataIFN2_Spain.rds")) -->
<!-- piesMenoresDataIFN2$ID = as.character(piesMenoresDataIFN2$ID) -->
<!-- piesMenoresDataIFN2$Species = as.character(piesMenoresDataIFN2$Species) -->
<!-- treeDataIFN2 = rbind(piesMayoresDataIFN2, piesMenoresDataIFN2) # Add pies menores -->

<!-- treeDataIFN3<-readRDS(paste0(IFN_path,"Products/IFN3/Rdata/treeDataIFN3_Spain.rds")) -->
<!-- treeDataIFN3 = treeDataIFN3[!(treeDataIFN3$OIF3 %in% c("888","999")),] # remove dead trees -->
<!-- regDataIFN3 = readRDS(paste0(IFN_path,"Products/IFN3/Rdata/regDataIFN3_Spain.rds")) -->
<!-- regTreeDataIFN3 = IFNread::extractRegTreeData(regDataIFN3, subsetVars=T) -->
<!-- treeDataIFN3 = rbind(treeDataIFN3[,c("ID","Species","DBH","H","N")], -->
<!--                      regTreeDataIFN3[,c("ID","Species","DBH","H","N")]) -->
<!-- treeDataIFN4<-readRDS(paste0(IFN_path,"Products/IFN4/Rdata/treeDataIFN4_Spain.rds")) -->
<!-- treeDataIFN4 = treeDataIFN4[!(treeDataIFN4$OIF4 %in% c("888","999")),] # remove dead trees -->
<!-- regDataIFN4 = readRDS(paste0(IFN_path,"Products/IFN4/Rdata/regDataIFN4_Spain.rds")) -->
<!-- regTreeDataIFN4 = IFNread::extractRegTreeData(regDataIFN4, subsetVars=T) -->
<!-- treeDataIFN4 = rbind(treeDataIFN4[,c("ID","Species","DBH","H","N")], -->
<!--                      regTreeDataIFN4[,c("ID","Species","DBH","H","N")]) -->

<!-- ``` -->



<!-- ### Merge IFN data -->
<!-- We merge the data sets into two tables, one for shrubs and the other for trees, taking only columns `ID` (plot identification), `Species` (code), `H` (height) and `DBH` (for trees). -->
<!-- ```{r merge_ifn} -->
<!-- shrubDataIFN <-bind_rows(shrubDataIFN2[,c("ID","Species", "H")], -->
<!--                          shrubDataIFN3[,c("ID","Species", "H")], -->
<!--                          shrubDataIFN4[,c("ID","Species", "H")]) -->

<!-- treeDataIFN <-bind_rows(treeDataIFN2[,c("ID","Species", "H", "DBH")], -->
<!--                         treeDataIFN3[,c("ID","Species", "H", "DBH")], -->
<!--                         treeDataIFN4[,c("ID","Species", "H", "DBH")]) -->
<!-- ``` -->

<!-- ### Growth form -->

<!-- To tell **medfate** which target entities are shrubs, trees (or both, since some species occur in both tree and shrub tables), we can simply take the species code vectors: -->
<!-- ```{r growth_form} -->
<!-- shrub_codes <- as.numeric(shrubDataIFN$Species) -->
<!-- tree_codes <- as.numeric(treeDataIFN$Species) -->
<!-- ``` -->

<!-- Before populating parameter `GrowthForm`, we can check how many species appear both as trees and shrubs: -->
<!-- ```{r} -->
<!-- sum(unique(shrub_codes) %in% unique(tree_codes)) -->
<!-- ``` -->
<!-- We can populate column `GrowthForm` of the species parameter table by calling function `populateGrowthForm()` of package **medfateutils**: -->
<!-- ```{r} -->
<!-- SpParams = populateGrowthForm(SpParams,  -->
<!--                               tree_codes, shrub_codes, -->
<!--                               fill_fromGenus = TRUE) -->

<!-- ``` -->
<!-- Some codes (particularly, those excluded above) may not be recognized according to the target entity definition of `SpParams` and are ignored. The option `fill_fromGenus = TRUE` was used to tell function `populateGrowthForm()` to impute growth form of species with missing value from the growth form of other species of the same genus, if homogeneous. Some species may still have missing growth form after the imputation process. We identify them: -->
<!-- ```{r} -->
<!-- missingGrowthFormSpecies = SpParams[is.na(SpParams$GrowthForm), c("Name")] -->
<!-- print(missingGrowthFormSpecies) -->
<!-- ``` -->
<!-- and manually set the growth form for them: -->
<!-- ```{r} -->
<!-- missingTrees<-c( "Dracaena drago", "Pleiomeris canariensis", "Quercus lusitanica", -->
<!--                  "Sideroxylon mirmulans")  -->
<!-- SpParams$GrowthForm[SpParams$Name %in% missingTrees] = "Tree" -->
<!-- missingShrubs<-c( "Dittrichia viscosa", "Periploca laevigata", "Withania frutescens")  -->
<!-- SpParams$GrowthForm[SpParams$Name %in% missingShrubs] = "Shrub" -->

<!-- ``` -->
<!-- We can inspect how many species were deemed "Tree", "Shrub" or "Tree/Shrub". -->
<!-- ```{r} -->
<!-- table(SpParams$GrowthForm, useNA = "ifany") -->
<!-- ``` -->


<!-- ### Maximum/median height -->

<!-- Maximum and median heights for each target entity can be obtained from observed height distributions. Thus, we define species identity and height (in cm) vectors from tree/shrub data: -->
<!-- ```{r maximum_med_height} -->
<!-- species_codes <- c(treeDataIFN$Species, shrubDataIFN$Species) -->
<!-- height_values <- c(treeDataIFN$H, shrubDataIFN$H)*100 # m to cm -->
<!-- ``` -->
<!-- and then call function `populateHeightParams()` to populate parameters `Hmax` and `Hmed`, corresponding to maximum and median heights. -->
<!-- ```{r} -->
<!-- SpParams <- populateHeightParams(SpParams, -->
<!--                                  species_codes, height_values) -->
<!-- ``` -->
<!-- By default, maximum height is defined as the 99\% quantile of the height distribution, whereas median height corresponds to the 50\% quantile.  -->

<!-- Parameters `Hmax` and `Hmed` are strict, meaning that medfate does not have inbuilt imputation procedures for them, so we must complete their columns in the parameter table. We can inspect the missing values for each growth form:  -->
<!-- ```{r} -->
<!-- table(is.na(SpParams$Hmax), SpParams$GrowthForm) -->
<!-- table(is.na(SpParams$Hmed), SpParams$GrowthForm) -->
<!-- ``` -->
<!-- and then use the species with known values for each growth from to impute species averages: -->
<!-- ```{r} -->
<!-- SpParams$Hmax[is.na(SpParams$Hmax) & SpParams$GrowthForm != "Shrub"] = round(mean(SpParams$Hmax[SpParams$GrowthForm != "Shrub"], na.rm=TRUE)) -->
<!-- SpParams$Hmax[is.na(SpParams$Hmax) & SpParams$GrowthForm == "Shrub"] = round(mean(SpParams$Hmax[SpParams$GrowthForm == "Shrub"], na.rm=TRUE)) -->

<!-- SpParams$Hmed[is.na(SpParams$Hmed) & SpParams$GrowthForm != "Shrub"] = round(mean(SpParams$Hmed[SpParams$GrowthForm != "Shrub"], na.rm=TRUE)) -->
<!-- SpParams$Hmed[is.na(SpParams$Hmed) & SpParams$GrowthForm == "Shrub"] = round(mean(SpParams$Hmed[SpParams$GrowthForm == "Shrub"], na.rm=TRUE)) -->
<!-- ``` -->

<!-- The final distribution of `Hmax` (cm) by growth form is: -->
<!-- ```{r} -->
<!-- tapply(SpParams$Hmax, SpParams$GrowthForm, summary) -->
<!-- ``` -->
<!-- and that of `Hmed` (cm) by growth form is: -->
<!-- ```{r} -->
<!-- tapply(SpParams$Hmed, SpParams$GrowthForm, summary) -->
<!-- ``` -->


<!-- ### Diameter/height ratio -->

<!-- Minimum and maximum diameter/height ratios are used in **medfate** to translate diameter growth into height growth. We first assemble three data vectors needed to populate these parameters (heights and diameters in cm): -->
<!-- ```{r HD_ratio} -->
<!-- tree_species_codes <- treeDataIFN$Species -->
<!-- tree_height_values <- treeDataIFN$H*100 # m to cm -->
<!-- tree_diameter_values <- treeDataIFN$DBH -->
<!-- ``` -->
<!-- and then call function `populateTreeDiameterHeightParams()` from package **medfateutils**: -->
<!-- ```{r} -->
<!-- sel <- (tree_diameter_values >1 & tree_diameter_values<7.5  -->
<!--         & tree_height_values >130 & tree_height_values < 2000) -->
<!-- sel[is.na(sel)] = FALSE -->
<!-- SpParams <- populateTreeDiameterHeightParams(SpParams, -->
<!--                                              tree_species_codes[sel],  -->
<!--                                              tree_height_values[sel], -->
<!--                                              tree_diameter_values[sel],  -->
<!--                                              quantile_fHDmax = 0.99) -->
<!-- ``` -->
<!-- This function fills parameters `fHDmin` and `fHDmax`, corresponding to the minimum and maximum values of the height-to-diameter ratio. The 5\% and 95\% quantile values are used for these parameters, respectively, by default.  -->

<!-- Missing values may occur after imputation:   -->
<!-- ```{r} -->
<!-- table(is.na(SpParams$fHDmin), SpParams$GrowthForm) -->
<!-- table(is.na(SpParams$fHDmax), SpParams$GrowthForm) -->
<!-- ``` -->
<!-- Since `fHDmin` and `fHDmax` are strict parameters, we use shrub and tree averages to fill the remaining cases: -->
<!-- ```{r} -->
<!-- SpParams$fHDmax[is.na(SpParams$fHDmax) & SpParams$GrowthForm != "Shrub"] = round(mean(SpParams$fHDmax, na.rm=TRUE)) -->
<!-- SpParams$fHDmin[is.na(SpParams$fHDmin) & SpParams$GrowthForm != "Shrub"] = round(mean(SpParams$fHDmin, na.rm=TRUE)) -->
<!-- ``` -->

<!-- The final distributions of `fHDmax` and `fHDmin` are: -->
<!-- ```{r} -->
<!-- summary(SpParams[SpParams$GrowthForm != "Shrub" ,c("fHDmax", "fHDmin")]) -->
<!-- ``` -->


## Allometric coefficients

Allometric equations are used in **medfate** to provide initial estimates of foliar biomass and leaf area of plant cohorts, which are key variables in simulations of forest functioning.

### Shrub allometric equations

Shrub allometric equations in **medfate** are of the form $y = a\cdot x ^b$, with $a$ and $b$ being allometric coefficients. We will draw shrub allometric coefficient from data included in package **medfuels**. First, we begin by loading from this package a data set that describes the shrub type for a number of Mediterranean species. Shrub types in **medfuels** are a combination of Raunkiaer life form and resprouting strategy. Whenever species level coefficients will be missing, we will impute allometric coefficients according to the shrub type specified in **medfuels**.
```{r shrub_allometries}
data("species_groups")
head(species_groups)
```

Here we manually correct the name and authority of some synonyms:
```{r}

species_groups[species_groups$name=="Rosmarinus officinalis","Genus"]<- "Salvia"
species_groups[species_groups$name=="Rosmarinus officinalis","Species"]<- "rosmarinus"
species_groups[species_groups$name=="Rosmarinus officinalis","author"]<- "Schleid."
species_groups[species_groups$name=="Rosmarinus officinalis","name"]<- "Salvia rosmarinus"

species_groups[species_groups$name=="Lithospermum fruticosum","Genus"]<- "Lithodora"
species_groups[species_groups$name=="Lithospermum fruticosum","Species"]<- "fruticosa"
species_groups[species_groups$name=="Lithospermum fruticosum","author"]<- "(L.) Griseb."
species_groups[species_groups$name=="Lithospermum fruticosum","name"]<- "Lithodora fruticosa"
```

We will first populate allometric coefficients $a$ and $b$ for the relationship between height and projected area. Allometric coefficients are in the following tables for species:
```{r}
data("sp_params_area")
head(sp_params_area)
```
and for shrub types:
```{r}
data("group_params_area")
head(group_params_area)
```
We are now ready to use function `populateShrubAllometriesFromMedfuels()` from package **medfateutils**:
```{r}

SpParams <- populateShrubAllometriesFromMedfuels(SpParams, 
                                             list("a_ash" = "a","b_ash" = "b"),
                                             sp_params_area, 
                                             group_params_area, 
                                             species_groups)
```
We can do the same for the relationship between shrub volume and biomass of fine fuels:
```{r}
data("sp_params_fine")
data("group_params_fine")
SpParams <- populateShrubAllometriesFromMedfuels(SpParams, 
                                             list("a_bsh" = "a","b_bsh" = "b"),
                                             sp_params_fine, 
                                             group_params_fine, 
                                             species_groups)
```
and for the relationship between shrub volume and total biomass:
```{r}
data("sp_params_total")
data("group_params_total")
SpParams <- populateShrubAllometriesFromMedfuels(SpParams, 
                                             list("a_btsh" = "a","b_btsh" = "b"),
                                             sp_params_total, 
                                             group_params_total, 
                                             species_groups)
```

### Shrub crown ratio and fraction of dead material
Although values not included in **medfuels**, we can use the same strategy to populate the fraction of dead material (`pDead` or $P_{dead}$) and the crown ratio (`cr`) for shrubs from a trait database. First we load the data frame:
```{r shrub_cr_pdead}
shrub_pdead_cr <- read.table(paste0(DB_path,"AllometryDatabases/ShrubAllometries/Shrub_Pdead_CR.txt"), 
                             sep="\t", dec=",", header=T) %>% 
  separate(Name, c("Genus", "Species"), sep=" ") %>% 
  unite("Name" , c("Genus", "Species"), sep=" ", na.rm=TRUE)
shrub_data<-data.frame(cr = shrub_pdead_cr$CR.mean,
                       pDead = shrub_pdead_cr$Pdead.mean,
                       row.names = shrub_pdead_cr$Name)
shrub_data
```
We then split the data frame into a table for species and a table for shrub types, and call function `populateShrubAllometriesFromMedfuels()`:
```{r}
sp_params_allom = shrub_data[1:27,, drop=FALSE]
group_params_allom = shrub_data[28:33,, drop=FALSE]
SpParams <- populateShrubAllometriesFromMedfuels(SpParams, 
                        list("cr" = "cr", "pDead" = "pDead"),
                        sp_params_allom, 
                        group_params_allom, 
                        species_groups)
```


### Tree allometries

Tree allometric models for foliar biomass were calibrated from field data obtained in IEFC, a forest inventory conducted in Catalonia (*Inventari Ecològic i Forestal de Catalunya*) more or less during the same years as the Spanish IFN3. We begin by loading the table containing regression coefficients:
```{r tree_allom}
fb_models <- openxlsx::read.xlsx(paste0(DB_path, "AllometryDatabases/TreeAllometries/TreeAllometries.xlsx"), sheet= "Tree_FB_models", rowNames = TRUE)
head(fb_models)
```
And we simply use function `populateTreeAllometries()` from package **medfateutils** to copy the coefficients onto `SpParams`:
```{r}
SpParams <- populateTreeAllometries(SpParams, fb_models, "foliarbiomass")
```
Note that some assignments for species not present in the source table were done using allometric coefficients that had been defined in the source table at the genus level. 

Crown width and crown ratio coefficients were obtained from the literature and are loaded here:
```{r}
cr_models <- openxlsx::read.xlsx(paste0(DB_path,"AllometryDatabases/TreeAllometries/TreeAllometries.xlsx"), 
                                 sheet= "Tree_CR_models", rowNames = TRUE)
head(cr_models)
```

```{r}
cw_models <- openxlsx::read.xlsx(paste0(DB_path,"AllometryDatabases/TreeAllometries/TreeAllometries.xlsx"), 
                                 sheet= "Tree_CW_models", rowNames = TRUE)
head(cw_models)
```
We use the same function to populate the corresponding columns in `SpParams`:
```{r}
SpParams <- populateTreeAllometries(SpParams, cr_models, "crownratio")
SpParams <- populateTreeAllometries(SpParams, cw_models, "crownwidth")
```


## Parameters drawn from plant trait databases

Although we used specific functions to map most of the parameters so far, package **medfateutils** offers a generic function, called `populateTraits()`, to help mapping plant traits from data bases into model parameters. In this section we will extensively use this function to fill a large number of parameters from TRY [@Kattge2011] and other plant trait data bases available to us.

### Trait mapping from TRY

The TRY database [@Kattge2011] is a primary source for plant traits. We created a list of TRY trait definitions that could be mapped to **medfate** parameters. The following table shows the mapping:

```{r try_correspondence}
TRY_mapping = openxlsx::read.xlsx(file.path(DB_path, "TraitDatabases/TRY/TraitCorrespondence.xlsx"))
knitr::kable(TRY_mapping[!is.na(TRY_mapping$Used) & TRY_mapping$Used=="YES",c(1,3,4)], 
             row.names = FALSE)
```

We then queried TRY for public data regarding the list of TraitIDs and split the result into one database for each trait, using the TRY TraitID as file name. Data for these plant traits will be progressively loaded as required in the following subsections.

### Additional trait databases

Additional plant trait databases can be used for parameter estimation. We begin by loading several trait data bases, that drawn from literature publications or assembled by ourselves. One of them is HydraTRY [@Mencuccini2019;@SanchezMartinez2020], which contains hydraulic plant traits:
```{r other_dbs}
fn_hydratry <- file.path(DB_path,"TraitDatabases/HydraTRY/HydraTRY.xlsx")
hydratry <- openxlsx::read.xlsx(fn_hydratry, rowNames = FALSE) 
```
Plant hydraulic traits can also be found in @Martin-StPaul2017:
```{r}
fn_martin_stpaul <- file.path(DB_path,"TraitDatabases/MartinStPaul_et_al_2017/DataBase.xlsx")
martin_stpaul <- read.xlsx(fn_martin_stpaul, sheet="ALL")
martin_stpaul<-martin_stpaul[,c("Species", "P50", "P12", "Ptlp")]
```
and @Choat2012:
```{r}
fn_choat <-file.path(DB_path,"TraitDatabases/Choat_et_al_2012/nature11688-s2.xls")
choat <- readxl::read_excel(fn_choat)
choat <-choat[-1, c("Species", "P50", "P88")] # Remove unit row
```

@Bartlett2012 provides values for leaf pressure-volume curves:
```{r}
fn_bartlett <- file.path(DB_path, "TraitDatabases/Bartlett_et_al_2012/Bartlett_2012_ELE_data.xlsx")
bartlett <- openxlsx::read.xlsx(fn_bartlett, rowNames = FALSE) 
```

@Duursma2018 is a good source of minimum stomatal conductance:
```{r}
fn_duursma <- file.path(DB_path, "TraitDatabases/Duursma_et_al_2018/gmindatabase.csv")
duursma <- read.csv2(fn_duursma, sep=",")
```

whereas @Hoshika2018 provides maximum stomatal conductance values: 
```{r}
fn_hoshika <- file.path(DB_path, "TraitDatabases/Hoshika_et_al_2018/Hoshika_et_al_2018.xlsx")
hoshika <- openxlsx::read.xlsx(fn_hoshika) 
hoshika$Species <- str_replace(string = hoshika$Species,pattern = "_", replacement = " ")
```

BROT2 database [@Tavsanoglu2018] is particularly useful for Mediterranean plants, containing many anatomical, phenological and morphological traits:
```{r}
fn_brot2 <- file.path(DB_path, "TraitDatabases/BROT2/BROT2_dat.csv")
brot2 <- read.csv2(fn_brot2)
```

Another source of data is the Global wood density database [@Zanne2009]:
```{r}
fn_gwdd <- file.path(DB_path, "TraitDatabases/GlobalWoodDensityDatabase/GlobalWoodDensityDatabase.xls")
gwdd <- readxl::read_excel(fn_gwdd, sheet=2)
```

and @Yebra2019 provides estimates of fine fuel moisture content for a large number of species worldwide:
```{r}
fn_yebra <- file.path(DB_path, "TraitDatabases/LFMC_Yebra_et_al_2019/Globe-LFMC-v2.xlsx")
fmc_yebra <- openxlsx::read.xlsx(fn_yebra, sheet=2)
```

Finally, @Morris2016 provides fractions of parenchymatic tissue within sapwood:
```{r}
fn_morris <- file.path(DB_path, 
                       "TraitDatabases/Morris_et_al_2016/nph13737-sup-0002-tables1.xlsx")
morris <- openxlsx::read.xlsx(fn_morris, sheet=2)
```


We complemented the formed data bases with our own compilation of plant traits, including leaf phenology:
```{r}
fn_pheno <- file.path(DB_path, "TraitDatabases/LeafPhenology/LeafPhenology.xlsx")
budburst <- as.data.frame(readxl::read_excel(fn_pheno, sheet=1))
senescence <- as.data.frame(readxl::read_excel(fn_pheno, sheet=2))
```

flammability traits:
```{r}
fn_flam <- file.path(DB_path, "TraitDatabases/FuelAttributes/FuelAttributes.xlsx")
flam <- readxl::read_excel(fn_flam, sheet=1)
```
the ratio of leaf dry mass to fine fuel dry mass:
```{r}
fn_ffr <- file.path(DB_path, "TraitDatabases/FineFuelRatio/FineFuelRatio.xlsx")
ffr <- readxl::read_excel(fn_ffr, sheet=1)
```
minimum/maximum fuel moisture contents:
```{r}
fn_fmc <- file.path(DB_path, "TraitDatabases/FMC/FMC.xlsx")
fmc_field <- readxl::read_excel(fn_fmc, sheet=1)
fmc_bib <- readxl::read_excel(fn_fmc, sheet=2)
```
and construction costs:
```{r}
fn_lcc <- file.path(DB_path, "TraitDatabases/ConstructionCosts/LeafConstructionCosts.xlsx")
lcc_data <- readxl::read_excel(fn_lcc, sheet=1)
fn_scc <- file.path(DB_path, "TraitDatabases/ConstructionCosts/SapwoodConstructionCosts.xlsx")
scc_data <- readxl::read_excel(fn_scc, sheet=1)
fn_fcc <- file.path(DB_path, "TraitDatabases/ConstructionCosts/FineRootConstructionCosts.xlsx")
fcc_data <- readxl::read_excel(fn_fcc, sheet=1)
```

### Plant/leaf classification

Plant and leaf classification parameters (i.e. plant life form, leaf phenology, leaf size class, leaf form) are strict in **medfate**, meaning they need to be provided for all target entities.

#### Life form

A large number of life form records can be found in TRY. We load TRY life form data (TraitID = 343) first: 
```{r life_form}
TRY_life_form  <-readRDS(file.path(DB_path, "TraitDatabases/TRY/TRY_traits/TRY_343.rds"))
```
Since life form descriptions in TRY are not standardized, we standardize life form values into six categories:
```{r}
TRY_life_form$OrigValueStr[TRY_life_form$OrigValueStr %in% 
                             c("Ch", "CH", "Cha", "chamaephyte", "Chamaephytes")] = "Chamaephyte"
TRY_life_form$OrigValueStr[TRY_life_form$OrigValueStr %in% 
                             c("Cryptophytes", "cryptophyte",
                               "G", "geophyte", "Geophytes")] = "Cryptophyte"
TRY_life_form$OrigValueStr[TRY_life_form$OrigValueStr %in% 
                             c("Ep", "epiphyte","Epiphytes", "liana")] = "Epiphyte"
TRY_life_form$OrigValueStr[TRY_life_form$OrigValueStr %in% 
                             c("H", "hemicryptophyte", "Hemicryptophytes")] = "Hemicryptophyte"
TRY_life_form$OrigValueStr[TRY_life_form$OrigValueStr %in% 
                             c("hydrophyte", "Hydrophytes")] = "Hydrophyte"
TRY_life_form$OrigValueStr[TRY_life_form$OrigValueStr %in% 
                             c("Mega- meso- and micro- phanerophyte",
                               "Phanerophyte (mega-,meso-, nano-)",
                               "Mega-, meso- and microphanerophyte",
                               "megaphanerophyte", "Megaphanerophyte",
                               "Micro-phanerophytes", "Nanophanerophyt",
                               "nanophanerophyte", "Nanophanerophyte",
                               "phanerophyte", "Phanerophytes", "shrub", "tree")] = "Phanerophyte"
TRY_life_form$OrigValueStr[TRY_life_form$OrigValueStr %in% 
                             c("T", "therophyte", "Therophyte (annual land plant)", "Therophytes")] = "Therophyte"
```
And we exclude a number of records by focusing on these categories:
```{r}
TRY_life_form<-TRY_life_form[TRY_life_form$OrigValueStr %in% c("Chamaephyte", "Cryptophyte",
                                                               "Epiphyte", "Hemicryptophyte", "Hydrophyte",
                                                               "Phanerophyte")]
```

We can now call `populateTraits()` to parse TRY values:
```{r}
SpParams <- populateTraits(SpParams, TRY_life_form,
                           trait_mapping = c("LifeForm" = "OrigValueStr"),
                           character_traits = TRUE,
                           taxon_column = "AccSpeciesName")
```
Check which species have missing values:
```{r}
SpParams[is.na(SpParams$LifeForm), c("Name", "LifeForm")]
```
Since `LifeForm` is a strict parameter, we need to fill missing values manually:
```{r}
SpParams$LifeForm[SpParams$Name %in% c("Arctostaphylos uva-ursi", "Corema spp.", "Genistella spp.", 
                                       "Halimium spp.", "Lithodora spp.", "Phlomis spp.",
                                       "Sideritis spp.", "Vella spp.")] = "Chamaephyte"
SpParams$LifeForm[SpParams$Name %in% c("Ampelodesmos mauritanica")] = "Hemicryptophyte"
SpParams$LifeForm[SpParams$Name %in% c("Anagyris spp.", "Apollonias barbujana", "Gleditsia triacanthos",
                                       "Heberdenia bahamensis", "Picconia excelsa", "Pleiomeris canariensis",
                                       "Paliurus spina-christi", "Retama spp.", "Securinega tinctoria",
                                       "Teline spp.",
                                       "Visnea mocanera", "Zizyphus lotus")] = "Phanerophyte"
```
Show the number of taxa for each life form in the species parameter table is:
```{r}
table(SpParams$LifeForm, useNA = "ifany")
```


#### Leaf shape

A large number of leaf shape records can be found in TRY (its name in the data base is "Leaf type"). We begin by loading TRY leaf form data (TraitID = 43):
```{r leaf_shape}
TRY_leaf_shape  <-readRDS(file.path(DB_path, "TraitDatabases/TRY/TRY_traits/TRY_43.rds"))
```
Since leaf shape descriptions in TRY are not standardized, we standardize leaf shape values into five categories:
```{r}
TRY_leaf_shape$OrigValueStr[TRY_leaf_shape$OrigValueStr %in% 
                             c("broadleaved", "broad-leaved", "B", "broadleaf")] = "Broad"
TRY_leaf_shape$OrigValueStr[TRY_leaf_shape$OrigValueStr %in% 
                              c("needleleaved", "needle-leaved","needle","needle-leaf", "N")] = "Needle"
TRY_leaf_shape$OrigValueStr[TRY_leaf_shape$OrigValueStr %in% 
                              c("scale-shaped", "scale", "scale-like", "scale-leaf")] = "Scale"
TRY_leaf_shape$OrigValueStr[TRY_leaf_shape$OrigValueStr %in% 
                              c("narrowleaved", "linear", "Cylindrical")] = "Linear"
TRY_leaf_shape$OrigValueStr[TRY_leaf_shape$OrigValueStr %in% 
                              c("photosynthetic stem")] = "Spines"
TRY_leaf_shape<-TRY_leaf_shape[TRY_leaf_shape$OrigValueStr %in% 
                                 c("Broad", "Needle", "Scale", "Linear", "Spines"),
                               c("AccSpeciesName", "OrigValueStr")]
names(TRY_leaf_shape) = c("Species", "LeafShape")
```
Leaf shape data is also available from database BROT2 [@Tavsanoglu2018]. We also standardize BROT2 leaf shape values into the same five categories:
```{r}
brot2_LeafShape <- brot2 %>% filter(Trait %in% c("LeafShape")) %>% 
  select(c("Taxon", "Data"))
brot2_LeafShape$Data <- paste(toupper(substring(brot2_LeafShape$Data, 1, 1)), 
                              substring(brot2_LeafShape$Data, 2), sep = "")
brot2_LeafShape$Data[brot2_LeafShape$Data=="Needle-like"] = "Needle"
brot2_LeafShape$Data[brot2_LeafShape$Data=="Scale-like"] = "Scale"
names(brot2_LeafShape) = c("Species", "LeafShape")
```
Finally, we merge the two source data sets and call `populateTraits()` to parse leaf shape values:
```{r}
LeafShape_data = rbind(brot2_LeafShape,
                       TRY_leaf_shape)
SpParams <- populateTraits(SpParams, LeafShape_data, 
                           trait_mapping = c("LeafShape" = "LeafShape"),
                           character_traits = TRUE,
                           taxon_column = "Species")
```

Show which species have missing values:
```{r}
SpParams[is.na(SpParams$LeafShape), c("Name", "LeafShape")]
```

and, like before, we need to fill missing values manually:
```{r}
SpParams$LeafShape[SpParams$Name %in% c("Genistella spp.")] = "Broad"
SpParams$LeafShape[SpParams$Name %in% c("Tetraclinis articulata")] = "Scale"
```
The final number of taxa for each leaf shape in the species parameter table is:
```{r}
table(SpParams$LeafShape, useNA = "ifany")
```

#### Leaf size
Parameter `LeafSize` of **medfate** is categorical, and describes the size of leaves as "Small", "Medium" or "Large". TRY includes a trait called "Leaf area", that contains leaf area in mm2. We load TRY leaf area data using: 
```{r leaf_size}
TRY_leaf_area  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_3110.rds"))
TRY_leaf_area <- TRY_leaf_area[TRY_leaf_area$ErrorRisk < 4,]
```
and classify TRY leaf area values into th`e three leaf size categories:
```{r}
TRY_leaf_area$LeafSize <-cut(as.numeric(TRY_leaf_area$StdValue), 
                             breaks = c(0,225, 2025,100000), labels = c("Small", "Medium", "Large")) 
TRY_LeafSize<-TRY_leaf_area[,c("AccSpeciesName", "LeafSize")]
names(TRY_LeafSize)[1] = "Species"
```
Database BROT2 [@Tavsanoglu2018] contains both leaf area measurements and categorical leaf size data. We thus split BROT2 data into leaf area values and leaf size strings:
```{r}
brot2_LeafArea <- brot2 %>% filter(Trait %in% c("LeafArea")) %>% 
  select(c("Taxon", "Trait", "Data", "Units"))
brot2_LeafSize <- brot2_LeafArea[brot2_LeafArea$Units=="[5]",]
brot2_LeafArea <- brot2_LeafArea[brot2_LeafArea$Units=="mm2",]
```
We then reclassify BROT2 leaf size data into three categories 
```{r}
brot2_LeafSize$Data[brot2_LeafSize$Data=="very small"] = "small"
brot2_LeafSize$Data[brot2_LeafSize$Data=="very large"] = "large"
brot2_LeafSize$Data <- paste(toupper(substring(brot2_LeafSize$Data, 1, 1)), 
                              substring(brot2_LeafSize$Data, 2), sep = "")
brot2_LeafSize <- brot2_LeafSize[,c("Taxon", "Data")]
names(brot2_LeafSize) = c("Species", "LeafSize")
```
and classify BRO2 leaf area values into three categories, as we did with TRY values:
```{r}
brot2_LeafArea$LeafSize = cut(as.numeric(brot2_LeafArea$Data), 
                              breaks = c(0,225, 2025,100000), labels = c("Small", "Medium", "Large")) 
brot2_LeafArea <- brot2_LeafArea[,c("Taxon", "LeafSize")]
names(brot2_LeafArea)[1] = "Species"
```
We can now merge the three source data tables and call `populateTraits()` to parse leaf size values:
```{r}
LeafSize_data = rbind(brot2_LeafSize,
                      brot2_LeafArea,
                      TRY_LeafSize)

SpParams <- populateTraits(SpParams, LeafSize_data, 
                           trait_mapping = c("LeafSize" = "LeafSize"),
                           character_traits = TRUE,
                           taxon_column = "Species")
```
Show species with missing values:
```{r}
SpParams[is.na(SpParams$LeafSize), c("Name", "LeafSize")]
```
As before, we need to fill missing values manually:
```{r}
SpParams$LeafSize[SpParams$Name %in% c("Apollonias barbujana",
                                       "Ampelodesmos mauritanica",
                                       "Dracaena drago", "Heberdenia bahamensis",
                                       "Picconia excelsa", "Pleiomeris canariensis",
                                       "Visnea mocanera")] = "Large"
SpParams$LeafSize[SpParams$Name %in% c("Anagyris spp.","Coriaria myrtifolia", 
                                       "Gleditsia triacanthos", "Securinega tinctoria",
                                       "Kleinia neriifolia", "Rhododendron spp.",
                                       "Sophora japonica")] = "Medium"
SpParams$LeafSize[SpParams$Name %in% c("Adenocarpus spp.", 
                                       "Cedrus spp.",
                                       "Chamaecyparis lawsoniana",
                                       "Chamaespartium tridentatum",
                                       "Cytisophyllum sessilifolium",
                                       "Daboecia cantabrica",
                                       "Erinacea spp.", "Genistella spp.",
                                       "Paliurus spina-christi", "Pseudotsuga menziesii",
                                       "Spartium spp.", "Teline spp.",  "Thuja spp.", "Vella spp.",
                                       "Zizyphus lotus")] = "Small"
```

The final number of taxa for each leaf size in the species parameter table is:
```{r}
table(SpParams$LeafSize, useNA = "ifany")
```

#### Phenology type
A large number of leaf phenology type records can be found in TRY (TraitID 37). We load TRY leaf phenology type data first: 
```{r phenology_type}
TRY_phenology  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_37.rds"))
```
As with several previous parameters, leaf phenology type strings need to be standardized. Five phenology categories are used in **medfate**, four of which can be mapped with TRY strings:
```{r}
TRY_phenology$OrigValueStr[TRY_phenology$OrigValueStr %in% 
                             c("evergreen", "E", "EV", "always summer green",
                               "Evergreen broad-leaved", "Evergreen scale-like",
                               "Evergreen needle-leaved")] = "oneflush-evergreen"
TRY_phenology$OrigValueStr[TRY_phenology$OrigValueStr %in% 
                             c("deciduous", "D", "Deciduous broad-leaved", "Deciduous",
                               "Nonevergreen", "winter deciduous")] = "winter-deciduous"
TRY_phenology$OrigValueStr[TRY_phenology$OrigValueStr %in% 
                             c("semi-deciduous",
                               "semi-evergreen")] = "winter-semideciduous"
TRY_phenology$OrigValueStr[TRY_phenology$OrigValueStr %in% 
                             c("drought semi-deciduous")] = "drought-semideciduous"
TRY_phenology<-TRY_phenology[TRY_phenology$OrigValueStr %in% c("progressive-evergreen",
                                                               "oneflush-evergreen", 
                                                               "winter-semideciduous",
                                                               "winter-deciduous",
                                                               "drought-semideciduous")]
TRY_phenology<-TRY_phenology[,c("AccSpeciesName", "OrigValueStr")]
names(TRY_phenology) = c("Species","LeafPhenology")
```
Leaf phenology types are also available in BROT2 [@Tavsanoglu2018], so we select them and map the existing values into **medfate** categories:
```{r}
brot2_LeafPhenology <- brot2 %>% filter(Trait %in% c("LeafPhenology")) %>% 
  select(c("Taxon", "Data"))
brot2_LeafPhenology$Data[brot2_LeafPhenology$Data=="drought semi-deciduous"] = "drought-semideciduous"
brot2_LeafPhenology$Data[brot2_LeafPhenology$Data=="evergreen"] = "oneflush-evergreen"
brot2_LeafPhenology$Data[brot2_LeafPhenology$Data=="winter deciduous"] = "winter-deciduous"
brot2_LeafPhenology$Data[brot2_LeafPhenology$Data=="winter semi-deciduous"] = "winter-semideciduous"
names(brot2_LeafPhenology) = c("Species","LeafPhenology")
```
We can now merge the two source data sets and call `populateTraits()` to parse leaf phenology types:
```{r}
LeafPhenology_data =rbind(brot2_LeafPhenology,
                          TRY_phenology)
SpParams <- populateTraits(SpParams, LeafPhenology_data, 
                           trait_mapping = c("PhenologyType" = "LeafPhenology"),
                           character_traits = TRUE,
                           taxon_column = "Species")
```
We show which target entities have missing values after imputation:
```{r}
SpParams[is.na(SpParams$PhenologyType), c("Name", "PhenologyType")]
```
And since leaf phenology type is a strict parameter, we need to manually fill missing values :
```{r}
SpParams$PhenologyType[SpParams$Name %in% c("Vella spp.", "Genistella spp.", "Securinega tinctoria")] = "winter-deciduous"
SpParams$PhenologyType[SpParams$Name %in% c("Pleiomeris canariensis")] = "oneflush-evergreen"
```
The final number of taxa for each phenology type in the species parameter table are:
```{r}
table(SpParams$PhenologyType, useNA = "ifany")
```
A notable problem observed is that there are no winter-semideciduous species, as they are frequently recorded simply as winter-deciduous.

### Plant size

We already populated some plant size parameters from forest inventory data. Here we draw values for parameters that are not normally measured in forest inventory data (for obvious reasons), such as maximum rooting depth.

#### Maximum rooting depth
TRY trait 'Rooting depth' (ID 6) contains measurements of rooting depth for multiple species. We first load this dataset:
```{r rooting_depth}
TRY_Z95 <-readRDS(file.path(DB_path, "TraitDatabases/TRY/TRY_traits/TRY_6.rds"))
```
We filter those entries that we consider correspond to maximum rooting depths, which we will map to parameter `Z95` (depth corresponding to 95\% of fine roots):
```{r}
TRY_Z95 <-TRY_Z95[TRY_Z95$OriglName %in% c("Max Rooting Depth (m)",
                                           "Rooting depth",
                                           "Rooting depth (m)",
                                           "Rooting depth_max",
                                           "RootDepth"),]
TRY_Z95 <- TRY_Z95[TRY_Z95$ErrorRisk < 3, c("AccSpeciesName", "StdValue")]
names(TRY_Z95) <- c("Species", "Z95")
```
Rooting depth values are also available from BROT2 [@Tavsanoglu2018] (some of these records are also in TRY). As before, we assume values to correspond to parameter `Z95`.
```{r}
brot2_Z95 <- brot2 %>% filter(Trait %in% c("RootDepth")) %>% 
  select(c("Taxon", "Data"))
brot2_Z95$Data <- as.numeric(brot2_Z95$Data)
names(brot2_Z95) <- c("Species", "Z95")
```
We then merge the two data sets, trim rooting values to lie within the interval [15 cm, 10 m] and call the populating function, while transforming units from m to mm:
```{r}
Z95_data = rbind(TRY_Z95,
                 brot2_Z95)
Z95_data$Z95<- pmax(Z95_data$Z95, 0.15) # Minimum of 15 cm depth
Z95_data$Z95<- pmin(Z95_data$Z95, 10) # Maximum of 10 m depth
SpParams <- populateTraits(SpParams, Z95_data, 
                           trait_mapping = c("Z95" = "Z95"),
                           scalar_functions = c("Z95" = function(x) {round(x*1000)}),
                           taxon_column = "Species")
```
Show species with missing values, grouped by growth form:
```{r}
table(is.na(SpParams$Z95), SpParams$GrowthForm)
```
Since `Z95` is a strict parameter, we fill missing values with growth form averages: 
```{r}
SpParams$Z95[is.na(SpParams$Z95) & SpParams$GrowthForm != "Shrub"] = round(mean(SpParams$Z95[SpParams$GrowthForm != "Shrub"], na.rm=TRUE))
SpParams$Z95[is.na(SpParams$Z95) & SpParams$GrowthForm == "Shrub"] = round(mean(SpParams$Z95[SpParams$GrowthForm == "Shrub"], na.rm=TRUE))
```

The distribution of $Z95$ values for shrub species and non-shrub species can be shown using:
```{r}
summary(SpParams[SpParams$GrowthForm == "Shrub",c("Z95")])
summary(SpParams[SpParams$GrowthForm != "Shrub",c("Z95")])
```

### Leaf phenology

#### Budburst and senescence

Most parameters regulating leaf phenological phases (i.e. bud dormancy, budburst, senescence, bud formation, ...) require calibration exercises using observed weather data and phenological records. 

We compiled a data base with published leaf budburst models [@Vitasse2011;@Asse2020]:
```{r leaf_pheno}
budburst
```
and leaf senescence models [@Delpierre2009; @Vitasse2011]:
```{r}
senescence
```
Note that some species are repeated. Since we cannot average model parameters, due to parameter correlation, in cases of more than one model by species we selected the one based on a larger sample size (table column `N`), and codified the result in a table column called `Accepted`. We use the leaf budburst model data base to populate leaf phenology parameters:
```{r}
SpParams <- populateTraits(SpParams, budburst[budburst$Accepted=="Y",], 
                           trait_mapping = c("t0gdd" = "t0gdd",
                                             "Tbgdd" = "Tbgdd",
                                             "Sgdd" = "Sgdd"),
                           taxon_column = "Species",
                           erase_previous = TRUE)
```
Similarly, we use the senescence model database to fill the parameters regulating leaf senescence:
```{r}
SpParams <- populateTraits(SpParams, senescence[senescence$Accepted=="Y",], 
                           trait_mapping = c("Phsen" = "Phsen",
                                             "Tbsen" = "Tbsen",
                                             "Ssen" = "Ssen",
                                             "xsen" = "xsen",
                                             "ysen" = "ysen"),
                           taxon_column = "Species",
                           erase_previous = TRUE)
```

#### Leaf duration
Leaf lifespan is available from TRY for many species (TraitID 12). We load records for this trait, which is already standardized into month units:
```{r leaf_duration}
TRY_LeafDuration  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_12.rds"))
TRY_LeafDuration <- TRY_LeafDuration[TRY_LeafDuration$ErrorRisk<3, c("AccSpeciesName", "StdValue")]
names(TRY_LeafDuration) <- c("Species", "LeafDuration")
```
Leaf lifespan is also available from BROT2 [@Tavsanoglu2018], also in number of months:
```{r}
names(hydratry)[names(hydratry)=="LL"] = "LeafDuration"
brot2_LeafDuration <- brot2 %>% filter(Trait %in% c("LeafLifespan")) %>% 
  select(c("Taxon", "Data"))
brot2_LeafDuration$Data <- as.numeric(brot2_LeafDuration$Data)
names(brot2_LeafDuration) <- c("Species", "LeafDuration")
```
We then merge the two data sources and call function `populateTraits()` to map leaf lifespan to parameter `LeafDuration` ($LD$), and we specify a transformation of units from months to years:
```{r}
LeafDuration_data<-rbind(hydratry[,c("Species","LeafDuration")],
                         TRY_LeafDuration,
                         brot2_LeafDuration)
SpParams <- populateTraits(SpParams, LeafDuration_data, 
                                trait_mapping = c("LeafDuration" = "LeafDuration"),
                                scalar_functions = c("LeafDuration" = function(x) {x/12}),
                                taxon_column = "Species")
```
In this case, and the following ones, the trait is not strict, so that we do not need to manually fill missing values.

### Morphology and anatomy

In this subsection we map several plant morphological and anatomical traits into model parameters. All such traits in **medfate** are imputable, so that missing values do not need to be handled here.

#### Specific leaf area

Specific leaf area ($SLA$), in mm2/mg or m2/kg, is available from several databases. We load records from TRY (TraitID = 3117), which are already standardized:
```{r SLA}
TRY_SLA <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_3117.rds"))
TRY_SLA <- TRY_SLA[TRY_SLA$ErrorRisk<4, c("AccSpeciesName", "StdValue")]
names(TRY_SLA) <- c("Species", "SLA")
```
We also select SLA records from BROT2 [@Tavsanoglu2018], already in the same units as TRY.
```{r}
brot2_SLA <- brot2 %>% filter(Trait %in% c("SLA")) %>% 
  select(c("Taxon", "Data"))
brot2_SLA$Data <- as.numeric(brot2_SLA$Data)
names(brot2_SLA) <- c("Species", "SLA")
```
And we merge these two sources with SLA values coming from database HydraTRY [@Mencuccini2019;@SanchezMartinez2020]. The whole set is used to call `populateTraits()` and populate $SLA$ parameter values without any transformation:
```{r}
SLA_data = rbind(hydratry[,c("Species","SLA")],
                 TRY_SLA,
                 brot2_SLA)
SpParams <- populateTraits(SpParams, SLA_data, 
                           trait_mapping = list("SLA" = "SLA") ,
                           taxon_column = "Species")
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("SLA")])
```

#### Leaf density

Leaf density (leaf dry mass per leaf volume, in $g \cdot cm^{-3}$) is available in TRY (TraitID = 48). Since we do not have other data sources we directly use this data set to populate parameter `LeafDensity` ($LD$):
```{r leaf_density}
TRY_LeafDensity  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_48.rds"))
TRY_LeafDensity<- TRY_LeafDensity[TRY_LeafDensity$ErrorRisk < 3,, drop= FALSE] 
SpParams <- populateTraits(SpParams, TRY_LeafDensity, 
                           trait_mapping = c("LeafDensity" = "StdValue"),
                           taxon_column = "AccSpeciesName")
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("LeafDensity")])
```
#### Wood density
Wood density (stem dry mass per fresh stem volume, in $g \cdot cm^{-3}$) is available in TRY (TraitID = 4):
```{r wood_density}
TRY_WoodDensity  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_4.rds"))
TRY_WoodDensity <- TRY_WoodDensity[TRY_WoodDensity$ErrorRisk<3, c("AccSpeciesName", "StdValue")]
names(TRY_WoodDensity) <- c("Species", "WoodDensity")
```

Wood density records are also available from the 'Global Wood Density Database' [@Zanne2009]:
```{r}
gwdd_WoodDensity<-gwdd[,c("Binomial", "Wood density (g/cm^3), oven dry mass/fresh volume")]
names(gwdd_WoodDensity)<-c("Species", "WoodDensity")
```
Wood density is also part of HydraTRY database [@Mencuccini2019;@SanchezMartinez2020]:
```{r}
names(hydratry)[names(hydratry)=="WD"] = "WoodDensity"
```
and we can also find wood density estimates in BROT2 [@Tavsanoglu2018]:
```{r}
brot2_WoodDensity <- brot2 %>% filter(Trait %in% c("StemDensity")) %>% 
  select(c("Taxon", "Data"))
brot2_WoodDensity$Data <- as.numeric(brot2_WoodDensity$Data)
names(brot2_WoodDensity) <- c("Species", "WoodDensity")
```
We merge all these four sources and call `populateTraits()`:
```{r}
WoodDensity_data<-rbind(hydratry[,c("Species","WoodDensity")],
                        TRY_WoodDensity,
                        brot2_WoodDensity,
                        gwdd_WoodDensity)
SpParams <- populateTraits(SpParams, WoodDensity_data, 
                           trait_mapping = c("WoodDensity" = "WoodDensity"),
                           taxon_column = "Species")
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("WoodDensity")])
```
#### Fine fuel ratio
The fine fuel ratio, i.e. the ratio between dry mass of leaves and the dry mass of leaves and branchlets (< 6.35 mm), is a parameter that is harder to get from trait data bases. Here we use our own compilation of values to populate parameter `r635` ($r_{6.35}$):
```{r ffratio}
SpParams <- populateTraits(SpParams, as.data.frame(ffr), 
                           trait_mapping = c("r635" = "r635"),
                           taxon_column = "Species")
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("r635")])
```
#### Fraction of dead material 

The fraction of standing dead biomass ($P_{dead}$) was already populated for shrubs in the section devoted to allometries, although several target entities still have missing values:

```{r pDead}
table(is.na(SpParams$pDead))
```
We tried to reduce some of these missing values with data from BROT2 [@Tavsanoglu2018]. We first select "DeadFuel" records from this data base:
```{r}
brot2_DF <- brot2 %>% filter(Trait %in% c("DeadFuel")) %>% 
  select(c("Taxon", "Trait", "Data", "Units")) %>% 
  filter(Units=="%")
brot2_DF$Data <- as.numeric(brot2_DF$Data)
```
and populate the species parameter table, while transforming percentages into the [0-1] interval:
```{r}
SpParams <- populateTraits(SpParams, brot2_DF, 
                           trait_mapping = c("pDead" = "Data"),
                           scalar_functions = c("pDead"= function(x){x/100}),
                           taxon_column = "Taxon",
                           erase_previous = FALSE)
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("pDead")])
```

#### Leaf area to sapwood area ratio
The leaf area to sapwood area ratio `Al2As` is also known as the inverse of the Huber value. This trait can be found in the TRY database (TraitID = 171):
```{r Al2As}
TRY_Al2As  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_171.rds"))
```
Since either `Al2As` or the Huber value may have been recorded, we need to carefully inspect what was actually measured, and which measurement units were used: 
```{r}
TRY_Al2As$StdValue[TRY_Al2As$OriglName=="Huber value"]<-1/as.numeric(TRY_Al2As$OrigValueStr[TRY_Al2As$OriglName=="Huber value"])
TRY_Al2As$StdValue[TRY_Al2As$OriglName=="leaf.area.per.sapwood.area"]<-TRY_Al2As$StdValue[TRY_Al2As$OriglName=="leaf.area.per.sapwood.area"]*100
TRY_Al2As$StdValue[TRY_Al2As$OriglName=="Sapwood: leaf area ratio"]<-10000/as.numeric(TRY_Al2As$OrigValueStr[TRY_Al2As$OriglName=="Sapwood: leaf area ratio"])
TRY_Al2As$StdValue[TRY_Al2As$OriglName=="The ratio of leaf area attached per unit sapwood cross-section area (m2 cm-2)"]<-10000*as.numeric(TRY_Al2As$OrigValueStr[TRY_Al2As$OriglName=="The ratio of leaf area attached per unit sapwood cross-section area (m2 cm-2)"])
TRY_Al2As$StdValue[TRY_Al2As$OriglName=="values at base of living crown, m2/cm2"]<-10000*as.numeric(TRY_Al2As$OrigValueStr[TRY_Al2As$OriglName=="values at base of living crown, m2/cm2"])
TRY_Al2As$StdValue[TRY_Al2As$OriglName=="values at breast height, m2/cm2"]<-10000*as.numeric(TRY_Al2As$OrigValueStr[TRY_Al2As$OriglName=="values at breast height, m2/cm2"])
TRY_Al2As <- TRY_Al2As[TRY_Al2As$ErrorRisk <3, c("AccSpeciesName", "StdValue")]
names(TRY_Al2As) <- c("Species", "Al2As")
```
We use a summary of the data to check the range of values and median ones, which should make sense according to our knowledge of this variable:
```{r}
summary(TRY_Al2As$Al2As)
```

Huber values are also available from HydraTRY [@Mencuccini2019;@SanchezMartinez2020], so we transform them into `Al2As`:
```{r}
hydratry_Al2As <-hydratry[,c("Species","Hv")]
hydratry_Al2As$Hv<- 10000/hydratry_Al2As$Hv
names(hydratry_Al2As) <- c("Species", "Al2As")
```
Now we are ready to merge the two datasets and perform the population of the species parameter table:
```{r}
Hv_data<-rbind(TRY_Al2As,
               hydratry_Al2As)
SpParams <- populateTraits(SpParams, Hv_data,
                           trait_mapping = c("Al2As" = "Al2As"),
                           taxon_column = "Species")
```

The distribution of parameter values is:
```{r}
summary(SpParams[,c("Al2As")])
```

#### Fraction of conduits within sapwood
We used the database of @Morris2016 as primary source to populate the fraction of sapwood corresponding to tracheids or vessels (the remaining being axial or radial parenchyma):
```{r conduit2sapwood}
SpParams <- populateTraits(SpParams, morris[morris$Organ=="stem",],
                           trait_mapping = c("conduit2sapwood" = "Radial.and.Axial.Parenchyma.(%)"),
                           taxon_column = "Species.name",
                           scalar_functions=list("conduit2sapwood"=function(x){1-(x/100)}))
```
Note that we had to transform the values in the data base, which refer to the percentage of radial/axial parenchymatic tissue. The distribution of parameter values is:
```{r}
summary(SpParams[,c("conduit2sapwood")])
```

#### Leaf width
Leaf width ($LW$ in cm) records can be obtained from TRY (TraitID = 145), so we directly use this data source to populate the trait without transformation:
```{r leaf_width}
TRY_LW <-readRDS(file.path(DB_path, "TraitDatabases/TRY/TRY_traits/TRY_145.rds"))
SpParams <- populateTraits(SpParams, TRY_LW[TRY_LW$ErrorRisk<3,],
                           trait_mapping = c("LeafWidth" = "StdValue"),
                           taxon_column = "AccSpeciesName")
```

The distribution of parameter values is:
```{r}
summary(SpParams[,c("LeafWidth")])
```
#### Specific root length

Finally, specific fine root length ($SRL$) is the fine root length per fine root dry mass and can also be retrieved from TRY (TraitID = 614):
```{r SRL}
TRY_SRL  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_614.rds"))
SpParams <- populateTraits(SpParams, TRY_SRL[TRY_SRL$ErrorRisk<3,], 
                           trait_mapping = c("SRL" = "StdValue"),
                           taxon_column = "AccSpeciesName")

```

The distribution of parameter values is:
```{r}
summary(SpParams[,c("SRL")])
```
### Tissue moisture
Plant tissue moisture parameters in **medfate** include those specifying maximum/minimum moisture content in leaves/branchlets as well as parameters describing the pressure-volume curve of leaves and stem tissues.

#### Minimum/maximum fuel moisture content
Fuel moisture content is usually defined as the mass of water per dry mass of fuel tissue. We used the database of @Yebra2019 as primary source of fuel moisture content. We first update the name of some species in the data base:
```{r FMC}
fmc_yebra$Species.collected[fmc_yebra$Species.collected=="Rosmarinus officinalis"] = "Salvia officinalis"
```
and call `populateTraits()` to determine parameters `minFMC` ($FMC_{min}$) and `maxFMC` ($FMC_{max}$) using the 5\% and 95\% lowest quantile values, respectively:
```{r}
SpParams <- populateTraits(SpParams, fmc_yebra, 
                           trait_mapping = c("minFMC" = "LFMC.value"),
                           summary_function = function(x) {as.numeric(quantile(x,probs=0.05))},
                           summary_params = NULL,
                           taxon_column = "Species.collected")
SpParams <- populateTraits(SpParams, fmc_yebra, 
                           trait_mapping = c("maxFMC" = "LFMC.value"),
                           summary_function = function(x) {as.numeric(quantile(x,probs=0.95))},
                           summary_params = NULL,
                           taxon_column = "Species.collected")
```

We now complement these values with additional data assembled by ourselves, from publications or field work:
```{r}
FMC_data <- rbind(fmc_field[,c("Species name", "MinFMC_DW", "MaxFMC_DW")],
                  fmc_bib[,c("Species name", "MinFMC_DW", "MaxFMC_DW")])
FMC_data$`Species name`[FMC_data$`Species name`=="Rosmarinus officinalis"] = "Salvia officinalis"
```
Note that the call to `populateTraits()` includes `erase_previous = FALSE` to avoid overwriting values on those estimated from @Yebra2019:
```{r}
SpParams <- populateTraits(SpParams, as.data.frame(FMC_data), 
                           trait_mapping = c("minFMC" = "MinFMC_DW",
                                             "maxFMC" = "MaxFMC_DW"),
                           taxon_column = "Species name", erase_previous = FALSE)
```
The distribution of $FMC_{min}$ and $FMC_{max}$ parameter values is:
```{r}
summary(SpParams[,c("minFMC", "maxFMC")])
```

#### Leaf pressure-volume curve

Leaf osmotic potential at full turgor (`LeafPI0` or $\pi_{0,leaf}$) and leaf elastic modulus (`LeafEPS` or $\epsilon_{leaf}$) describe leaf pressure-volume curves. These traits can be obtained from TRY [@Kattge2011] and @Bartlett2012. We begin by changing the column names of this second database:
```{r PV_curves}
names(bartlett)[names(bartlett)=="πo.(MPa)"] = "LeafPI0"
names(bartlett)[names(bartlett)=="ε.(MPa)"] = "LeafEPS"
```
we then load TRY data for $\pi_{0,leaf}$ (TraitID = 188):
```{r}
TRY_LeafPI0  <-readRDS(file.path(DB_path, "TraitDatabases/TRY/TRY_traits/TRY_188.rds"))
TRY_LeafPI0 <- TRY_LeafPI0[TRY_LeafPI0$ErrorRisk<3, c("AccSpeciesName", "StdValue")]
names(TRY_LeafPI0) <- c("Species", "LeafPI0")
TRY_LeafPI0$LeafPI0 <- -1*TRY_LeafPI0$LeafPI0 
```
and TRY data for $\epsilon_{leaf}$ (TraitID = 190):
```{r}
TRY_LeafEPS  <-readRDS(file.path(DB_path, "TraitDatabases/TRY/TRY_traits/TRY_190.rds"))
TRY_LeafEPS$StdValue <- as.numeric(TRY_LeafEPS$OrigValueStr) 
TRY_LeafEPS <- TRY_LeafEPS[, c("AccSpeciesName", "StdValue"), drop = FALSE]
names(TRY_LeafEPS) <- c("Species", "LeafEPS")
```
We then fill $\pi_{0,leaf}$ with the TRY and @Bartlett2012 data sources merged:
```{r}
LeafPI0_data = rbind(TRY_LeafPI0[, c("Species", "LeafPI0")],
                     bartlett[,c("Species", "LeafPI0")])
SpParams <- populateTraits(SpParams, LeafPI0_data, 
                           trait_mapping = c("LeafPI0" = "LeafPI0"),
                           taxon_column = "Species")
```
And we do the same for $\epsilon_{leaf}$:
```{r}
LeafEPS_data = rbind(TRY_LeafEPS[, c("Species", "LeafEPS")],
                     bartlett[,c("Species", "LeafEPS")])
SpParams <- populateTraits(SpParams, LeafEPS_data, 
                           trait_mapping = c("LeafEPS" = "LeafEPS"),
                           taxon_column = "Species")
```

Finally, we use @Bartlett2012 to provide suitable values for the fraction of apoplastic tissue (`LeafAF` or $f_{apo,leaf}$):
```{r}
SpParams <- populateTraits(SpParams, bartlett, 
                           trait_mapping = c("LeafAF" = "af"),
                           taxon_column = "Species")
```

The statistical distributions of the three parameters is summarized below:
```{r}
summary(SpParams[,c("LeafPI0", "LeafEPS", "LeafAF")])
```
### Fuel flammability

Fuel flammability properties are needed to predict potential fire behavior on forest stands on the basis of its composition and structure.

#### Heat content and surface area to volume ratio

We compiled  in a database bibliographic values of calorific heat value ($h$, in $J \cdot kg^{-1}$) and the surface area to volume ratio ($\sigma$, in $m^2 \cdot m^{-3}$) for different Mediterranean plant species. We use this data set to populate the corresponding **medfate** parameters:

```{r heat_content_SAV}
SpParams <- populateTraits(SpParams, as.data.frame(flam), 
                           trait_mapping = c("HeatContent" = "High calorific value [J/kg]",
                                             "SAV" = "SurfaceAreaVolume[m2/m3]"),
                           taxon_column = "Species name")
```
The statistical distributions of the two parameters is summarized below:
```{r}
summary(SpParams[,c("HeatContent", "SAV")])
```
#### Lignin content
The lignin content per dry mass of leaves is included in TRY (TraitID = 87) with standardized units mg/g. To parse this trait into `LigninPercent` ($LI$) we simply divide values by 10.

```{r lignin}
TRY_Lignin  <-readRDS(file.path(DB_path, "TraitDatabases/TRY/TRY_traits/TRY_87.rds"))

SpParams <- populateTraits(SpParams, TRY_Lignin[TRY_Lignin$ErrorRisk<3,], 
                           trait_mapping = c("LigninPercent" = "StdValue"),
                           scalar_functions = c("LigninPercent" = function(x){x/10}), #mg/g to % 
                           taxon_column = "AccSpeciesName")
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("LigninPercent")])
```

### Transpiration/photosynthesis (Granier's model)
The three following subsections provide suitable values for parameters regulating transpiration and photosynthesis under the Granier transpiration model.

#### Water potential determining stomatal closure

Parameter $\Psi_{extract}$ (`Psi_Extract`) is the water potential at which transpiration is 50% of its maximum, for a given species. We decided to populate values for this parameter using the water potential corresponding to turgor loss point ($\Psi_{tlp}$), which is available in @Bartlett2012 and HydraTRY [@Mencuccini2019;@SanchezMartinez2020]:

```{r Psi_Extract}
names(bartlett)[names(bartlett)=="πtlp.(MPa)"] = "Ptlp"
names(hydratry)[names(hydratry)=="PItlp"] = "Ptlp"
```
However, we did not take $\Psi_{tlp}$ directly as a value for $\Psi_{extract}$ but assumed that $\Psi_{tlp}$ corresponded to 5% of maximum transpiration. Using $\Psi_{tlp}$ and the equation that relates  $\Psi_{extract}$ with relative whole plant transpiration $\Psi_{extract}$ is $\Psi_{extract} = \left[ln(0.5)/ln(0.05)\right]^{(1/r)} \cdot \Psi_{tlp}$, where $r = 3$. The mapping is done by calling `populateTraits()` on a merged dataset:
```{r}
Ptlp_data = rbind(martin_stpaul[, c("Species", "Ptlp")],
                  hydratry[, c("Species", "Ptlp")],
                  bartlett[,c("Species", "Ptlp")])
SpParams <- populateTraits(SpParams, Ptlp_data, 
                           trait_mapping = list("Psi_Extract" = "Ptlp"),
                           taxon_column = "Species",
                           scalar_functions = c("Psi_Extract" = function(x){x*(log(0.5)/log(0.05))^(1/3)}))
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("Psi_Extract")])
```
#### Critical water potential
Parameter $\Psi_{critic}$ (`Psi_Critic`) is the water potential corresponding to 50\% of loss of hydraulic conductance in the stem, also known as $\Psi_{50}$. This trait can be found in three data bases [@Martin-StPaul2017;@Choat2012;@Mencuccini2019;@SanchezMartinez2020]: 
```{r Psi_Critic}
P50_data = rbind(martin_stpaul[, c("Species", "P50")],
                 choat[, c("Species", "P50")],
                 hydratry[, c("Species", "P50")])
P50_data$P50 <- as.numeric(P50_data$P50)
```
TRY also contains data from xylem vulnerability curves, which mostly overlaps other data sources, but from which we can perhaps obtain additional $\Psi_{50}$ records [@Pausas2016;@Manzoni2013]:
```{r}
TRY_VC <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_719.rds"))
VC_str <- c("Reference P50",
            "Xylem water potential at which 50% of conductivity is lost (P50)",
            "Stem P50")
TRY_P50 <- TRY_VC[TRY_VC$DataName %in% VC_str, c("AccSpeciesName", "OrigValueStr")]
names(TRY_P50)<-c("Species", "P50")
```
We convert string $\Psi_{50}$ records to numeric values and change the sign of positive values:
```{r}
TRY_P50$P50 <- as.numeric(TRY_P50$P50)
TRY_P50 <- TRY_P50[!is.na(TRY_P50$P50),]
TRY_P50$P50[TRY_P50$P50>0] <- -TRY_P50$P50[TRY_P50$P50>0]
```
And we can now merge TRY data with the other sources:
```{r}
P50_data = rbind(P50_data, TRY_P50)
```

The mapping is done by calling `populateTraits()` on a merged dataset:
```{r}
SpParams <- populateTraits(SpParams, P50_data, 
                           trait_mapping = list("Psi_Critic" = "P50"),
                           taxon_column = "Species")
```

We want to ensure that $\Psi_{critic} \leq \Psi_{extract}$ by using:
```{r}
SpParams$Psi_Critic = pmin(SpParams$Psi_Critic, SpParams$Psi_Extract, na.rm=TRUE)
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("Psi_Critic")])
```

### Transpiration/photosynthesis (Sperry's model)

Simulations with the Sperry's transpiration model require many species-specific parameters, several of them unavailable in plant trait data bases. Here we make an effort to provide suitable values for some of them, particularly stem hydraulic efficiency and vulnerability. Others are estimated in **medfate** via inbuilt imputation or scaling procedures, as described in medfate's reference book. 

#### Minimum/maximum stomatal conductance
Minimum and maximum values of leaf conductance to water vapor ($G_{swmin}$ and $G_{swmax}$) can be drawn from databases of @Duursma2018 and @Hoshika2018, respectively. We first populate $G_{swmin}$ using data from @Duursma2018, while changing the units from $mmol\cdot s^{-1} \cdot m^{-2}$ to $mol\cdot s^{-1} \cdot m^{-2}$:
```{r stomatal_cond}
duursma$gmin <- as.numeric(duursma$gmin)
SpParams <- populateTraits(SpParams, duursma, 
                           trait_mapping = list("Gswmin" = "gmin"),
                           taxon_column = "species",
                           scalar_functions = list("Gswmin" = function(x) {x/1000})) 
```
We then populate $G_{swmax}$ using data from @Hoshika2018 without any change in units:
```{r}
SpParams <- populateTraits(SpParams, hoshika, 
                           trait_mapping = list("Gswmax" = "gmax.(mol.m-2.s-1)") ,
                           taxon_column = "Species")
```
The statistical distributions of the two parameters is summarized below:
```{r}
summary(SpParams[,c("Gswmin", "Gswmax")])
```

#### Stem maximum hydraulic conductivity
Stem maximum hydraulic conductivity ($K_{stem, max, ref}$) is an important parameter used in **medfate** to estimate stem maximum conductance, which defines plant hydraulic efficiency. Here we populate some values from HydraTRY [@Mencuccini2019;@SanchezMartinez2020]: 

```{r Kmax_stemxylem}
SpParams <- populateTraits(SpParams, hydratry, 
                           trait_mapping = c("Kmax_stemxylem" = "Ks"),
                           taxon_column = "Species")
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("Kmax_stemxylem")])
```
#### Stem hydraulic vulnerability
Stem hydraulic vulnerability is described in **medfate** using Weibull-type curves, which are described using two parameters. Trait databases normally report vulnerability using water potential corresponding to different levels of conductance loss, i.e. $\Psi_{12}$, $\Psi_{50}$, $\Psi_{88}$, etc. Here we first estimate Weibull coefficients from @Martin-StPaul2017 data on $\Psi_{12}$, $\Psi_{50}$:
```{r hydraulic_vulnerability}
martin_stpaul$VCstem_c <- NA
martin_stpaul$VCstem_d <- NA
for(i in 1:nrow(martin_stpaul)) {
  if((!is.na(martin_stpaul$P50[i])) && (!is.na(martin_stpaul$P12[i]))) {
    weib = hydraulics_psi2Weibull(psi50 = martin_stpaul$P50[i],
                                  psi12 = martin_stpaul$P12[i])
    martin_stpaul$VCstem_c[i] = weib[["c"]]
    martin_stpaul$VCstem_d[i] = weib[["d"]]
  }
}
```
and we do something similar from @Choat2012 data on $\Psi_{50}$, $\Psi_{88}$:
```{r}
choat$P50 = as.numeric(choat$P50)
choat$P88 = as.numeric(choat$P88)
choat$VCstem_c <- NA
choat$VCstem_d <- NA
for(i in 1:nrow(choat)) {
  if((!is.na(choat$P50[i])) && (!is.na(choat$P88[i]))) {
    weib = hydraulics_psi2Weibull(psi50 = choat$P50[i],
                                  psi88 = choat$P88[i])
    choat$VCstem_c[i] = weib[["c"]]
    choat$VCstem_d[i] = weib[["d"]]
  }
}
```



We then merge the two data sets and populate Weibull coefficients using `populateTraits()`:
```{r}
VC_data = rbind(martin_stpaul[, c("Species", "VCstem_c", "VCstem_d")],
                choat[, c("Species", "VCstem_c", "VCstem_d")])
SpParams <- populateTraits(SpParams, VC_data, 
                           trait_mapping = list("VCstem_c" = "VCstem_c",
                                                "VCstem_d" = "VCstem_d"),
                           taxon_column = "Species")
```

The statistical distributions of the stem vulnerability parameters is summarized below:
```{r}
summary(SpParams[,c("VCstem_c", "VCstem_d")])
```


#### Photosynthetic capacity

TRY offers $V_{max}$ values (TraitID = 186), which we take as estimates for **medfate**'s $V_{max298}$ the maximum Rubisco carboxilation rate at 25 degrees Celsius: 
```{r}
TRY_Vmax298  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_186.rds"))
SpParams <- populateTraits(SpParams, TRY_Vmax298[TRY_Vmax298$ErrorRisk<3, ], 
                           trait_mapping = c("Vmax298" = "StdValue"),
                           taxon_column = "AccSpeciesName")
```
The same happens with TRY's $J_{max}$ values (TraitID = 269), which we take as estimates for **medfate**'s $J_{max298}$ the maximum electron transport rate at 25 degrees Celsius: 
```{r}
TRY_Jmax298  <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_269.rds"))
SpParams <- populateTraits(SpParams, TRY_Jmax298[TRY_Jmax298$ErrorRisk<3, ], 
                           trait_mapping = c("Jmax298" = "StdValue"),
                           taxon_column = "AccSpeciesName")
```

The statistical distributions of the two parameters is summarized below:
```{r}
summary(SpParams[,c("Vmax298", "Jmax298")])
```

### Carbon balance and growth
Carbon balance and growth simulations require many parameters some of which are hard to measure experimentally.

#### Maintenance respiration rates
Baseline maintenance respiration rates for each tissue (leaf, sapwood and fine roots) are species-specific parameters in the model. Here we first leaf dark respiration rates per leaf mass values from TRY (TraitID = 41):

```{r }
TRY_RERleaf <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_41.rds"))
TRY_RERleaf <- TRY_RERleaf[TRY_RERleaf$ErrorRisk<3, c("AccSpeciesName", "StdValue")]
```

We now call the population function, while transforming units of micro mol C g-1 s-1 to g gluc g-1 d-1:
```{r}
SpParams <- populateTraits(SpParams, TRY_RERleaf, 
                           trait_mapping = c("RERleaf" = "StdValue"),
                           scalar_functions = c("RERleaf" = function(x){24.0*3600.0*(x/6.0)*(1e-6)*180.156}),
                           taxon_column = "AccSpeciesName")
```

```{r}
summary(SpParams[,c("RERleaf")])
```

Sapwood respiration is much harder to measure and database values are not retrieved at present.

#### Nitrogen content of tissues
The model allows estimating leaf and fine root respiration rates from nitrogen concentration. Hence it is useful to draw this latter parameter from trait databases. Sapwood respiration rates are given a default value when missing, but we also draw sapwood N concentrations. Here we first extract nitrogen concentration per leaf mass values from TRY (TraitID = 14):
```{r }
TRY_Nleaf <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_14.rds"))
TRY_Nleaf <- TRY_Nleaf[TRY_Nleaf$ErrorRisk<3, c("AccSpeciesName", "StdValue")]
```
We now call the population function:
```{r}
SpParams <- populateTraits(SpParams, TRY_Nleaf, 
                           trait_mapping = c("Nleaf" = "StdValue"),
                           taxon_column = "AccSpeciesName")
```
And we do the same for sapwood and fine roots (TraitIDs = 1229 and 475, respectively):
```{r nitrogen_content}
TRY_Nsapwood <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_1229.rds"))
TRY_Nfineroot <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_475.rds"))
TRY_Nsapwood <- TRY_Nsapwood[TRY_Nsapwood$ErrorRisk<3, c("AccSpeciesName", "StdValue")]
TRY_Nfineroot <- TRY_Nfineroot[TRY_Nfineroot$ErrorRisk<3, c("AccSpeciesName", "StdValue")]
SpParams <- populateTraits(SpParams, TRY_Nsapwood, 
                           trait_mapping = c("Nsapwood" = "StdValue"),
                           taxon_column = "AccSpeciesName")
SpParams <- populateTraits(SpParams, TRY_Nfineroot, 
                           trait_mapping = c("Nfineroot" = "StdValue"),
                           taxon_column = "AccSpeciesName")
```

The statistical distributions of the three parameters is summarized below:
```{r}
summary(SpParams[,c("Nleaf", "Nsapwood" ,"Nfineroot")])
```
#### Wood carbon content
Stem carbon content per stem dry mass is given by TRY (TraitID = 407) and can be easily mapped, after transforming units from mg/g to g/g: 
```{r WoodC}
TRY_WoodC <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_407.rds"))
SpParams <- populateTraits(SpParams, TRY_WoodC[TRY_WoodC$ErrorRisk<3,], 
                           trait_mapping = c("WoodC" = "StdValue"),
                           scalar_functions = c("WoodC" = function(x){x/1000}), 
                           taxon_column = "AccSpeciesName")
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("WoodC")])
```


#### Construction costs
Tissue construction costs are parsed from our own trait compilation from different sources:
```{r CCleaf}
 SpParams <- populateTraits(SpParams, as.data.frame(lcc_data), 
                            trait_mapping = c("CCleaf" = "CC (g glu g−1)"),
                            taxon_column = "Species")
 SpParams <- populateTraits(SpParams, as.data.frame(scc_data), 
                            trait_mapping = c("CCsapwood" = "CC (g glu g−1)"),
                            taxon_column = "Species")
 SpParams <- populateTraits(SpParams, as.data.frame(fcc_data), 
                            trait_mapping = c("CCfineroot" = "CC (g glu g−1)"),
                            taxon_column = "Species")
```
The distribution of parameter values is:
```{r}
summary(SpParams[,c("CCleaf", "CCsapwood", "CCfineroot")])
```

#### Relative starch for sapwood growth
Relative starch for sapwood growth ($RSSG$) is a threshold indicating when sapwood formation is possible. For relative starch levels below this threshold sapwood growth is not possible, so that carbon is saved for maintenance and replacement of leaves/fine roots. This parameter is related to shade tolerance, because a high $RSSG$ allows the plant to survive with low carbon assimilation at the cost of not growing. 

Shadow tolerance is available at TRY (TraitID = 603), but data is in several scales depending on the data set. We selected data set 49, which corresponds to @Niinemets2006 and describes shade tolerance in a scale from 1 to 5:
```{r RSSG}
TRY_ShadeTol <-readRDS(file.path(DB_path,"TraitDatabases/TRY/TRY_traits/TRY_603.rds"))
TRY_ShadeTol <- TRY_ShadeTol[TRY_ShadeTol$DatasetID==49,]
```

We estimated $RSSG$ as a linear function of shade tolerance, with 1 corresponding to 25% of relative starch and 5 corresponding to 95%:
```{r}
TRY_ShadeTol$OrigValueStr<- as.numeric(TRY_ShadeTol$OrigValueStr)
SpParams <- populateTraits(SpParams, TRY_ShadeTol, 
                           trait_mapping = c("RSSG" = "OrigValueStr"),
                           scalar_functions = c("RSSG" = function(x){min(0.95,0.25+0.70*((x-1)/2))}), 
                           taxon_column = "AccSpeciesName")
```

The distribution of parameter values after estimation is:
```{r}
summary(SpParams[,c("RSSG")])
```


<!-- ```{r save_SpParams, eval = params$spparams_rebuild, include = FALSE} -->
<!--   wb <- openxlsx::createWorkbook() -->
<!--   openxlsx::addWorksheet(wb, "InitialSpParamsMED") -->
<!--   openxlsx::writeDataTable(wb, "InitialSpParamsMED", SpParams) -->
<!--   openxlsx::saveWorkbook(wb,"../../data-raw/InitialSpParamsMED.xlsx", overwrite=TRUE) -->
<!-- ``` -->

## References