---
title: "Species parameter tables"
author: "Miquel De Cáceres"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: TRUE
bibliography: ../medfate.bib
params:
  spparams_rebuild: FALSE
  IFN_rebuild: FALSE
---

## Introduction



This vignette describes the procedures used to obtain the species parameter data tables (`SpParamsES`, `SpParamsFR` and `SpParamsUS`) included in package **medfate**. Each species parameter table is meant to be used on forest plot locations belonging to a given National Forest Inventory (NFI). Hence, the taxon entities chosen (rows in `SpParams`) are intended to represent plant taxa reported in the NFI. 

>**IMPORTANT**: The vignette is not self-contained, in the sense that it cannot be reproduced without access to data sets that are not included. Nevertheless, it is intended to serve as example of species parameterization for other regions.

### Required packages`
Installed version of **medfate** should be **ver 2.9** or higher. 
```{r, eval = FALSE}
install.packages("medfate")
```

If the version at CRAN is older, the package should be installed from GitHub via `remotes::install_github()`:
```{r, eval = FALSE}
remotes::install_github("emf-creaf/medfate")
```

Package **medfateutils** is required for the species parameterization process, as it includes several functions that facilitate parsing data sources into parameter values. Package **medfateutils** is installed only from GitHub:

```{r, eval = FALSE}
remotes::install_github("emf-creaf/medfateutils")
```

Package **medfuels** is also needed because it includes allometric coefficients for Mediterranean shrub species. As before, package **medfuels** is installed only from GitHub:

```{r, eval = FALSE}
remotes::install_github("spif-ctfc/medfuels")
```

Once we have **medfate**, **medfateutils** and  **medfuels**, we load them and other common packages that we will employ in this vignette:
```{r message=FALSE}
library(medfate)
library(medfateutils)
library(medfuels)
# library(IFNread)
library(tidyverse)
library(openxlsx)
```

### Target parameters

The set of species parameters needed to run the models included in **medfate** are described in a dataframe called `SpParamsDefinition`: 


```{r}
data("SpParamsDefinition")
SpParamsDefinition$ParameterName
```
It is obvious that such a sheer number of parameters cannot be searched for manually, even for a small number of target species. Hence, it is important to draw values from allometric and plant trait data bases. Not all model parameters have their counterpart in publicly-available databases, and even when using data bases it is unlikely that one will find appropriate parameter values for all species. To help with this situation, **medfate** has *inbuilt imputation procedures* to estimate missing values for many parameters before starting model simulations. However, imputed parameter estimates will normally be worse than estimates coming from data bases. Hence, we should put a strong effort in finding source parameter data before relying on inbuilt imputation.


### Steps to build a species parameter table

The following sections describe different steps that we used to obtain the species parameter table, which can be grouped into five main tasks:

 1. Initialize the parameter table with target taxonomic entities.
 2. Draw species parameters from forest inventory (IFN in our case) data.
 3. Populate tree and shrub allometric coefficients from suitable allometric databases.
 4. Populate plant functional traits from available data bases
 
 

## References