---
title: "Water pools and species interactions"
author: "Miquel De Caceres (CREAF)"
date: "`r Sys.Date()`"
description: >
   Preliminary results comparing simulations with different assumptions of root overlap between species
output:
  rmarkdown::html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(cowplot)
library(medfate)

base_pools = "/home/miquel/OneDrive/mcaceres_work/model_development/medfate_development/PlantWaterPools/"
```

## About this vignette

By default, simulations in **medfate** assume that soil moisture under all plant cohorts is the same (i.e. water sources corresponding to vertical soil layers are shared among cohorts). Therefore, the models neglect spatial variation in soil moisture observed in mixed forest stands (Schume et al. 2004). Variations in soil moisture beneath plant cohorts (and, implicitly, horizontal variation of soil moisture) can be simulated in **medfate** if required by the user (see control parameter `rhizosphereOverlap`). This involves considering that a given plant cohort will perform water uptake from the **water pool** surrounding its roots, whereas it may not have access to the water beneath other plants. However, there can exist some degree of horizontal overlap between water pools exploited by different plants.

The aim of this vignette is to illustrate the sensitivity of the advanced water balance model (De Cáceres et al. 2021) to the consideration of independent water pools. 

```{r, echo = FALSE}
WB_FB_sperry_total = readRDS(paste0(base_pools,"Rdata/Output/fb_sperry_total.rds"))
WB_FB_sperry_partial = readRDS(paste0(base_pools,"Rdata/Output/fb_sperry_partial.rds"))
WB_FB_sperry_none = readRDS(paste0(base_pools,"Rdata/Output/fb_sperry_none.rds"))
WB_PR_sperry_total = readRDS(paste0(base_pools,"Rdata/Output/pr_sperry_total.rds"))
WB_PR_sperry_partial = readRDS(paste0(base_pools,"Rdata/Output/pr_sperry_partial.rds"))
WB_PR_sperry_none = readRDS(paste0(base_pools,"Rdata/Output/pr_sperry_none.rds"))
WB_PU_sperry_total = readRDS(paste0(base_pools,"Rdata/Output/pu_sperry_total.rds"))
WB_PU_sperry_partial = readRDS(paste0(base_pools,"Rdata/Output/pu_sperry_partial.rds"))
WB_PU_sperry_none = readRDS(paste0(base_pools,"Rdata/Output/pu_sperry_none.rds"))
# WB_CB_sperry_total = readRDS(paste0(base_pools,"Rdata/Output/cb_sperry_total.rds"))
# WB_CB_sperry_partial = readRDS(paste0(base_pools,"Rdata/Output/cb_sperry_partial.rds"))
# WB_CB_sperry_none = readRDS(paste0(base_pools,"Rdata/Output/cb_sperry_none.rds"))
WB_ES_sperry_total = readRDS(paste0(base_pools,"Rdata/Output/es_sperry_total.rds"))
WB_ES_sperry_partial = readRDS(paste0(base_pools,"Rdata/Output/es_sperry_partial.rds"))
WB_ES_sperry_none = readRDS(paste0(base_pools,"Rdata/Output/es_sperry_none.rds"))
# WB_RO_sperry_total = readRDS(paste0(base_pools,"Rdata/Output/ro_sperry_total.rds"))
# WB_RO_sperry_partial = readRDS(paste0(base_pools,"Rdata/Output/ro_sperry_partial.rds"))
# WB_RO_sperry_none = readRDS(paste0(base_pools,"Rdata/Output/ro_sperry_none.rds"))

WB_FB_sureau_total = readRDS(paste0(base_pools,"Rdata/Output/fb_sureau_total.rds"))
WB_FB_sureau_partial = readRDS(paste0(base_pools,"Rdata/Output/fb_sureau_partial.rds"))
WB_FB_sureau_none = readRDS(paste0(base_pools,"Rdata/Output/fb_sureau_none.rds"))
WB_PR_sureau_total = readRDS(paste0(base_pools,"Rdata/Output/pr_sureau_total.rds"))
WB_PR_sureau_partial = readRDS(paste0(base_pools,"Rdata/Output/pr_sureau_partial.rds"))
WB_PR_sureau_none = readRDS(paste0(base_pools,"Rdata/Output/pr_sureau_none.rds"))
WB_PU_sureau_total = readRDS(paste0(base_pools,"Rdata/Output/pu_sureau_total.rds"))
WB_PU_sureau_partial = readRDS(paste0(base_pools,"Rdata/Output/pu_sureau_partial.rds"))
WB_PU_sureau_none = readRDS(paste0(base_pools,"Rdata/Output/pu_sureau_none.rds"))
# WB_CB_sureau_total = readRDS(paste0(base_pools,"Rdata/Output/cb_sureau_total.rds"))
# WB_CB_sureau_partial = readRDS(paste0(base_pools,"Rdata/Output/cb_sureau_partial.rds"))
# WB_CB_sureau_none = readRDS(paste0(base_pools,"Rdata/Output/cb_sureau_none.rds"))
WB_ES_sureau_total = readRDS(paste0(base_pools,"Rdata/Output/es_sureau_total.rds"))
WB_ES_sureau_partial = readRDS(paste0(base_pools,"Rdata/Output/es_sureau_partial.rds"))
WB_ES_sureau_none = readRDS(paste0(base_pools,"Rdata/Output/es_sureau_none.rds"))
# WB_RO_sureau_total = readRDS(paste0(base_pools,"Rdata/Output/ro_sureau_total.rds"))
# WB_RO_sureau_partial = readRDS(paste0(base_pools,"Rdata/Output/ro_sureau_partial.rds"))
# WB_RO_sureau_none = readRDS(paste0(base_pools,"Rdata/Output/ro_sureau_none.rds"))
```

```{r, echo = FALSE, warning = FALSE}
plot_one<-function(plot_name, sim_total, sim_partial, sim_none, type, ylim = NULL) {
g_total = plot(sim_total, type, bySpecies = TRUE, ylim = ylim) + ggplot2::labs(title= paste0("Total overlap (shared)"))+ theme(axis.text.x = element_blank(), axis.ticks = element_blank())+ylab(plot_name)
g_partial = plot(sim_partial, type, bySpecies = TRUE, ylim = ylim) + ggplot2::labs(title= paste0("Partial overlap"))+ theme(axis.text.x = element_blank(), axis.ticks = element_blank())+ylab("")
g_none = plot(sim_none, type, bySpecies = TRUE, ylim = ylim) + ggplot2::labs(title= paste0("No overlap (independent)"))+ theme(axis.text.x = element_blank(), axis.ticks = element_blank())+ylab("")

l = get_legend(g_total)
g_total <- g_total+theme(legend.position = "none")
g_partial<-g_partial +theme(legend.position = "none")
g_none<-g_none +theme(legend.position = "none")
plot_grid(g_total,g_partial, g_none, l, nrow=1, rel_widths = c(1,1, 1, 0.5))
}
```

```{r, echo = FALSE, warning = FALSE}
plot_several<-function(type, ylim = NULL) {
fb_sperry_plot = plot_one("FONTBL. - SPERRY", WB_FB_sperry_total, WB_FB_sperry_partial, WB_FB_sperry_none, type, ylim)
fb_sureau_plot = plot_one("FONTBL. - SUREAU", WB_FB_sureau_total, WB_FB_sureau_partial, WB_FB_sureau_none, type, ylim)
pr_sperry_plot = plot_one("PRADES - SPERRY", WB_PR_sperry_total, WB_PR_sperry_partial, WB_PR_sperry_none, type, ylim)
pr_sureau_plot = plot_one("PRADES - SUREAU", WB_PR_sureau_total, WB_PR_sureau_partial, WB_PR_sureau_none, type, ylim)
pu_sperry_plot = plot_one("PUECHABON - SPERRY", WB_PU_sperry_total, WB_PU_sperry_partial, WB_PU_sperry_none, type, ylim)
pu_sureau_plot = plot_one("PUECHABON - SUREAU", WB_PU_sureau_total, WB_PU_sureau_partial, WB_PU_sureau_none, type, ylim)
es_sperry_plot = plot_one("ARMALLON. - SPERRY", WB_ES_sperry_total, WB_ES_sperry_partial, WB_ES_sperry_none, type, ylim)
es_sureau_plot = plot_one("ARMALLON. - SUREAU", WB_ES_sureau_total, WB_ES_sureau_partial, WB_ES_sureau_none, type, ylim)
# ro_plot = plot_one("RONDA (2014)", WB_RO_sperry_total, WB_RO_sperry_partial, WB_RO_sperry_none, type, ylim)
cowplot::plot_grid(fb_sperry_plot, fb_sureau_plot, 
                   pr_sperry_plot, pr_sureau_plot,
                   pu_sperry_plot, pu_sureau_plot,
                   es_sperry_plot, es_sureau_plot,
                   nrow = 8)
}

```

## Simulations
As target forest stands we take five experimental plots whose data was already used when presenting the water balance model in De Cáceres et al. (2021). An evaluation of model performance in these experimental plots is given in [*Stand-level evaluation*](https://emf-creaf.github.io/medfate/articles/evaluation/StandLevelEvaluation.html). 

For each of these plots we performed two simulations with the advanced water balance model (`transpirationMode = "Sperry"`), one using the default assumption of complete rhizosphere overlap (`rhizosphereOverlap = "total"`) and the other assuming complete independence of water pools (`rhizosphereOverlap = "none"`). Recovery from embolism was assumed to depend on soil moisture (`stemCavitationRecovery = "rate"` and `leafCavitationRecovery = "rate"`). We performed a burn-in simulation for the available weather of the forest plot (1 to 3 years, depending on the plot), so that soil moisture, plant water potentials and xylem percent loss conductance (PLC) could be equilibrated. After this burn-in period, we run again the model for one year only, to simplify the examination of the results.


## Results
### Transpiration per leaf area unit (l·m-2)

```{r, echo = FALSE, fig=TRUE, fig.align="center", fig.width=10, fig.height = 15, warning = FALSE, message = FALSE}
plot_several(type = "TranspirationPerLeaf", ylim = c(0,3.5))
```


### Soil water potential (MPa)


```{r, echo = FALSE, fig=TRUE, fig.align="center", fig.width=10, fig.height = 15, message = FALSE, warning = FALSE}
plot_several(type = "SoilPsi", ylim = c(-8,0))
```


### Root-crown water potential (MPa)

```{r, echo = FALSE, fig=TRUE, fig.align="center", fig.width=10, fig.height = 15, message = FALSE, warning = FALSE}
plot_several(type = "RootPsi", ylim = c(-8,0))
```


### Range of leaf water potential (MPa)


```{r, echo = FALSE, fig=TRUE, fig.align="center", fig.width=10, fig.height = 15, warning = FALSE, message = FALSE}
plot_several(type = "LeafPsiRange", ylim = c(-10,0))
```

### Stem percent loss conductance (%)


```{r, echo = FALSE, fig=TRUE, fig.align="center", fig.width=10, fig.height = 15, warning = FALSE, message = FALSE}
plot_several(type = "StemPLC", ylim = c(0,100))
```



## Bibliography

+ De Cáceres, M., M. Mencuccini, N. Martin-StPaul, J.-M. Limousin, L. Coll, R. Poyatos, A. Cabon, V. Granda, A. Forner, F. Valladares, and J. Martínez-Vilalta. 2021. Unravelling the effect of species mixing on water use and drought stress in Mediterranean forests: A modelling approach. Agricultural and Forest Meteorology 296:108233.
+ Schume, H., G. Jost, and H. Hager. 2004. Soil water depletion and recharge patterns in mixed and pure forest stands of European beech and Norway spruce. Journal of Hydrology 289:258–274.